
Pre and Post Bennett Dam Construction Continuous Stations - From Tidyhydat & the Water Survey of CA
variables Calculated: IHA, ice, and Percent Change
Run "Eflows_FUNCTIONS.R" before running this file.


```{r}

library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
library(timetk)
library(Kendall)

```



```{r}
#get stns for this analysis
stn_all <- hy_daily_flows(station_number = c('07KC001', '07HA001', '07NB001', '07DA001')) 

stn_all <- stn_all %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1920/10/1"), as.Date("2022/09/30"), by="day"))

#add water year
stn_all <- addWaterYear(stn_all) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year
stn_all_split <- split(stn_all, stn_all$STATION_NUMBER )
stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear)
stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER")

rm(stn_all_split, stn_all_split_doy)
```


# Data tidying - Pre dam stns 1947-1967
```{r}

stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1956-10-01" & Date < "1967-10-01") 

#delete years that have >14 days missing data using calc_rle function from open-source functions for this project
lst_stns_cln <- split(stn_pre, stn_pre$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist by the station numbers
stns_ready_pre <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_daymonthyear_pre <- stns_ready_pre %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)


```

# IHA calculations - Pre dam stns
```{r}

#make a list of dfs- split by station number
lst_stns_pre <- split(stns_daymonthyear_pre, stns_daymonthyear_pre$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_pre_IHA <- lapply(lst_stns_pre, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_pre <- bind_rows(lst_stns_pre_IHA, .id = "STATION_NUMBER")
View(IHA_pre)

```


# Tidying - Post dam stns 1972-2023
```{r}
stn_post <- stn_all %>%
mutate(Date = as.Date(Date)) %>%
group_by(STATION_NUMBER) %>%
filter(Date >= "1972-10-01" & Date < "2022-10-01") 

#delete years that have >14 days missing data using calc_rle function from open-source functions for this project
lst_stns_cln <- split(stn_post, stn_post$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist by the station numbers
stns_ready_post <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_daymonthyear_post <- stns_ready_post %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

# Calculate IHA - Post dam stns

```{r}
#make a list of dfs- split by station number
lst_stns_post <- split(stns_daymonthyear_post, stns_daymonthyear_post$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_post_IHA <- lapply(lst_stns_post, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_post <- bind_rows(lst_stns_post_IHA, .id = "STATION_NUMBER")
View(IHA_post)


```

#Tidying - prep for ice variables, pre and post
```{r}

#Date must be in yyyy-mm-dd format for ice variables- so re-run tidying code for pre and post dfs without the line that mutates to dd-mm-yyyy format. 

stns_yearmonthday_pre <- stns_ready_pre %>%
  mutate(Date = as.Date(Date)) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

stns_yearmonthday_post <- stns_ready_post %>%
  mutate(Date = as.Date(Date)) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

# Calculate ice variables - pre
```{r}
#make a list of dfs- split by station number
stn_pre_split <- split(stns_yearmonthday_pre, stns_yearmonthday_pre$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_pre_g1 <- lapply(stn_pre_split, Group_1_ice_cover)
stn_pre_g2 <- lapply(stn_pre_split, Group_2_freeze_thaw)
stn_pre_g3 <- lapply(stn_pre_split, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_pre_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_pre_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_pre_g3, .id = "STATION_NUMBER")

#df_final <- cbind.data.frame(output1, output2, output3)
```


```{r}

#Deal with each output df having different lengths and not being able to combine them

#find max and min years for each result df
min_yro1 <- min(output1$waterYear)
max_yr01 <- max(output1$waterYear)

min_yro2 <- min(output2$waterYear)
max_yr02 <- max(output2$waterYear)

min_yro3 <- min(output3$waterYear)
max_yr03 <- max(output3$waterYear)

#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = min_yro1:max_yr01) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = min_yro2:max_yr02) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = min_yro3:max_yr03) #fills NA for values for missing years 

df_ice_pre <- cbind(output1_ex, output2_ex, output3_ex)

df_ice_pre_final <- df_ice_pre[!duplicated(as.list(df_ice_pre))] #removed duplicated columns

View(df_ice_pre_final)
#write.csv(df_ice_pre_final, "ice1.csv")

```

#Calculate ice variables - post

```{r}
#make a list of dfs- split by station number
stn_post_split <- split(stns_yearmonthday_post, stns_yearmonthday_post$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_post_g1 <- lapply(stn_post_split, Group_1_ice_cover)
stn_post_g2 <- lapply(stn_post_split, Group_2_freeze_thaw)
stn_post_g3 <- lapply(stn_post_split, Group_3_freshet)

#all output lists into a df
output1 <- bind_rows(stn_post_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_post_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_post_g3, .id = "STATION_NUMBER")

```


```{r}
#Deal with each output df having different lengths and not being able to combine them

#find max and min years for each result df
min_yr01 <- min(output1$waterYear)
max_yr01 <- max(output1$waterYear)

min_yr02 <- min(output2$waterYear)
max_yr02 <- max(output2$waterYear)

min_yr03 <- min(output3$waterYear)
max_yr03 <- max(output3$waterYear)


#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = min_yr01:max_yr01) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = min_yr02:max_yr02) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = min_yr03:max_yr03) #fills NA for values for missing years 

df_ice_post <- cbind(output1_ex, output2_ex, output3_ex)

df_ice_post_final <- df_ice_post[!duplicated(as.list(df_ice_post))] #removed duplicated columns

View(df_ice_post_final)

#write.csv(df_ice_post_final, "df_ice_post_final.csv")

```

#Percent change, IHA variables 
```{r}
#data prep, IHA-pre
#calc the medians value for each stn number for the IHA-pre df
IHA_medians_pre <- IHA_pre %>%
  group_by(STATION_NUMBER) %>%
  summarise(across(October:Reversals, median)) #Change to Jan if doing cal year

#data prep, IHA-post
#find the min and max year for the IHA_post df so there is an even number of years for each station

min_yr <-min(IHA_post$Year)
max_yr <- max(IHA_post$Year)

IHA_pst <- IHA_post %>%
  mutate(Year = as.integer(Year)) %>%
  complete(STATION_NUMBER, Year = min_yr:max_yr) #fills NA for values for missing years 

```

#Calculate percent change, IHA
```{r}

df_lst <- list()

for (i in unique(IHA_medians_pre$STATION_NUMBER)) {

calc <- calc_percent_change(IHA_medians_pre, IHA_pst, stn = i, year_col = "Year") #calc perc. change for all stns
calc$i <- i #add i (stn # to the calc df)
df_lst[[i]] <- calc #add calc df to the list
}

final_pc_IHA <- bind_rows(df_lst) %>%
 rename("Station_Number" = "i") 

View(final_pc_IHA)

#write.csv(final_pc_IHA, "PC_IHA.csv")

```

#Percent change, Ice variables 

```{r}

#get rid of dates and rely on doy to figure out time changes (cant avg a date)

ice_pre <- df_ice_pre_final %>%
  select(-c(Freeze_Date, Thaw_Date, Station_Number, Freshet_Date)) %>%
  rename(Freshet_Doy = Day_of_year)

ice_post <- df_ice_post_final %>% 
  select(-c(Freeze_Date, Thaw_Date, Station_Number, Freshet_Date)) %>%
  rename(Freshet_Doy = Day_of_year) 

#data prep, ice-pre
#calc the median for each stn number for the ice-pre df
ice_medians_pre <- ice_pre %>%
  group_by(STATION_NUMBER) %>%
  na.omit() %>%
  summarise(across(ice_coverage_wy:Freshet_Flow, median)) #Change to Jan if doing cal year

#data prep, ice-post
#find the min and max year for the ice_post df so there is an even number of years for each station

min_yr <-min(ice_post$waterYear)
max_yr <- max(ice_post$waterYear)

ice_pst <- ice_post %>%
  mutate(waterYear = as.integer(waterYear)) %>%
  complete(STATION_NUMBER, waterYear = min_yr:max_yr) #fills NA for values for missing years 

```

Calculate percent change
```{r}
df_lst <- list()

for (i in unique(ice_medians_pre$STATION_NUMBER)) {

calc <- calc_percent_change(ice_medians_pre, ice_pst, stn = i, year_col = "waterYear") #calc perc. change for all stns
calc$i <- i #add i (stn # to the calc df)
df_lst[[i]] <- calc #add calc df to the list
}

final_pc_ice <- bind_rows(df_lst) %>%
 rename("Station_Number" = "i") 

View(final_pc_ice)

#write.csv(final_pc_ice, "PC_ice.csv")
```


#Mann-Kendall Trend Analysis - ice variables 
```{r}

DY <- calc_MK(df_ice_post_final, parameter = Day_of_year, start = 1973) %>%
  rename("Result_Freshet_Onset" = 'P_Value' )

icecover <- calc_MK(df_ice_post_final, parameter = ice_coverage_wy, start = 1973)%>%
  rename("Result_ice_cover" = 'P_Value' )

frdoy <- calc_MK(df_ice_post_final, parameter = Freeze_DOY, start = 1973) %>%
  rename("Result_freeze_doy" = 'P_Value' )

thawdoy <- calc_MK(df_ice_post_final, parameter = Thaw_DOY, start = 1973)%>%
  rename("Result_thaw_doy" = 'P_Value' )

result <- cbind(DY, icecover, frdoy, thawdoy)
MK_result <- result[!duplicated(as.list(result))] #removed duplicated columns

MK_result_final_ice <- MK_result %>% mutate_if(is.numeric, ~round(., 3)) #round 
View(MK_result_final_ice)
#write.csv(MK_result_final_ice, "MKres.csv")

```

#Mann-Kendall Trend Analysis - IHA variables 

```{r}
#Expand IHA_post df to have all years for all stations so they can be used in the MK function

for (i in unique(IHA_post$STATION_NUMBER)) {
  min_yr <- min(IHA_post$Year)
  max_yr <- max(IHA_post$Year)
  IHA_post_expand <- IHA_post %>%
    mutate(Year = as.integer(Year)) %>%
    complete(STATION_NUMBER, Year = min_yr:max_yr)
  return(IHA_post_expand)
}

IHA_post_expand <- IHA_post_expand %>%
rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min",
          "High_pulse_number" = "High pulse number",
          "Low_pulse_number" = "Low pulse number",
          "High_pulse_length" = "High pulse length",
          "Low_pulse_length" = "Low pulse length")

```


```{r}
#Use calc_Mk to do a MK for all Stations for selected variables

Reversals <- calc_MK(IHA_post_expand, parameter = Reversals, start = 1973) %>%
  rename("Result_Reversals" = 'P_Value')

One_Day_Min <- calc_MK(IHA_post_expand, parameter = One_Day_Min , start = 1973) %>% rename("Result_One_Day_Min" = 'P_Value')

One_Day_Max <-  calc_MK(IHA_post_expand, parameter = One_Day_Max , start = 1973) %>% rename("Result_1_Day_Max" = 'P_Value')

Seven_Day_Min <- calc_MK(IHA_post_expand, parameter = Seven_Day_Min , start = 1973) %>% rename("Result_7_Day_Min" = 'P_Value')

Seven_Day_Max <-  calc_MK(IHA_post_expand, parameter = Seven_Day_Max , start = 1973) %>% rename("Result_7_Day_Max" = 'P_Value')

Thirty_Day_Min <- calc_MK(IHA_post_expand, parameter = Thirty_Day_Min , start = 1973) %>% rename("Result_30_Day_Min" = 'P_Value')

Thirty_Day_Max <-  calc_MK(IHA_post_expand, parameter = Thirty_Day_Max , start = 1973) %>% rename("Result_30_Day_Max" = 'P_Value')

High_pulse_number <-  calc_MK(IHA_post_expand, parameter = High_pulse_number , start = 1973) %>% rename("Result_High_pulse_number" = 'P_Value')

Low_pulse_number <-  calc_MK(IHA_post_expand, parameter = Low_pulse_number , start = 1973) %>% rename("Result_Low_pulse_number" = 'P_Value')

High_pulse_length <-  calc_MK(IHA_post_expand, parameter = High_pulse_length , start = 1973) %>% rename("Result_High_pulse_length" = 'P_Value')

Low_pulse_length <-  calc_MK(IHA_post_expand, parameter = Low_pulse_length , start = 1973) %>% rename("Result_Low_pulse_length" = 'P_Value')

result_MK_IHA <- cbind(Reversals, One_Day_Min, One_Day_Max, Seven_Day_Min, Seven_Day_Max, Thirty_Day_Min, Thirty_Day_Max, High_pulse_number, Low_pulse_number, High_pulse_length, Low_pulse_length )

MK_result_final <- result_MK_IHA[!duplicated(as.list(result_MK_IHA))] #removed duplicated columns

#MK_result_final_IHA <- result %>% mutate_if(is.numeric, ~round(., 3)) #round 

View(result_MK_IHA)

#write.csv(result_MK_IHA, "IHAresult1.csv")
```



```{r}

```

