
Pre and Post Bennett Dam Construction Continuous Stations - From Tidyhydat & the Water Survey of CA
Variables Calculated: IHA, ICE, and Percent Change
Run "Eflows_FUNCTIONS.R" before running this file.


```{r}

library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
library(timetk)

```



```{r}
#get stns for this analysis
stn_all <- hy_daily_flows(station_number = c('07KC001', '07HA001', '07NB001', '07DA001')) 

stn_all <- stn_all %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1920/10/1"), as.Date("2022/09/30"), by="day"))

#add water year
stn_all <- addWaterYear(stn_all) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year
stn_all_split <- split(stn_all, stn_all$STATION_NUMBER )
stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear)
stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER")

rm(stn_all_split, stn_all_split_doy)
```


# Pre dam stns 1947-1967
```{r}

stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1956-10-01" & Date < "1967-10-01") 

#print out rows with > 14 Nas in a row
lst_stns_pre <- split(stn_pre, stn_pre$STATION_NUMBER )
lst_stns_pre_rle <- lapply(lst_stns_pre, calc_rle)

for (i in lst_stns_pre_rle) {
  for (j in i)
    stn <- unique(i$STATION_NUMBER)
    wy <- unique(i$waterYear)
    print("delete the following station # and water year:")
    print(stn)
    print(wy)
}


#delete the rows with >14 NAs, printed from above, and fill any values that are missing with preceding values 
stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1956-10-01" & Date < "1967-10-01") %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1964)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 1959)) %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

# Pre dam stns IHA
```{r}

#make a list of dfs- split by station number
lst_stns_pre <- split(stn_pre, stn_pre$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_pre_IHA <- lapply(lst_stns_pre, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_pre <- bind_rows(lst_stns_pre_IHA, .id = "STATION_NUMBER")
View(IHA_pre)


```


# Post dam stns 1972-2023
```{r}
stn_post <- stn_all %>%
mutate(Date = as.Date(Date)) %>%
group_by(STATION_NUMBER) %>%
filter(Date >= "1972-10-01" & Date < "2022-10-01") %>%
  select(-c(Symbol))


#print out rows with >14 Nas in a row, and then manually delete that year
lst_stns_post<- split(stn_post, stn_post$STATION_NUMBER )
lst_stns_post_rle <- lapply(lst_stns_post, calc_rle)

for (i in lst_stns_post_rle) {
  for (j in i)
    stn <- unique(i$STATION_NUMBER)
    wy <- unique(i$waterYear)
    print("delete the following station # and water year:")
    print(stn)
    print(wy)
}


#delete the printed values from above and fill any values that are missing with preceding values 

stn_post <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1972-10-01" & Date < "2022-10-01") %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 2022)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2022)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2008)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2009)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2022)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2022)) %>%
  select(-c(Symbol)) %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

# Post dam stns IHA

```{r}
#make a list of dfs- split by station number
lst_stns_post <- split(stn_post, stn_post$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_post_IHA <- lapply(lst_stns_post, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_post <- bind_rows(lst_stns_post_IHA, .id = "STATION_NUMBER")
View(IHA_post)


```


```{r}

#Date must be in yyyy-mm-dd format for ice variables- so re-run tidying code for pre and post dfs without the line that mutates to dd-mm-yyyy format. 
#don't forget to keep the symbol column

stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1956-10-01" & Date < "1967-10-01") %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1964)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 1959)) %>%
  #mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)


stn_post <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1972-10-01" & Date < "2022-10-01") %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 2022)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2022)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2008)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2009)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2022)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2022)) %>%
  #mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

#ICE Variables - pre
```{r}
#make a list of dfs- split by station number
stn_pre_split <- split(stn_pre, stn_pre$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_pre_g1 <- lapply(stn_pre_split, Group_1_ice_cover)
stn_pre_g2 <- lapply(stn_pre_split, Group_2_freeze_thaw)
stn_pre_g3 <- lapply(stn_pre_split, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_pre_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_pre_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_pre_g3, .id = "STATION_NUMBER")

#df_final <- cbind.data.frame(output1, output2, output3)
```


```{r}

#Deal with each output df having different lengths and not being able to combine them

#find max and min years for each result df
minyro1 <- min(output1$waterYear)
maxyr01 <- max(output1$waterYear)

minyro2 <- min(output2$waterYear)
maxyr02 <- max(output2$waterYear)

minyro3 <- min(output3$waterYear)
maxyr03 <- max(output3$waterYear)


#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro1:maxyr01) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro2:maxyr02) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro3:maxyr03) #fills NA for values for missing years 

df_ICE_pre <- cbind(output1_ex, output2_ex, output3_ex)

df_ICE_pre_final <- df_ICE_pre[!duplicated(as.list(df_ICE_pre))] #removed duplicated columns

View(df_ICE_pre_final)

```

#ICE Variables - post

```{r}
#make a list of dfs- split by station number
stn_post_split <- split(stn_post, stn_post$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_post_g1 <- lapply(stn_post_split, Group_1_ice_cover)
stn_post_g2 <- lapply(stn_post_split, Group_2_freeze_thaw)
stn_post_g3 <- lapply(stn_post_split, Group_3_freshet)

#all output lists into a df
output1 <- bind_rows(stn_post_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_post_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_post_g3, .id = "STATION_NUMBER")

```


```{r}
#Deal with each output df having different lengths and not being able to combine them

#find max and min years for each result df
minyro1 <- min(output1$waterYear)
maxyr01 <- max(output1$waterYear)

minyro2 <- min(output2$waterYear)
maxyr02 <- max(output2$waterYear)

minyro3 <- min(output3$waterYear)
maxyr03 <- max(output3$waterYear)


#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro1:maxyr01) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro2:maxyr02) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro3:maxyr03) #fills NA for values for missing years 

df_ICE_post <- cbind(output1_ex, output2_ex, output3_ex)

df_ICE_post_final <- df_ICE_post[!duplicated(as.list(df_ICE_post))] #removed duplicated columns

View(df_ICE_post_final)

```

#Percent change, IHA Variables 
```{r}
#data prep, IHA-pre
#calc the medians value for each stn number for the IHA-pre df
IHA_medians_pre <- IHA_pre %>%
  group_by(STATION_NUMBER) %>%
  summarise(across(October:Reversals, median)) #Change to Jan if doing cal year

#data prep, IHA-post
#find the min and max year for the IHA_post df so there is an even number of years for each station

minyr <-min(IHA_post$Year)
maxyr <- max(IHA_post$Year)

IHA_pst <- IHA_post %>%
  mutate(Year = as.integer(Year)) %>%
  complete(STATION_NUMBER, Year = minyr:maxyr) #fills NA for values for missing years 

```

Calculate percent change
```{r}

df_lst <- list()

for (i in unique(IHA_medians_pre$STATION_NUMBER)) {

calc <- calc_percent_change(IHA_medians_pre, IHA_pst, stn = i, year_col = "Year") #calc perc. change for all stns
calc$i <- i #add i (stn # to the calc df)
df_lst[[i]] <- calc #add calc df to the list
}

final_pc_IHA <- bind_rows(df_lst) %>%
 rename("Station_Number" = "i") 

View(final_pc_IHA)

#write.csv(final_pc_IHA, "PC_IHA.csv")

```

#Percent change, Ice Variables 

```{r}

#get rid of dates and rely on doy to figure out time changes (cant avg a date)

ICE_pre <- df_ICE_pre_final %>%
  select(-c(Freeze_Date, Thaw_Date, Station_Number, Freshet_Date)) %>%
  rename(Freshet_Doy = Day_of_year)

ICE_post <- df_ICE_post_final %>% 
  select(-c(Freeze_Date, Thaw_Date, Station_Number, Freshet_Date)) %>%
  rename(Freshet_Doy = Day_of_year) 

#data prep, ICE-pre
#calc the median for each stn number for the ICE-pre df
ICE_medians_pre <- ICE_pre %>%
  group_by(STATION_NUMBER) %>%
  na.omit() %>%
  summarise(across(Ice_coverage_wy:Freshet_Flow, median)) #Change to Jan if doing cal year

#data prep, ICE-post
#find the min and max year for the ICE_post df so there is an even number of years for each station

minyr <-min(ICE_post$waterYear)
maxyr <- max(ICE_post$waterYear)

ICE_pst <- ICE_post %>%
  mutate(waterYear = as.integer(waterYear)) %>%
  complete(STATION_NUMBER, waterYear = minyr:maxyr) #fills NA for values for missing years 

```

Calculate percent change
```{r}
df_lst <- list()

for (i in unique(ICE_medians_pre$STATION_NUMBER)) {

calc <- calc_percent_change(ICE_medians_pre, ICE_pst, stn = i, year_col = "waterYear") #calc perc. change for all stns
calc$i <- i #add i (stn # to the calc df)
df_lst[[i]] <- calc #add calc df to the list
}

final_pc_ICE <- bind_rows(df_lst) %>%
 rename("Station_Number" = "i") 

View(final_pc_ICE)

#write.csv(final_pc_ICE, "PC_ICE.csv")
```



```{r}



```


```{r}

```



```{r}

```

