
Pre and Post Climate Change Continuous Stations - From Tidyhydat & the Water Survey of CA
Variables Calculated: IHA, ICE, and Percent Change
Run "Eflows_FUNCTIONS.R" before running this file.


```{r}
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)

```



```{r}

#get stns for this analysis
stn_all <- hy_daily_flows(station_number = c('07AF002', '07BC002', '07BE001',  '07CD001', '07DA001', '07EC003', '07EE007', '07EF001', '07FB001', '07FD001', '07FD002', '07GE001','07GH002', '07GJ001', '07HA001', '07HC001', '07KC001', '07MA003', '07MB001', '07NB001'))

stn_all <- stn_all %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1920/10/1"), as.Date("2022/09/30"), by="day")) 


#add water year
stn_all <- addWaterYear(stn_all) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year
stn_all_split <- split(stn_all, stn_all$STATION_NUMBER )
stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear)
stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER")

rm(stn_all_split, stn_all_split_doy)

```

# Pre climate change stns 1975-2005
```{r}

stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1975-10-01" & Date < "2005-10-01")

#print out rows with >14 Nas in a row, and then manually delete that year
lst_stns_pre <- split(stn_pre, stn_pre$STATION_NUMBER )
lst_stns_pre_rle <- lapply(lst_stns_pre, calc_rle)

for (i in lst_stns_pre_rle) {
  for (j in i)
    stn <- unique(i$STATION_NUMBER)
    wy <- unique(i$waterYear)
    print("delete the following station # and water year:")
    print(stn)
    print(wy)
}


#delete the printed values from above and fill any values that are missing with preceding values 
stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1975-10-01" & Date < "2005-10-01") %>%
  filter(!c(STATION_NUMBER == "07EC003" & waterYear == 1976)) %>%
  filter(!c(STATION_NUMBER == "07EE007" & waterYear == 2005)) %>%
  filter(!c(STATION_NUMBER == "07MA003" & waterYear == 1989)) %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

# Pre climate change stns IHA
```{r}

#make a list of dfs- split by station number
lst_stns_pre <- split(stn_pre, stn_pre$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_pre_IHA <- lapply(lst_stns_pre, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_pre <- bind_rows(lst_stns_pre_IHA, .id = "STATION_NUMBER")
View(IHA_pre)


```


# Post climate change stns 2006-2023
```{r}

stn_post <- stn_all %>%
mutate(Date = as.Date(Date)) %>%
group_by(STATION_NUMBER) %>%
filter(Date >= "2006-10-01" & Date < "2021-10-01") 


#print out rows with >14 Nas in a row, and then manually delete that year
lst_stns_post<- split(stn_post, stn_post$STATION_NUMBER )
lst_stns_post_rle <- lapply(lst_stns_post, calc_rle)

for (i in lst_stns_post_rle) {
  for (j in i)
    stn <- unique(i$STATION_NUMBER)
    wy <- unique(i$waterYear)
    print("delete the following station # and water year:")
    print(stn)
    print(wy)
}



#delete the printed values from above and fill any values that are missing with preceding values 
stn_post <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "2006-10-01" & Date < "2021-10-01") %>%
  filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2017)) %>%
    filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2018)) %>%
    filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07EE007" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07EE007" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2019)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2011)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2012)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2015)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2016)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2017)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07GJ001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2016)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2017)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2008)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2009)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2021)) %>%
   filter(!c(STATION_NUMBER == "07MA003" & waterYear == 2021)) %>%
   filter(!c(STATION_NUMBER == "07MB001" & waterYear == 2021)) %>%
   filter(!c(STATION_NUMBER == "07MB001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
   filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
   filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)


```

# Post climate change stns IHA

```{r}
#make a list of dfs- split by station number
lst_stns_post <- split(stn_post, stn_post$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_post_IHA <- lapply(lst_stns_post, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
IHA_post <- bind_rows(lst_stns_post_IHA, .id = "STATION_NUMBER")

View(IHA_post)

```


# Pre climate change stns Ice Variables
```{r}
#Date must be in yyyy-mm-dd format for ice variables- so re-run tidying code for pre and post dfs without the line that mutates to dd-mm-yyyy format. 
#don't forget to keep the symbol column

stn_pre <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "1975-10-01" & Date < "2005-10-01") %>%
  filter(!c(STATION_NUMBER == "07EC003" & waterYear == 1976)) %>%
  filter(!c(STATION_NUMBER == "07EE007" & waterYear == 2005)) %>%
  filter(!c(STATION_NUMBER == "07MA003" & waterYear == 1989)) %>%
 #mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)


stn_post <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(Date >= "2006-10-01" & Date < "2021-10-01") %>%
  filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2017)) %>%
    filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2018)) %>%
    filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07EE007" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07EE007" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2019)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2011)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2012)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2015)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2016)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2017)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07GE001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07GJ001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2016)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2017)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07HC001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2008)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2009)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2021)) %>%
   filter(!c(STATION_NUMBER == "07MA003" & waterYear == 2021)) %>%
   filter(!c(STATION_NUMBER == "07MB001" & waterYear == 2021)) %>%
   filter(!c(STATION_NUMBER == "07MB001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
   filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
   filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  #mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```


#ICE Variables - pre
```{r}

#make a list of dfs- split by station number
stn_pre_split <- split(stn_pre, stn_pre$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_pre_g1 <- lapply(stn_pre_split, Group_1_ice_cover)
stn_pre_g2 <- lapply(stn_pre_split, Group_2_freeze_thaw)
stn_pre_g3 <- lapply(stn_pre_split, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_pre_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_pre_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_pre_g3, .id = "STATION_NUMBER")

#df_final <- cbind.data.frame(output1, output2, output3)

```

#ICE Variables - post

```{r}
#make a list of dfs- split by station number
stn_post_split <- split(stn_post, stn_post$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_post_g1 <- lapply(stn_post_split, Group_1_ice_cover)
stn_post_g2 <- lapply(stn_post_split, Group_2_freeze_thaw)
stn_post_g3 <- lapply(stn_post_split, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_post_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_post_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_post_g3, .id = "STATION_NUMBER")

#df_final <- cbind.data.frame(output1, output2, output3)

```

```{r}

```

```{r}

```


```{r}

```


