
Long term, Continuous Stations - From Tidyhydat & the Water Survey of CA.

Variables Calculated: IHA, ICE
Run "Eflows_FUNCTIONS.R" before running this file.


```{r}

library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
library(timetk)
library(forecast)
library(Kendall)
library(trend)
library(outliers)


```


```{r}
#get stns for this analysis
stn_all <- hy_daily_flows(station_number = c("07BE001", "07BC002", "07CD001", "07DA001", "07GH002", "07GJ001", "07HA001", "07NB001", '07AF002')) 

stn_all <- stn_all %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1920/10/1"), as.Date("2022/09/30"), by="day"))

#add water year
stn_all <- addWaterYear(stn_all) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year
stn_all_split <- split(stn_all, stn_all$STATION_NUMBER )
stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear)
stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER")

rm(stn_all_split, stn_all_split_doy)
```


# Tidy Stations
```{r}

stn_cln <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1955) %>%
  filter(waterYear != 2022)

#print out rows with > 14 Nas in a row
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

for (i in lst_stns_cln_rle) {
  for (j in i)
    stn <- unique(i$STATION_NUMBER)
    wy <- unique(i$waterYear)
    print("delete the following station # and water year:")
    print(stn)
    print(wy)
}


#delete the rows with >14 NAs, printed from above, and fill any values that are missing with preceding values 
stn_cln <- stn_all %>%
    filter(waterYear > 1955) %>%
  filter(waterYear != 2022) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1959)) %>%        
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07GJ001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1964)) %>%
  filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2017)) %>%
  filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1960)) %>%  
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1961)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07CD001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07CD001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07CD001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1964)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1970)) %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

# IHA calcs
```{r}

#make a list of dfs- split by station number
lst_stns <- split(stn_cln, stn_cln$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")
View(IHA)

#write.csv(IHA, "IHAlongterm.csv")

```


```{r}

#Date must be in yyyy-mm-dd format for ice variables- so re-run tidying code for pre and post dfs without the line that mutates to dd-mm-yyyy format. 
#don't forget to keep the symbol column
stn_cln <- stn_all %>%
  filter(waterYear > 1955) %>%
  filter(waterYear != 2022) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 1959)) %>%        
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07HA001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07GJ001" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07GH002" & waterYear == 1964)) %>%
  filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2017)) %>%
  filter(!c(STATION_NUMBER == "07AF002" & waterYear == 2018)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1960)) %>%  
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 1961)) %>%
  filter(!c(STATION_NUMBER == "07BC002" & waterYear == 2021)) %>%
  filter(!c(STATION_NUMBER == "07CD001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07CD001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07CD001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1956)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1957)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1958)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1959)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1964)) %>%
  filter(!c(STATION_NUMBER == "07DA001" & waterYear == 1970)) %>%
  #mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

#ICE Variables - pre
```{r}
#make a list of dfs- split by station number
lst_stns <- split(stn_cln, stn_cln$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_g1 <- lapply(lst_stns, Group_1_ice_cover)
stn_g2 <- lapply(lst_stns, Group_2_freeze_thaw)
stn_g3 <- lapply(lst_stns, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_g3, .id = "STATION_NUMBER")

```


```{r}
#Deal with each output df having different lengths and not being able to combine them

#find max and min years for each result df
minyro1 <- min(output1$waterYear)
maxyr01 <- max(output1$waterYear)

minyro2 <- min(output2$waterYear)
maxyr02 <- max(output2$waterYear)

minyro3 <- min(output3$waterYear)
maxyr03 <- max(output3$waterYear)


#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro1:maxyr01) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro2:maxyr02) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro3:maxyr03) #fills NA for values for missing years 

df_ICE <- cbind(output1_ex, output2_ex, output3_ex)

df_ICE_final <- df_ICE[!duplicated(as.list(df_ICE))] #removed duplicated columns

View(df_ICE_final)

#write.csv(df_ICE_final, "ICE.csv")
```

#Mann-Kendall Trend Analysis - Ice Variables

```{r}

DY <- calc_MK(df_ICE_final, parameter = Day_of_year, start = 1956) %>%
  rename("Result_Freshet_Onset" = 'P_Value' )

Icecover <- calc_MK(df_ICE_final, parameter = Ice_coverage_wy, start = 1956)%>%
  rename("Result_Ice_cover" = 'P_Value' )

frdoy <- calc_MK(df_ICE_final, parameter = Freeze_DOY, start = 1956) %>%
  rename("Result_freeze_doy" = 'P_Value' )

thawdoy <- calc_MK(df_ICE_final, parameter = Thaw_DOY, start = 1956)%>%
  rename("Result_thaw_doy" = 'P_Value' )

result <- cbind(DY, Icecover, frdoy, thawdoy)
MK_result_final <- result[!duplicated(as.list(result))] #removed duplicated columns

MK_result_final_Ice <- MK_result_final %>% mutate_if(is.numeric, ~round(., 3)) #round 

View(MK_result_final_Ice)
#write.csv(MK_result_final_Ice, "MKres.csv")

STOP
```

#Mann-Kendall Trend Analysis - IHA Variables 

```{r}
#Expand IHA df to have all years for all stations so they can be used in the MK function

for (i in unique(IHA$STATION_NUMBER)) {
  minyr <- min(IHA$Year)
  maxyr <- max(IHA$Year)
  IHA_post_expand <- IHA %>%
    mutate(Year = as.integer(Year)) %>%
    complete(STATION_NUMBER, Year = minyr:maxyr)
  return(IHA_post_expand)
}

IHA_post_expand <- IHA_post_expand %>%
rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min",
          "High_pulse_number" = "High pulse number",
          "Low_pulse_number" = "Low pulse number",
          "High_pulse_length" = "High pulse length",
          "Low_pulse_length" = "Low pulse length")
```


```{r}
#Use calc_Mk to do a MK for all Stations for selected variables

Reversals <- calc_MK(IHA_post_expand, parameter = Reversals, start = 1956) %>%
  rename("Result_Reversals" = 'P_Value')

One_Day_Min <- calc_MK(IHA_post_expand, parameter = One_Day_Min , start = 1956) %>% rename("Result_One_Day_Min" = 'P_Value')

One_Day_Max <-  calc_MK(IHA_post_expand, parameter = One_Day_Max , start = 1956) %>% rename("Result_1_Day_Max" = 'P_Value')

Seven_Day_Min <- calc_MK(IHA_post_expand, parameter = Seven_Day_Min , start = 1956) %>% rename("Result_7_Day_Min" = 'P_Value')

Seven_Day_Max <-  calc_MK(IHA_post_expand, parameter = Seven_Day_Max , start = 1956) %>% rename("Result_7_Day_Max" = 'P_Value')

Thirty_Day_Min <- calc_MK(IHA_post_expand, parameter = Thirty_Day_Min , start = 1956) %>% rename("Result_30_Day_Min" = 'P_Value')

Thirty_Day_Max <-  calc_MK(IHA_post_expand, parameter = Thirty_Day_Max , start = 1956) %>% rename("Result_30_Day_Max" = 'P_Value')

High_pulse_number <-  calc_MK(IHA_post_expand, parameter = High_pulse_number , start = 1956) %>% rename("Result_High_pulse_number" = 'P_Value')

Low_pulse_number <-  calc_MK(IHA_post_expand, parameter = Low_pulse_number , start = 1956) %>% rename("Result_Low_pulse_number" = 'P_Value')

High_pulse_length <-  calc_MK(IHA_post_expand, parameter = High_pulse_length , start = 1956) %>% rename("Result_High_pulse_length" = 'P_Value')

Low_pulse_length <-  calc_MK(IHA_post_expand, parameter = Low_pulse_length , start = 1956) %>% rename("Result_Low_pulse_length" = 'P_Value')

result_MK_IHA <- cbind(Reversals, One_Day_Min, One_Day_Max, Seven_Day_Min, Seven_Day_Max, Thirty_Day_Min, Thirty_Day_Max, High_pulse_number, Low_pulse_number, High_pulse_length, Low_pulse_length )

View(result_MK_IHA)

#write.csv(result_MK_IHA, "IHAresult5.csv")

```

#MK/ARIMA Trend Analysis - working and testing ARIMA model
( ignore and use code above)

```{r}
stn_07BE001 <- stn_07BE001 %>%
  dplyr::filter(lubridate::year(Date) > 1955) %>%
  dplyr::filter(lubridate::year(Date) < 2022) 

TS_07BE001 <- ts(stn_07BE001$Value, 
        frequency = 365, #daily data (no NAs present, and full seq of years are present)
         start = c(1956,1)) #1 is day of the year that data starts ie Jan 1

plot.ts(TS_07BE001) #additive model could work, bc fluctuations in the data are more or less constant over time. 

```

```{r}
TS_07BE001_timecompon <- decompose(TS_07BE001) #decompose into seasonal comp, irreg, trend comp.

plot(TS_07BE001_timecompon) #doesn't look like there is a trend

#subtract the estimated seasonal comp from the original time series to get rid of seasonal component 
TS_07BE001_seasonadj <- TS_07BE001 - TS_07BE001_timecompon$seasonal

plot(TS_07BE001_seasonadj) #this contains the trend component and the irregular component
```


```{r}
## Mann-Kendall test

MK <- MannKendall(TS_07BE001) #p = .70
MK_seas <- MannKendall(TS_07BE001_seasonadj) #p = .2
summary(MK)

```


```{r}
#try Seasonal MK using Monthly data
##NOTE: will be off by one year. Need to re-do the IHA calcs for the calendar year and not WY.

#get df of monthly statistics using IHA calcs
mnth_07BE001 <- IHA %>%
  filter(STATION_NUMBER == "07BE001") %>%
  select(Year:September) %>% 
  pivot_longer(c('October', 'November', 'December', 'January', 'February', 'March', 'April', 'May', 'June', 'July', 'August', 'September'), names_to = "Month", values_to = "Value") %>%
  mutate(Month = recode(Month, 
                  January = 01,
                        February = 02,
                        March = 03,
                        April = 04,
                        May = 05,
                        June = 06,
                        July = 07,
                        August = 08,
                        September = 09,
                        October = 10,
                        November = 11,
                        December = 12))


TS_mnth <- ts(mnth_07BE001$Value, frequency = 12, start = c(1956, 10))
```


```{r}

#decompose monthly data and remove seasonal component
TS_mnth_decomp <- decompose(TS_mnth)
TS_mnth_adj <- TS_mnth - TS_mnth_decomp$seasonal

##DECOMPOSED TS
#MK on decomposed data 
MannKendall(TS_mnth_adj) #.35
SeasonalMannKendall(TS_mnth_adj) #p = .78

#Sen's slope and Pettitt's test on decomposed data
sens.slope(TS_mnth_adj) 
sea.sens.slope(TS_mnth_adj)
pettitt.test(TS_mnth_adj) #get diff pvalues??? p = 00.00129

#NON DECOMPOSED TS
#MK on non decomposed data 
MannKendall(TS_mnth) #p = .875
SeasonalMannKendall(TS_mnth) #p= .79 can do: smk.test(TS_mnth)

#Sen's slope and Pettitt's test on non-decomposed data 
sens.slope(TS_mnth) 
sea.sens.slope(TS_mnth)
pettitt.test(TS_mnth) #p= .6315 

#use original df and input "K" to get the date of the probably change point.
mnth_07BE001[667,] #suggests change point in 2011, but p = .6315

?SeasonalMannKendall()
#score is computed separately for each month 
#unsure if this works for our data bc we'd expect more flow in winter and less flow in summer. Trend is not monotonic. 
#can we deal with the lack of monotonic trends by doing a MK only across summer months, or winter months, to see a monotonic trend? Ie less flow in summer? And more in winter? 

#to do this, would need to figure out how to make a time series obj just for a certain month?


```




```{r}
#Try a MK test just for one month: August

library("TTR")

mnth_07BE001_aug <- mnth_07BE001 %>%
  filter(Month == 8)
  
TS_Aug <- ts(mnth_07BE001_aug$Value, frequency = 1, start = c(1956, 1))

MannKendall(TS_Aug) #p = .12
sens.slope(TS_Aug) #m = -1.5
pettitt.test(TS_Aug) #p = .06, change pnt 46
mnth_07BE001_aug[46,] #year = 2001

plot(TS_Aug)

#smoothing with a moving avg (can't decomp seasonally, there is no seasonal component)
TS_Aug_SMA <- SMA(TS_Aug, n = 9)
plot(TS_Aug_SMA)


#looking at outliers 
library(outliers)
grubbs.test(mnth_07BE001_aug$Value) #1989 is an outlier

#remove outlier then re-try test

mnth_07BE001_aug_rmout <- mnth_07BE001 %>%
  filter(Month == 8) %>%
  filter(Year != 1989)

TS_Aug_rm <- ts(mnth_07BE001_aug_rmout$Value, frequency = 1, start = c(1956,1))

MannKendall(TS_Aug_rm) #p = .12
sens.slope(TS_Aug_rm) #m = -1.5
pettitt.test(TS_Aug_rm) #p = .07, change pnt 46
mnth_07BE001_aug[46,] #year = 2001

#removing outlier didn't do much 
```

#ARIMA
```{r}
TS_mnth_diff <- diff(TS_mnth, differences = 2)

plot.ts(TS_mnth_diff)

acf(TS_mnth_diff, lag.max = 20) #1,2 exceed sign bounds
acf(TS_mnth_diff, lag.max = 20, plot = FALSE)

pacf(TS_mnth_diff, lag.max = 20)
pacf(TS_mnth_diff, lag.max = 20, plot = FALSE)

library('forecast')
auto.arima(TS_mnth)
TS_mnth_arima <- arima(TS_mnth, order = c(1,1,1) )
TS_mnth_forecast <- forecast(TS_mnth_arima, h = 5)

plot(TS_mnth_forecast)

#are the forecast errors N distributed with mean zero, constant variance 
acf(TS_mnth_forecast$residuals, lag.max = 20)
Box.test(TS_mnth_forecast$residuals, lag = 20, type = "Ljung-Box") #high evidence for non-zero correlations in the forecast errors 

plot.ts(TS_mnth_forecast$residuals)
plotForecastErrors(TS_mnth_forecast$residuals)

hist(TS_mnth_forecast$residuals)

d <- density(TS_mnth_forecast$residuals)
plot(d)

```

#try MK with yearly timing of freshet for a station on the PEace or Slave River
```{r}

stn_07HA001 <- df_ICE_final %>%
  filter(STATION_NUMBER == "07HA001") %>%
  filter(waterYear > 1958) %>%
  filter(waterYear < 2021)

TS_07HA001 <- ts(stn_07HA001$Day_of_year, 
        frequency = 1,
         start = c(1959,1))

plot.ts(TS_07HA001) 
MK <- MannKendall(TS_07HA001) #p = .016
sens.slope(TS_07HA001) #won't work with NAS
pettitt.test(TS_07HA001) #won't work with NAs


```


#Try stns on the peace, slave for timing of freshet 

```{r}
stns_ice <- tidyhydat::hy_daily_flows(station_number = c("07HA001", "07EF001", "07FD002", "07NB001", "07KC001"))

stns_ice <- stns_ice %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1920/10/1"), as.Date("2022/09/30"), by="day"))

#add water year
stns_ice <- addWaterYear(stns_ice) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year
stn_ice_split <- split(stns_ice, stns_ice$STATION_NUMBER )
stn_ice_split_doy <- lapply(stn_ice_split, calc_day_of_wyear)
stn_ice <- bind_rows(stn_ice_split_doy, .id = "STATION_NUMBER")

rm(stn_ice_split, stn_ice_split_doy)
```


```{r}
stn_cln_ice <- stn_ice %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1961) %>%
  filter(waterYear < 2021)

#print out rows with > 14 Nas in a row
lst_stns_cln_ice <- split(stn_cln_ice, stn_cln_ice$STATION_NUMBER )
lst_stns_ice_rle <- lapply(lst_stns_cln_ice, calc_rle)

for (i in lst_stns_ice_rle) {
  for (j in i)
    stn <- unique(i$STATION_NUMBER)
    wy <- unique(i$waterYear)
    print("delete the following station # and water year:")
    print(stn)
    print(wy)
}


stn_cln_ice <- stn_ice %>%
  filter(waterYear > 1961) %>%
  filter(waterYear < 2021) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2019)) %>%
  filter(!c(STATION_NUMBER == "07EF001" & waterYear == 2020)) %>%
  filter(!c(STATION_NUMBER == "07FD002" & waterYear == 2011)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2008)) %>%   
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2009)) %>%
  filter(!c(STATION_NUMBER == "07KC001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2013)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2014)) %>%
  filter(!c(STATION_NUMBER == "07NB001" & waterYear == 2020)) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)
```


```{r}
#make a list of dfs- split by station number
lst_stns <- split(stn_cln_ice, stn_cln_ice$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_g1 <- lapply(lst_stns, Group_1_ice_cover)
stn_g3 <- lapply(lst_stns, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_g1, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_g3, .id = "STATION_NUMBER")

#find max and min years for each result df to make sure have a complete set of years
minyr01 <- min(output1$waterYear)
maxyr01 <- max(output1$waterYear)

minyr03 <- min(output3$waterYear)
maxyr03 <- max(output3$waterYear)

#expand data set to include all years for all output dfs
output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyr03:maxyr03) #fills NA for values for missing years
```

#MK timing of freshet 
```{r}
stn_07KC001 <- output3_ex %>%
  filter(STATION_NUMBER == "07KC001") %>%
  filter(Day_of_year != 154) %>% #2011
    filter(Day_of_year != 155) %>% #2015
      filter(Day_of_year != 256) %>% #2012
        filter(Day_of_year != 240) %>% #1978
        na.omit()

grubbs.test(stn_07KC001$Day_of_year)

#"07HA001", "", "07FD002", "07NB001", "07KC001"))

TS_07KC001 <- ts(stn_07KC001$Day_of_year, frequency = 1, start = c(1962,1))

plot(TS_07KC001)
MannKendall(TS_07KC001) #p = .007
sens.slope(TS_07KC001) #.38, .45: pos slope means increasing?
pettitt.test(TS_07KC001) #28-1998, with outliers removed

plot.ts(TS_07KC001)

TS_07KC001_diff <- diff(TS_07KC001, differences = 1)
plot.ts(TS_07KC001_diff)

#freshet happening later
```


```{r}

stn_07HA001 <- output3_ex %>%
  filter(STATION_NUMBER == "07HA001") %>%
  na.omit()

TS_07HA001 <- ts(stn_07HA001$Day_of_year, 
        frequency = 1,
         start = c(1962,1))

plot.ts(TS_07HA001) 
MK <- MannKendall(TS_07HA001) #p = .05 without the NAs???
sens.slope(TS_07HA001) #.32... increasing!
pettitt.test(TS_07HA001) 


```


```{r}

stn_07KC001 <- output3_ex %>%
  filter(STATION_NUMBER == "07HA001") 


```

#MK ice coverage
```{r}

stn_07HA001 <- output1_ex %>%
  filter(STATION_NUMBER == "07HA001")

TS_07HA001 <- ts(stn_07HA001$Ice_coverage_wy, frequency = 1, start = c(1962, 1))

plot.ts(TS_07HA001)

MannKendall(TS_07HA001) # p = 0.0
sens.slope(TS_07HA001) #-1.15
pettitt.test(TS_07HA001) #chng pnt
stn_07HA001[25,] #1986

```

```{r}
#ice coverage
stn_07KC001 <- output1_ex %>%
  filter(STATION_NUMBER == "07KC001") %>%
  filter(waterYear != 2006) %>%
  filter(waterYear != 1968) %>%
  na.omit()

grubbs.test(stn_07KC001$Ice_coverage_wy)

TS_07KC001 <- ts(stn_07KC001$Ice_coverage_wy, frequency = 1, start = c(1962,1))

MannKendall(TS_07KC001) #p = .054
sens.slope(TS_07KC001)  #-.2
pettitt.test(TS_07KC001) #chng pnt 35, 1997

```


