
Bayesian Belief Network: Hydrological Datasets

Flow and water level data as well as IHA variables and ice cycle variables. 

Run "Eflows_FUNCTIONS.R" file before running this script

```{r}

library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(dataRetrieval)
library(knitr)

```

##FLOW DATA

#1 extract flow data and add column for water year and day of the year

```{r}
#get stns to analyze
stn_all <- tidyhydat::hy_daily_flows(station_number = c('07KC001','07NB001', '07KE001','07DA001')) 

stn_all <- stn_all %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1960/10/1"), as.Date("2022/09/30"), by="day"))

#add water year
stn_all <- addWaterYear(stn_all) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year from "Eflows_FUNCTIONS.R" function 
stn_all_split <- split(stn_all, stn_all$STATION_NUMBER )
stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear)
stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER")

rm(stn_all_split, stn_all_split_doy)

```

#2 export csvs for flow data
Note: these csvs are untidied flow data. No estimates, no deleting years with >14 days of missing data, etc. See below for tidied data

```{r}

df1 <- stn_all %>% filter(STATION_NUMBER == "07KC001")
df2 <- stn_all %>% filter(STATION_NUMBER == "07NB001 ")
df3 <- stn_all %>% filter(STATION_NUMBER == "07KE001 ")
df4 <- stn_all %>% filter(STATION_NUMBER == "07DA001 ")

write.csv(df1, "PeaceRiver_07KC001_Flow.csv")
write.csv(df2, "SlaveRiver_07NB001_Flow.csv")
write.csv(df3, "BirchRiver_07KE001_Flow.csv")
write.csv(df4, "AthabascaRiver_07DA001_Flow.csv")

```

#3 Make three separate datasets in prep for IHA calculations: 
-full year
-open water data (April-October) 
-winter/closed water (November-March)

```{r}
stn_all <- stn_all # full dataset

#open water data (change numbers in filter argument to select diff months)
stn_openwater <- stn_all %>% 
filter(month(Date) %in% c(4,5,6,7,8,9,10))  

#closed water data (change numbers in filter argument to select diff months)
stn_closed <- stn_all %>%
  filter(month(Date) %in% c(11,12,1,2,3))

```

#4 Prep for IHA calculations: Get rid of years that have more than 14 consecutive days of missing data, and fill in missing values
-full dataset
```{r}

#data prep for full dataset
stn_cln <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1960) %>%
  filter(waterYear != 2022)

#delete years that have >14 days missing data using calc_rle function from functions for this project
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist the station numbers
stns_ready_full <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_dmy<- stns_ready_full %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

-open water data

```{r}

#data prep for open water data
stn_cln <- stn_openwater %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1960) %>%
  filter(waterYear != 2022)

#delete years that have >14 days missing data using calc_rle function from functions for this project
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist the station numbers
stns_ready_openwater <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_dmy_openwater <- stns_ready_openwater %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

-Winter/closed water data

```{r}

#data prep for winter/closed water
stn_cln <- stn_openwater %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1960) %>%
  filter(waterYear != 2022)

#delete years that have >14 days missing data using calc_rle function from functions for this project
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist the station numbers
stns_ready_closed <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_dmy_closedwater <- stns_ready_closed %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

#5 IHA calculations
-full year
-open water
-closed water

```{r}

#IHA calculations: -full year
#make a list of dfs- split by station number
lst_stns <- split(stns_dmy, stns_dmy$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_fullyr <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

###########################

#IHA calculations: -open water
#make a list of dfs- split by station number
lst_stns <- split(stns_dmy_openwater, stns_dmy_openwater$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_openwater <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

###########################

#IHA calculations: -closed water
#make a list of dfs- split by station number
lst_stns <- split(stns_dmy_closedwater, stns_dmy_closedwater$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_closedwater <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

```

#6 Export IHA variables to csv with a separate csv by station number

```{r}
#select variables for each dataset

IHA_fullyr_s <- IHA_fullyr %>% 
  select("STATION_NUMBER", "Year", "1 Day Max", "30 Day Min", "30 Day Max", "Low pulse length", "High pulse length", "Reversals")

IHA_openwater_s <- IHA_openwater  %>% 
  select("STATION_NUMBER", "Year", "1 Day Max", "30 Day Min", "30 Day Max", "Low pulse length", "High pulse length", "Reversals")

IHA_closedwater_s <- IHA_closedwater  %>%
select("STATION_NUMBER", "Year", "1 Day Max", "30 Day Min", "30 Day Max", "Low pulse length", "High pulse length", "Reversals")


#write a csv for each station number- full year, IHA
for (i in unique(IHA_fullyr_s$STATION_NUMBER)) {
  dfsubset <- IHA_fullyr_s[IHA_fullyr_s$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_fullyr_flow_IHA.csv"))
}

#write a csv for each station number- open water, IHA
for (i in unique(IHA_openwater_s$STATION_NUMBER)) {
  dfsubset <- IHA_openwater_s[IHA_openwater_s$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_openwater_flow_IHA.csv"))
}

#write a csv for each station number- closed water, IHA
for (i in unique(IHA_closedwater_s$STATION_NUMBER)) {
  dfsubset <- IHA_closedwater_s[IHA_closedwater_s$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_closed_flow_IHA.csv"))
}

```


#7 Calculate Ice variables
-full dataset 
```{r}

#data prep 
#change the date format, and fill any values that are missing with preceding values 
stns_yearmonthday <- stns_ready_full %>%
  mutate(Date = as.Date(Date)) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup')

# make a list of dfs- split by station number
lst_stns <- split(stns_yearmonthday, stns_yearmonthday$STATION_NUMBER )

#apply ice variable function to all dfs in the list, by station 
stn_g1 <- lapply(lst_stns, Group_1_ice_cover)
stn_g2 <- lapply(lst_stns, Group_2_freeze_thaw)
stn_g3 <- lapply(lst_stns, Group_3_freshet)

#Unnlist the outputs
output1 <- bind_rows(stn_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_g3, .id = "STATION_NUMBER")

#Deal with each output df having different lengths and not being able to combine them
#find max and min years for each result df
minyro1 <- min(output1$waterYear)
maxyro1 <- max(output1$waterYear)

minyro2 <- min(output2$waterYear)
maxyro2 <- max(output2$waterYear)

minyro3 <- min(output3$waterYear)
maxyro3 <- max(output3$waterYear)

#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro1:maxyro1) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro2:maxyro2) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro3:maxyro3) #fills NA for values for missing years 

df_ICE_full <- cbind(output1_ex, output2_ex, output3_ex)

df_ICE_full_final <- df_ICE_full[!duplicated(as.list(df_ICE_full))] #removed duplicated columns (delete this if columns other than station number are the same)


```

-open water
```{r}

#data prep 
#change the date format, and fill any values that are missing with preceding values 
stns_yearmonthday <- stns_ready_openwater %>%
  mutate(Date = as.Date(Date)) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup')

# make a list of dfs- split by station number
lst_stns <- split(stns_yearmonthday, stns_yearmonthday$STATION_NUMBER )

#apply ice variable function to all dfs in the list, by station 
stn_g1 <- lapply(lst_stns, Group_1_ice_cover)
stn_g2 <- lapply(lst_stns, Group_2_freeze_thaw)
stn_g3 <- lapply(lst_stns, Group_3_freshet)

#Unnlist the outputs
output1 <- bind_rows(stn_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_g3, .id = "STATION_NUMBER")

#Deal with each output df having different lengths and not being able to combine them
#find max and min years for each result df
minyro1 <- min(output1$waterYear)
maxyro1 <- max(output1$waterYear)

minyro2 <- min(output2$waterYear)
maxyro2 <- max(output2$waterYear)

minyro3 <- min(output3$waterYear)
maxyro3 <- max(output3$waterYear)

#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro1:maxyro1) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro2:maxyro2) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro3:maxyro3) #fills NA for values for missing years 

df_ICE_opwater <- cbind(output1_ex, output2_ex, output3_ex)

df_ICE_opwater_final <- df_ICE_opwater[!duplicated(as.list(df_ICE_opwater))] #removed duplicated columns (delete this if columns other than station number are the same)


```

-closed water
```{r}

#data prep 
#change the date format, and fill any values that are missing with preceding values 
stns_yearmonthday <- stns_ready_closed %>%
  mutate(Date = as.Date(Date)) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup')

# make a list of dfs- split by station number
lst_stns <- split(stns_yearmonthday, stns_yearmonthday$STATION_NUMBER )

#apply ice variable function to all dfs in the list, by station 
stn_g1 <- lapply(lst_stns, Group_1_ice_cover)
stn_g2 <- lapply(lst_stns, Group_2_freeze_thaw)
stn_g3 <- lapply(lst_stns, Group_3_freshet)

#Unnlist the outputs
output1 <- bind_rows(stn_g1, .id = "STATION_NUMBER") 
output2 <- bind_rows(stn_g2, .id = "STATION_NUMBER")
output3 <- bind_rows(stn_g3, .id = "STATION_NUMBER")

#Deal with each output df having different lengths and not being able to combine them
#find max and min years for each result df
minyro1 <- min(output1$waterYear)
maxyro1 <- max(output1$waterYear)

minyro2 <- min(output2$waterYear)
maxyro2 <- max(output2$waterYear)

minyro3 <- min(output3$waterYear)
maxyro3 <- max(output3$waterYear)

#expand data set to include all years for all output dfs
output1_ex <- output1 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro1:maxyro1) #fills NA for values for missing years 

output2_ex <- output2 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro2:maxyro2) #fills NA for values for missing years 

output3_ex <- output3 %>%
mutate(waterYear = as.integer(waterYear)) %>%
complete(STATION_NUMBER, waterYear = minyro3:maxyro3) #fills NA for values for missing years 

df_ICE_clwater <- cbind(output1_ex, output2_ex, output3_ex)

df_ICE_clwater_final <- df_ICE_clwater[!duplicated(as.list(df_ICE_clwater))] #removed duplicated columns (delete this if columns other than station number are the same)

```

#8 Export Ice variables to csv

```{r}

#write a csv for each station number- full year, Ice variables
for (i in unique(df_ICE_full_final$STATION_NUMBER)) {
  dfsubset <- df_ICE_full_final[df_ICE_full_final$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_fullyr_Icevar.csv"))
}

#write a csv for each station number- open water, Ice variables
for (i in unique(df_ICE_opwater_final$STATION_NUMBER)) {
  dfsubset <- df_ICE_opwater_final[df_ICE_opwater_final$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_openwater_Icevar.csv"))
}

#write a csv for each station number- closed water, Ice variables
for (i in unique(df_ICE_clwater_final$STATION_NUMBER)) {
  dfsubset <- df_ICE_clwater_final[df_ICE_clwater_final$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_closedwater_Icevar.csv"))
}


```

##WATER LEVEL DATA

#1 extract water level for stations and add column for water year and day of the year
```{r}

stns_wl <- tidyhydat::hy_daily_levels(station_number = c("07KF002", "07KF003", "07MD001"))

stn_all <- stns_wl %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
complete(Date = seq.Date(as.Date("1960/10/1"), as.Date("2022/09/30"), by="day"))

#add water year
stn_all <- addWaterYear(stn_all) %>%
mutate(waterYear = as.character(waterYear))

#add day of the year from "Eflows_FUNCTIONS.R" function 
stn_all_split <- split(stn_all, stn_all$STATION_NUMBER )
stn_all_split_doy <- lapply(stn_all_split, calc_day_of_wyear)
stn_all <- bind_rows(stn_all_split_doy, .id = "STATION_NUMBER")

rm(stn_all_split, stn_all_split_doy)
```

#2 export water level data to csvs
Note: these csvs are untidied flow data. No estimates, no deleting years with >14 days of missing data, etc. See below for tidied data

```{r}

df1 <- stn_all %>% filter(STATION_NUMBER == "07KF002")
df2 <- stn_all %>% filter(STATION_NUMBER == "07KF003")
df3 <- stn_all %>% filter(STATION_NUMBER == "07MD001")

write.csv(df1, "LakeClaire_07KF002_WL.csv")
write.csv(df2, "MamawiLake_07KF003_WL.csv")
write.csv(df3, "LakeAthabasca_07MD001_WL.csv")

```

#3 Make three separate datasets in prep for IHA calculations: 
-full year
-open water data (April-October) 
-winter/closed water (November-March)
```{r}

stn_all <- stn_all # full dataset

#open water data (change numbers in filter argument to select diff months)
stn_openwater <- stn_all %>% 
filter(month(Date) %in% c(4,5,6,7,8,9,10))  

#closed water data (change numbers in filter argument to select diff months)
stn_closed <- stn_all %>%
  filter(month(Date) %in% c(11,12,1,2,3))

```

#4 Prep for IHA calculations: Get rid of years that have more than 14 consecutive days of missing data, and fill in missing values
-full dataset

```{r}

#data prep for full dataset
stn_cln <- stn_all %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1960) %>%
  filter(waterYear != 2022)

#delete years that have >14 days missing data using calc_rle function from functions for this project
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist the station numbers
stns_ready_full_wl <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_dmy_wl <- stns_ready_full_wl %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

-open water

```{r}

#data prep for full dataset
stn_cln <- stn_openwater %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1960) %>%
  filter(waterYear != 2022)

#delete years that have >14 days missing data using calc_rle function from functions for this project
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist the station numbers
stns_ready_opwater_wl <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_opwater_wl <- stns_ready_opwater_wl %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)
```

-closed water

```{r}

#data prep for full dataset
stn_cln <- stn_closed %>%
  mutate(Date = as.Date(Date)) %>%
  filter(waterYear > 1960) %>%
  filter(waterYear != 2022)

#delete years that have >14 days missing data using calc_rle function from functions for this project
lst_stns_cln <- split(stn_cln, stn_cln$STATION_NUMBER )
lst_stns_cln_rle <- lapply(lst_stns_cln, calc_rle)

#unlist the station numbers
stns_ready_cl_wl <- bind_rows(lst_stns_cln_rle, .id = "STATION_NUMBER") 

#change date format and fill any values that are missing with preceding values 
stns_dmy_cl_wl <- stns_ready_cl_wl %>%
  mutate(Date = format(Date,"%d-%m-%Y")) %>%
  group_by(STATION_NUMBER) %>%
  fill(Value, .direction = 'downup') #fill missing values going down then up (ie at start of df)

```

#5 IHA calculations
-full year
-open water
-closed water

```{r}

#IHA calculations: -full year
#make a list of dfs- split by station number
lst_stns <- split(stns_dmy_wl, stns_dmy_wl$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_fullyr_wl <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

###########################

#IHA calculations: -open water
#make a list of dfs- split by station number
lst_stns <- split(stns_opwater_wl, stns_opwater_wl$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_openwater_wl <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

###########################

#IHA calculations: -closed water
#make a list of dfs- split by station number
lst_stns <- split(stns_dmy_cl_wl, stns_dmy_cl_wl$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_IHA <- lapply(lst_stns, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_closedwater_wl <- bind_rows(lst_stns_IHA, .id = "STATION_NUMBER")

```


#5 Export as csv for each station
```{r}

#select variables for each dataset

IHA_fullyrwl_S <- IHA_fullyr_wl %>% 
  select("STATION_NUMBER", "Year", "1 Day Max", "30 Day Min", "30 Day Max", "Low pulse length", "High pulse length", "Reversals")

IHA_openwater_wl_s <- IHA_openwater_wl  %>% 
  select("STATION_NUMBER", "Year", "1 Day Max", "30 Day Min", "30 Day Max", "Low pulse length", "High pulse length", "Reversals")

IHA_closedwater_wl_s <- IHA_closedwater_wl  %>%
select("STATION_NUMBER", "Year", "1 Day Max", "30 Day Min", "30 Day Max", "Low pulse length", "High pulse length", "Reversals")


#write a csv for each station number- full year, IHA
for (i in unique(IHA_fullyrwl_S$STATION_NUMBER)) {
  dfsubset <- IHA_fullyrwl_S[IHA_fullyrwl_S$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_fullyr_wl_IHA.csv"))
}

#write a csv for each station number- open water, IHA
for (i in unique(IHA_openwater_wl_s$STATION_NUMBER)) {
  dfsubset <- IHA_openwater_wl_s[IHA_openwater_wl_s$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_openwater_wl_IHA.csv"))
}

#write a csv for each station number- closed water, IHA
for (i in unique(IHA_closedwater_wl_s$STATION_NUMBER)) {
  dfsubset <- IHA_closedwater_wl_s[IHA_closedwater_wl_s$STATION_NUMBER == i,]
  write.csv(dfsubset, paste0(i, "_closed_wl_IHA.csv"))
}

```


