# Change in Water Flow on the Peace-Athabasca Delta 

Median percent change in water flow for ~30 stations on the PAD, for two periods, 1975 - 2005, and 2006-2020. 

These periods were chosen to assess changes in river flows that may be related to the impacts of climate change, while also controlling for the effects of the Bennett dam, installed and running from 1972.

```{r, echo=FALSE, include = FALSE}
library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(grid)
library(gridExtra)
library(viridis)
library(sqldf)
#install.packages("IHA", repos="http://R-Forge.R-project.org", dependencies = TRUE)

```


```{r, echo = FALSE}
#IHA FUNCTION
calc_IHA <- function(data){
  
flow_data <-  {{data}} #%>%
#  select(Date, Value)
  
flow_data <- zoo(flow_data$Value, order.by = as.Date(as.character(flow_data$Date), format = "%d-%m-%Y"))
## Run IHA analyses
group1_output <- group1(flow_data, year = "calendar", FUN = median)
group2_output <- group2(flow_data, year = "calendar", mimic.tnc = TRUE)
group3_output <- group3(flow_data, year = "calendar", mimic.tnc = FALSE)
#group4_output <- group4(flow_data, year = "calendar")
group5_output <- group5(flow_data, year = "calendar")

## Convert outputs
group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
#group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)

## Create output dataframe 
#Need to figure out how to join all 4 dfs even if they are diff sizes - use reduce? 
#group4 data left out for now (low/high pulse numbers)

IHA_output <- bind_cols(list(group1_output, group2_output, group3_output, group5_output)) #group4_output

#make the years the column names instead of the rownames
IHA_output <- tibble::rownames_to_column(IHA_output, "Year") 

}

```

```{r, echo = FALSE, include = FALSE}
#Station metadata with lat/longs:

stn_meta <- tidyhydat::hy_stations(station_number = c('07AD001', '07AF002', '07AH001', '07BC002', '07BE001', '07BK007', '07CD001', '07CA006', '07CD004', '07CD005', '07DA001', '07DB001', '07DD002', '07EF001', '07EE007', '07EC003', '07FD001', '07FB001', '07FD012', '07GJ001', '07GH002', '07GE001', '07HA001', '07HC001', '07JD002', '07KC001', '07LE002', '07LB002', '07MB001', '07NB001', '07MA003'  ))

stn_meta <- stn_meta %>%
  select(STATION_NUMBER, STATION_NAME, PROV_TERR_STATE_LOC, LATITUDE, LONGITUDE, HYD_STATUS) 
#write.csv(stn_meta, "Stns_Climate_Change_LAT_LONG.csv")
rm(stn_meta)
```

Get all the flow data for the stations that were selected using the tile plots:
```{r}

stns_all <- tidyhydat::hy_daily_flows(station_number = c('07AD002', '07AF002', '07AH001', '07BC002', '07BE001', '07BK007', '07CD001', '07CA006', '07CD004', '07CD005', '07DA001', '07DB001', '07DD002', '07EF001', '07EE007', '07EC003', '07FD001', '07FB001', '07FD012', '07GJ001', '07GH002', '07GE001', '07HA001', '07HC001', '07JD002', '07KC001', '07LE002', '07LB002', '07MB001', '07NB001', '07MA003'))

```

## Pre period (1975 - 2005 ) data tidying

```{r}
#Keep only the 1 March - 31 October data from 1975 - 2005

stns_pre <- stns_all %>%
mutate(Date = as.Date(Date)) %>%
  filter(month(Date) %in% c(3,4,5,6,7,8,9,10)) %>%
  filter(year(Date) %in% c(1975, 1976, 1977, 1978, 1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005)) 
```


```{r}
#delete years with partial data 

#Returns df with number of days of data in each year.
count_pre <- stns_pre %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
  complete(Date = seq.Date(as.Date("1975/3/1"), as.Date("2005/10/31"), by="day")) %>%
  #make a separate column for year
  mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y")) %>%
  group_by(STATION_NUMBER, Year) %>%
  count(is.na(Value)) %>%
  rename("Is_NA" = "is.na(Value)")  %>%
  pivot_wider(names_from = Is_NA, values_from = n) %>%
  rename('Number_values' = "FALSE",
         "Number_NAs" = "TRUE" )  %>%
  select(-c(Number_NAs)) %>%
  #na.omit(Number_values) %>%
 mutate("Watershed" = str_extract(STATION_NUMBER, ".{2}(.)") )%>%
mutate_if(is.numeric, ~replace(., is.na(.), 0))

#select all stations with less than 235 days a year (max is 245 a year)
pre_to_del <-  count_pre %>%
filter(Number_values <= 235 )

#get list of stns #s and years to delete, subtract these stations from stns_pre df.
pre_to_del <- unique(pre_to_del$STATION_NUMBER)

```

Remove the years with partial data from the pre-station data frame in prep for IHA calculations (from "pre_to_del")

```{r}
#remove years with partial data from pre df before calculations
stns_pre <- stns_pre %>%
mutate(Year = year(Date))  %>%
    filter(!c(STATION_NUMBER == "07CD004" & Year == "1978"), 
       !c(STATION_NUMBER == "07CD004" & Year == "1992"),
       !c(STATION_NUMBER == "07CD004" & Year == "1993"),
       !c(STATION_NUMBER == "07CD005" & Year == "1975"),
       !c(STATION_NUMBER == "07DB001" & Year == "1991"),
       !c(STATION_NUMBER == "07DB001" & Year == "1992"),
       !c(STATION_NUMBER == "07DD002" & Year == "1991"),
       !c(STATION_NUMBER == "07DD002" & Year == "1992"),
       !c(STATION_NUMBER == "07EE007" & Year == "2005"),
       !c(STATION_NUMBER == "07FD012" & Year == "1975"),
       !c(STATION_NUMBER == "07LB002" & Year == "1977"),
       !c(STATION_NUMBER == "07LE002" & Year == "1995"),
       !c(STATION_NUMBER == "07LE002" & Year == "1998"),
       !c(STATION_NUMBER == "07LE002" & Year == "2000"),
       !c(STATION_NUMBER == "07MA003" & Year == "1975"),
       !c(STATION_NUMBER == "07MA003" & Year == "1989")  
       )%>%
      mutate(Date = format(Date,"%d-%m-%Y")) 
  

```

## IHA variables: Pre period (1975 - 2005)

```{r, warning= FALSE}
#make a list of dfs- split by station number
lst_stns_pre <- split(stns_pre, stns_pre$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_pre_IHA <- lapply(lst_stns_pre, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
IHA_pre <- bind_rows(lst_stns_pre_IHA, .id = "Station_Number") %>%
  filter(Year != "2006")

```


```{r}

#expand so that all years from 1975-2005 are present for all stations
IHA_pre <- IHA_pre %>%
  mutate(Year = as.integer(Year)) %>%
    complete(Station_Number, Year = 1975:2005) %>%
    mutate("Watershed" = substring(Station_Number, 3,3))


```

## Post period (2006 - 2020) data tidying

```{r}

#make df for 1 March - 31 October data from 2006-2020
stns_pst <- stns_all %>%
mutate(Date = as.Date(Date)) %>%
  filter(month(Date) %in% c(3,4,5,6,7,8,9,10)) %>%
  filter(year(Date) %in% c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)) 
#make a separate column for year - this was done after the previous code was written, and was done in order to get rid of the Stn-year combos proceeding, ie don't need the code below if keeping all years and stations

```


```{r}
#delete years with partial data 

#Returns df with number of days of data in each year.
post_count <- stns_pst %>%
group_by(STATION_NUMBER) %>%
#make a complete set of days for each year 
  complete(Date = seq.Date(as.Date("2006/3/1"), as.Date("2020/10/31"), by="day")) %>%
  #make a separate column for year
  mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y")) %>%
  group_by(STATION_NUMBER, Year) %>%
  count(is.na(Value)) %>%
  rename("Is_NA" = "is.na(Value)")  %>%
  pivot_wider(names_from = Is_NA, values_from = n) %>%
  rename('Number_values' = "FALSE",
         "Number_NAs" = "TRUE" )  %>%
  select(-c(Number_NAs)) %>%
  #na.omit(Number_values) %>%
 mutate("Watershed" = str_extract(STATION_NUMBER, ".{2}(.)") )%>%
mutate_if(is.numeric, ~replace(., is.na(.), 0))

```

Get rid of all stations that have stns with less than 235 days of data - post data 

```{r}
#select all stations with less than 235 days a year (max is 245 a year)

post_to_del <-  post_count %>%
filter(Number_values <= 235 )

#get a list of stns and years to delete
post_to_del <- unique(post_to_del$STATION_NUMBER)


```

Remove the years with partial data from the post-station data frame in prep for IHA calculations (from "post_to_del")

```{r}

stns_pst <- stns_pst %>%
mutate(Year = year(Date))  %>%
#years to get rid of for each station number:
filter(!c(STATION_NUMBER == "07CA006" & Year == "2013"), 
       !c(STATION_NUMBER == "07CA006" & Year == "2014"),
       !c(STATION_NUMBER == "07EE007" & Year == "2020"),
       !c(STATION_NUMBER == "07EF001" & Year == "2019"),
       !c(STATION_NUMBER == "07FD012" & Year == "2011"),
       !c(STATION_NUMBER == "07KC001" & Year == "2009"),
       !c(STATION_NUMBER == "07MB001" & Year == "2020"),
       !c(STATION_NUMBER == "07NB001" & Year == "2013"),
       !c(STATION_NUMBER == "07NB001" & Year == "2020") ) %>%
      mutate(Date = format(Date,"%d-%m-%Y")) 

```


## IHA variables: Post period (2006 - 2020 )

```{r, warning=FALSE}

#make a list of dfs by station number
stns_pst_lst <- split(stns_pst, stns_pst$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_pst_IHA <- lapply(stns_pst_lst, calc_IHA)

IHA_pst <- bind_rows(lst_stns_pst_IHA, .id = "Station_Number") %>%
  filter(Year != "2021")

```

##################################
## Calculate Percent Change 

```{r}
#expand so that all years from 2006-2020 are present for all stations
IHA_pst <- IHA_pst %>%
  mutate(Year = as.integer(Year)) %>%
    complete(Station_Number, Year = 2006:2020) 
```

Make a df for the medians for the 1975-2005 period 

```{r}

#find median of all IHA variables for 1975-2005
IHA_medians_pre <- IHA_pre %>%
  group_by(Station_Number) %>%
  summarise(across(January:Reversals, median )) ###Change to october if doing water year 

```


Function to calculate the percent change for each station
```{r}

calc_perc_change <- function(data_pre, data_pst, stn){

years_post <- {{data_pst}}$Year

IHA_medians_pre <- {{data_pre}} %>%
  filter(Station_Number == {{stn}}) %>% 
  select(-c(Station_Number))

IHA_pst <- {{data_pst}} %>%
  filter(Station_Number == {{stn}})%>% 
  select(-c(Station_Number, Year))

IHA_pre_expand_rows <- IHA_medians_pre[rep(1, nrow(IHA_pst)),]

output <- ((IHA_pst - IHA_pre_expand_rows) /IHA_pre_expand_rows ) * 100

percent_change <- merge(years_post, output, by.x = 0, by.y = 0) %>%
  rename("Year" = "x")
}

```

Calculate percent change
```{r}
df_lst <- list()

for (i in unique(IHA_medians_pre$Station_Number)) {

calc <- calc_perc_change(IHA_medians_pre, IHA_pst, stn = i) #calc perc. change for all stns
calc$i <- i #add i (stn # to the calc df)
df_lst[[i]] <- calc #add calc df to the list
}

```


```{r}
#expand list of dfs to one df, add watershed and rename columns 

final_pc <- bind_rows(df_lst) %>%
 rename("Station_Number" = "i") %>%
  mutate("Watershed" = substring(Station_Number, 3,3))%>%
    rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min")

```


```{r, include = FALSE, echo = FALSE}

IHA_pst <- IHA_pst %>%
  mutate("Watershed" = substring(Station_Number, 3,3)) %>%
    rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min")


IHA_pre <- IHA_pre  %>%
   mutate("Watershed" = substring(Station_Number, 3,3)) %>%
    rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min")

```


```{r, echo = FALSE, include = FALSE}
 #graphing code 

#plot of all stations, coloured by watershed, on one graph
plt_all <-  function(df, xvar, title ){
ggplot(data = {{df}}, aes(y= {{xvar}}, x = factor(Year), group = Station_Number, colour = Watershed)) +
  geom_line(linewidth = 0.75) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  scale_x_discrete(breaks = seq(1975, 2006, 2)) +
    theme_bw() +
ggtitle({{title}}) +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE)
}


#ggplot(df, aes(x = factor(Year), y= Reversals, group = Station_Number, colour = Station_Number)) +
 # geom_line() +
  #scale_colour_viridis(discrete = TRUE)


```


```{r}

#plot of all stations, coloured and faceted by watershed

plt_facet <- function(df, yvar, title){
ggplot(data = {{df}}, aes(y={{yvar}}, x = factor(Year), group = Station_Number, colour = Watershed )) +
  geom_line(linewidth = 0.75) +
  facet_wrap(~Watershed) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  #scale_x_discrete(breaks = seq(1975, 2006, 2)) +
    theme_bw() +
ggtitle({{title}}) +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE) +
  theme(legend.position = "none")  +
  theme(axis.text.x = element_blank())
}

```


# Graphs

## Graphs with just a few stations 

```{r}


final_pc %>%
  filter(Watershed == "B") %>%
ggplot( aes(y= Reversals, x = factor(Year), group = Station_Number, colour = Station_Number)) +
  geom_line(linewidth = 0.75) +
  geom_hline(yintercept = 0, linetype = "dotted") +
 # scale_x_discrete(breaks = seq(1975, 2006, 2)) +
  ylim(-50, 50) +
   theme_classic() +
ggtitle(" Percent Change for Reversals compared to 1975-2005  ") +
  xlab("Year") +
  ylab("Percent change, Reversals") +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE) #+
  #theme(legend.position = "None")



```
```{r}


IHA_pre %>%
  filter(Watershed == "B") %>%
ggplot( aes(y= Reversals, x = factor(Year), group = Station_Number, colour = Station_Number)) +
  geom_line(linewidth = 0.75) +
  geom_hline(yintercept = 0, linetype = "dotted") +
 scale_x_discrete(breaks = seq(1975, 2006, 2)) +
  ylim(0, 80) +
   theme_classic() +
ggtitle("  Reversals 1975 - 2005 ") +
  xlab("Year") +
  ylab("Reversals") +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE) #+
  #theme(legend.position = "None")

```
```{r}

IHA_pst %>%
  filter(Watershed == "B") %>%
ggplot( aes(y= Reversals, x = factor(Year), group = Station_Number, colour = Station_Number)) +
  geom_line(linewidth = 0.75) +
  geom_hline(yintercept = 0, linetype = "dotted") +
 # scale_x_discrete(breaks = seq(1975, 2006, 2)) +
  ylim(0, 80) +
   theme_classic() +
ggtitle("  Reversals 2006 - 2020 ") +
  xlab("Year") +
  ylab("Reversals") +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE) #+
  #theme(legend.position = "None")


```


## Reversals

```{r, echo = FALSE, fig.height= 6, fig.width= 8}

##Reversals, Pre: 1975-2005
df <- IHA_pre %>%
  filter(Year != "2006") 

plt_facet(df, Reversals, "Reversals, 1975-2005" )

```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}

##Reversals, Post: 2006-2020

df <- IHA_pst %>%
  filter(Year != "2021") 

plt_facet(df,Reversals, "Reversals, 2006-2020" )

```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##Reversals, Percent Change 

df <- final_pc %>%
    filter(Year != "2021") 

plt_facet(df, Reversals, "Percent Change from Median Reversal Values for 1975-2005")

```


## 1 Day Minimum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Min, Pre: 1975-2005
df <- IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, One_Day_Min, "One Day Minimum, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Min, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021")

plt_facet(df, One_Day_Min, "One Day Minimum, 2006-2020")


```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Min, Percent Change
df <- final_pc %>%
    filter(Year != "2021")

plt_facet(df, One_Day_Min, "Percent Change from Median One Day Minimum Values for 1975-2005")

```
 
## 1 Day Maximum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Max, Pre: 1975-2005
df <- IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, One_Day_Max, "1 Day Max, 1975-2005")

```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Max, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, One_Day_Max, "1 Day Max, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Max, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, One_Day_Max, "Percent Change from Median 1 Day Maximum Values for 1975-2005")

```

## 7 Day Minimum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Min, Pre: 1975-2005

df <- IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Seven_Day_Min, "7 Day Min, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Min, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021")

plt_facet(df, Seven_Day_Min, "7 Day Min, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Min, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, Seven_Day_Min, "Percent Change from Median 7 Day Minimum Values for 1975-2005")

```

## 7 Day Maximum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Max, Pre: 1975-2005
df <- IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Seven_Day_Max, "7 Day Max, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Max, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, Seven_Day_Max, "7 Day Max, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Max, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, Seven_Day_Max, "Percent Change from Median 7 Day Maximum Values for 1975-2005")

```

## 30 Day Minimum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Min, Pre: 1975-2005
df <- IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Thirty_Day_Min, "30 Day Min, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Min, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, Thirty_Day_Min, "30 Day Min, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Min, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, Thirty_Day_Min, "Percent Change from Median 30 Day Minimum Values for 1975-2005")

```

## 30 Day Maximum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Max, Pre: 1975-2005
df <- IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Thirty_Day_Max, "30 Day Max, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Max, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, Thirty_Day_Max, "30 Day Max, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Max, Percent Change
df <- final_pc %>%
    filter(Year != "2021")

plt_facet(df, Thirty_Day_Max, "Percent Change from Median 30 Day Maximum Values for 1975-2005")

```



```{r}

```



```{r}

	
	
```

```{r}

```


```{r}

```


```{r}

```


```{r}
```

```{r, include = FALSE, echo = FALSE}
#ignore this code, keep just in case though

#Plotting options:

#make a graphing function to plot one at a time

plt <- function(data, variable) {
ggplot() +
  geom_line(data = {{data}}, aes(x = Year, y = {{variable}}), group = 1, colour = "red", linewidth = 0.75)  + 
      geom_point() +
        scale_x_discrete(breaks = seq(1975, 2006, 2))
}

#make a list of all plot objects
#pltlist <- lapply(lst_stns_march_oct_IHA, plt) #only works if plt function has x value already in it, can't figure out how to use lapply and choosing a diff x value

```


```{r}
