# Change in Water Flow on the Peace-Athabasca Delta 

Median percent change in water flow for ~30 stations on the PAD, for two periods, 1975 - 2005, and 2006-2020. 

These periods were chosen to assess changes in river flows that may be related to the impacts of climate change, while also controlling for the effects of the Bennet dam, installed and running from 1972.

```{r, echo=FALSE, include = FALSE}
library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(grid)
library(gridExtra)
library(viridis)
```


```{r, echo = FALSE}
#IHA FUNCTION
calc_IHA <- function(data){
  
flow_data <-  {{data}} #%>%
#  select(Date, Value)
  

flow_data <- zoo(flow_data$Value, order.by = as.Date(as.character(flow_data$Date), format = "%d-%m-%Y"))
## Run IHA analyses
group1_output <- group1(flow_data, year = "water", FUN = median)
group2_output <- group2(flow_data, year = "water", mimic.tnc = TRUE)
group3_output <- group3(flow_data, year = "water", mimic.tnc = FALSE)
#group4_output <- group4(flow_data, year = "water")
group5_output <- group5(flow_data, year = "water")

## Convert outputs
group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
#group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)

## Create output dataframe 
#Need to figure out how to join all 4 dfs even if they are diff sizes - use reduce? 
#group4 data left out for now (low/high pulse numbers)

IHA_output <- bind_cols(list(group1_output, group2_output, group3_output, group5_output)) #group4_output

#make the years the column names instead of the rownames
IHA_output <- tibble::rownames_to_column(IHA_output, "Year") 

}

```

```{r, echo = FALSE, include = FALSE}
#Station metadata with lat/longs:

stn_meta <- tidyhydat::hy_stations(station_number = c('07AD001', '07AF002', '07AH001', '07BC002', '07BE001', '07BK007', '07CD001', '07CA006', '07CD004', '07CD005', '07DA001', '07DB001', '07DD002', '07EF001', '07EE007', '07EC003', '07FD001', '07FB001', '07FD012', '07GJ001', '07GH002', '07GE001', '07HA001', '07HC001', '07JD002', '07KC001', '07LE002', '07LB002', '07MB001', '07NB001', '07MA003'  ))

stn_meta <- stn_meta %>%
  select(STATION_NUMBER, STATION_NAME, PROV_TERR_STATE_LOC, LATITUDE, LONGITUDE, HYD_STATUS) 
#write.csv(stn_meta, "Stns_Climate_Change_LAT_LONG.csv")
rm(stn_meta)
```

Get all the flow data for the stations that were selected using the tile plots:
```{r}

stns_all <- tidyhydat::hy_daily_flows(station_number = c('07AD002', '07AF002', '07AH001', '07BC002', '07BE001', '07BK007', '07CD001', '07CA006', '07CD004', '07CD005', '07DA001', '07DB001', '07DD002', '07EF001', '07EE007', '07EC003', '07FD001', '07FB001', '07FD012', '07GJ001', '07GH002', '07GE001', '07HA001', '07HC001', '07JD002', '07KC001', '07LE002', '07LB002', '07MB001', '07NB001', '07MA003'))

```

Keep only the 1 March - 31 October data from 1975 - 2005
```{r}
stns_march_oct <- stns_all %>%
mutate(Date = as.Date(Date)) %>%
  filter(month(Date) %in% c(3,4,5,6,7,8,9,10)) %>%
  filter(year(Date) %in% c(1975, 1976, 1977, 1978, 1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005)) %>%
  mutate(Date = format(Date,"%d-%m-%Y"))

```

## Calculate IHA variables for 1975 - 2005 (pre-period)
```{r, warning= FALSE}
#make a list of dfs- split by station number
lst_stns_march_oct <- split(stns_march_oct, stns_march_oct$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_march_oct_IHA <- lapply(lst_stns_march_oct, calc_IHA)

#Combine all IHA outputs into one df (unnlist the dfs)
#make a "watershed" column for graphing later
df_combined_IHA_pre <- bind_rows(lst_stns_march_oct_IHA, .id = "Station_Number") 

```

##################################
## Calculate Percent Change 

Make a df for the medians for the 1975-2005 period 
```{r}

#find median of all IHA variables for 1975-2005
IHA_medians_pre <- df_combined_IHA_pre %>%
  group_by(Station_Number) %>%
  summarise(across(October:Reversals, median ))

```

Make a df of flow data for 2006-2020 ("post" period)
```{r}

#make df for 1 March - 31 October data from 2006-2020
stns_all_pst <- stns_all %>%
mutate(Date = as.Date(Date)) %>%
  filter(month(Date) %in% c(3,4,5,6,7,8,9,10)) %>%
  filter(year(Date) %in% c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)) %>%
  mutate(Date = format(Date,"%d-%m-%Y"))

```

## Calculate IHA variables for post period
```{r, warning=FALSE}

#make a list of dfs by station number
stns_all_pst_mar_oct <- split(stns_all_pst, stns_all_pst$STATION_NUMBER )

#apply calc_IHA function to all dfs in the list
lst_stns_mar_oct_IHA <- lapply(stns_all_pst_mar_oct, calc_IHA)

IHA_pst <- bind_rows(lst_stns_mar_oct_IHA, .id = "Station_Number") 
```

Function to calculate the percent change for each station
```{r}

calc_perc_change <- function(data_pre, data_pst, stn){ #, data_pst, stn

years_post <- {{data_pst}}$Year

IHA_medians_pre <- {{data_pre}} %>%
  filter(Station_Number == {{stn}}) %>% 
  select(-c(Station_Number))

IHA_pst <- {{data_pst}} %>%
  filter(Station_Number == {{stn}})%>% 
  select(-c(Station_Number, Year))

IHA_pre_expand_rows <- IHA_medians_pre[rep(1, nrow(IHA_pst)),]

output <- ((IHA_pst - IHA_pre_expand_rows) /IHA_pre_expand_rows ) * 100

percent_change <- merge(years_post, output, by.x = 0, by.y = 0) %>%
  rename("Year" = "x")
}

```


```{r}
df_lst <- list()

for (i in unique(IHA_medians_pre$Station_Number)) {

calc <- calc_perc_change(IHA_medians_pre, IHA_pst, stn = i) #calc perc. change for all stns
calc$i <- i #add i (stn # to the calc df)
df_lst[[i]] <- calc #add calc df to the list
}

final_pc <- bind_rows(df_lst) %>%
  rename("Station_Number" = "i") %>%
  mutate("Watershed" = substring(Station_Number, 3,3))%>%
    rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min")

```


```{r, include = FALSE, echo = FALSE}

IHA_pst <- IHA_pst %>%
  mutate("Watershed" = substring(Station_Number, 3,3)) %>%
    rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min")


df_combined_IHA_pre <- df_combined_IHA_pre  %>%
   mutate("Watershed" = substring(Station_Number, 3,3)) %>%
    rename("One_Day_Max" = "1 Day Max",
           "One_Day_Min" = '1 Day Min',
           "Three_Day_Max" = "3 Day Max",
           "Three_Day_Min" = "3 Day Min",
           "Seven_Day_Min" ="7 Day Min",
          "Seven_Day_Max" = "7 Day Max",
          "Thirty_Day_Max" = "30 Day Max",
          "Thirty_Day_Min" = "30 Day Min",
          "Ninety_Day_Max" = "90 Day Max",
          "Ninety_Day_Min" = "90 Day Min")

```


```{r, echo = FALSE, include = FALSE}
 #graphing code 

#plot of all stations, coloured by watershed, on one graph
plt_all <-  function(df, xvar, title ){
ggplot(data = {{df}}, aes(y={{xvar}}, x = Year, group = Station_Number, colour = Watershed)) +
  geom_line(linewidth = 0.75) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  scale_x_discrete(breaks = seq(1975, 2006, 2)) +
    theme_bw() +
ggtitle({{title}}) +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE)
}

```


```{r}

#plot of all stations, coloured and faceted by watershed

plt_facet <- function(df, yvar, title){
ggplot(data = {{df}}, aes(y={{yvar}}, x = Year, group = Station_Number, colour = Watershed)) +
  geom_line(linewidth = 0.75) +
  facet_wrap(~Watershed) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  #scale_x_discrete(breaks = seq(1975, 2006, 2)) +
    theme_bw() +
ggtitle({{title}}) +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_viridis(discrete = TRUE) +
  theme(legend.position = "none")  +
  theme(axis.text.x = element_blank())
}


```


# Graphs

## Reversals

```{r, echo = FALSE, fig.height= 6, fig.width= 8}

##Reversals, Pre: 1975-2005
df <- df_combined_IHA_pre %>%
  filter(Year != "2006") 

plt_facet(df, Reversals, "Reversals, 1975-2005" )


```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}

##Reversals, Post: 2006-2020

df <- IHA_pst %>%
  filter(Year != "2021") 

plt_facet(df,Reversals, "Reversals, 2006-2020" )


```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##Reversals, Percent Change 

df <- final_pc %>%
    filter(Year != "2021") 

plt_facet(df, Reversals, "Percent Change from Median Reversal Values for 1975-2005")

```


## 1 Day Minimum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Min, Pre: 1975-2005
df <- df_combined_IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, One_Day_Min, "One Day Minimum, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Min, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, One_Day_Min, "One Day Minimum, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Min, Percent Change
df <- final_pc %>%
    filter(Year != "2021")

plt_facet(df, One_Day_Min, "Percent Change from Median One Day Minimum Values for 1975-2005")

```
 
## 1 Day Maximum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Max, Pre: 1975-2005
df <- df_combined_IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, One_Day_Max, "1 Day Max, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Max, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, One_Day_Max, "1 Day Max, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##1 Day Max, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, One_Day_Max, "Percent Change from Median 1 Day Maximum Values for 1975-2005")

```

## 7 Day Minimum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Min, Pre: 1975-2005

df <- df_combined_IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Seven_Day_Min, "7 Day Min, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Min, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021")

plt_facet(df, Seven_Day_Min, "7 Day Min, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Min, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, Seven_Day_Min, "Percent Change from Median 7 Day Minimum Values for 1975-2005")

```

## 7 Day Maximum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Max, Pre: 1975-2005
df <- df_combined_IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Seven_Day_Max, "7 Day Max, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Max, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, Seven_Day_Max, "7 Day Max, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##7 Day Max, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, Seven_Day_Max, "Percent Change from Median 7 Day Maximum Values for 1975-2005")

```

## 30 Day Minimum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Min, Pre: 1975-2005
df <- df_combined_IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Thirty_Day_Min, "30 Day Min, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Min, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, Thirty_Day_Min, "30 Day Min, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Min, Percent Change
df <- final_pc  %>%
    filter(Year != "2021")

plt_facet(df, Thirty_Day_Min, "Percent Change from Median 30 Day Minimum Values for 1975-2005")

```

## 30 Day Maximum

```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Max, Pre: 1975-2005
df <- df_combined_IHA_pre %>%
  filter(Year != "2006")

plt_facet(df, Thirty_Day_Max, "30 Day Max, 1975-2005")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Max, 2006-2020
df <- IHA_pst %>%
  filter(Year != "2021") 
plt_facet(df, Thirty_Day_Max, "30 Day Max, 2006-2020")
```


```{r, echo = FALSE, fig.height= 6, fig.width= 8}
##30 Day Max, Percent Change
df <- final_pc %>%
    filter(Year != "2021")

plt_facet(df, Thirty_Day_Max, "Percent Change from Median 30 Day Maximum Values for 1975-2005")

```



```{r}




```


```{r}


```


```{r}


```


```{r, include = FALSE, echo = FALSE}
#ignore this code, keep just in case though

#Plotting options:

#make a graphing function to plot one at a time

#plt <- function(data, variable) {
#ggplot() +
#  geom_line(data = {{data}}, aes(x = Year, y = {{variable}}), group = 1, colour = #"red", linewidth = 0.75)  + 
 #     geom_point() +
#        scale_x_discrete(breaks = seq(1975, 2006, 2))
#}

#make a list of all plot objects
#pltlist <- lapply(lst_stns_march_oct_IHA, plt) #only works if plt function has x value already in it, can't figure out how to use lapply and choosing a diff x value

```


```{r}

