---
title: "Code to Save"
output: html_notebook
---

##Code snippets for use in the WBNP EF project##

```{r}

#compare two lists of station numbers to see if any are in common
#This is useful for finding the same pre and post dam stations

unique(postdam_stations[postdam_stations %in% predam_stations])

```

##Extract data and put into a list of df. This uses trycatch so it will continue to extract data even if a station number is null

```{r}

#Flow only with trycatch 
dfFlow_subset <- list()
for (st in stations_subset) {data <- tryCatch(tidyhydat::hy_daily_flows(station_number = (st)), error = function(e) NULL)
if (!is.null(data)){
dfFlow_subset[[st]] <- data.frame(data) 
}
}

```

## checking that the above worked

```{r}

for (df in dfFlow_subset) {
  print(unique(df$STATION_NUMBER, df$Parameter))
  print(unique(df$Parameter)) 
}

```

#keep this, it's useful if need to go back to using metadata to graph- ie just looking at year and not percentage of years
```{r}

#Create a new column with a year for every single year between the "year_from" and "year_to" columns. 

Stations_metadata_FLOW <- Stations_metadata_onelist %>%
  dplyr::select(-c(MEASUREMENT)) %>%
 dplyr::filter(DATA_TYPE == "Flow" ) %>%
tidyr::unite("Unique_Id", c("STATION_NUMBER", "OPERATION", "DATA_TYPE" ), sep = "_", remove = FALSE ) %>%
dplyr::mutate(Year_from = as.integer(Year_from), Year_to = as.integer(Year_to)) %>%
dplyr::mutate(year = map2(Year_from, Year_to, ~seq(.x, .y), .id = "year")) %>%
 tidyr::unnest(cols = c(year))

```

## GRAPHING FUNCTIONS AND MISC GRAPH CODE ##

Line graph of one IHA output variable at a time. Coloured by time period (pre/post dam).

```{r}

ggplot(data = IHA_both, aes(x =Year, y = Reversals, group = 1) ) +
          geom_line(aes(colour = Time_Period) ) +
      geom_point(aes(colour = Time_Period)) +
  ggtitle( "07NB001 \n Reversals") +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.position=c(0.87, 0.8), legend.title=element_text(size=8), legend.text=element_text(size=7)) +
  geom_line(data = IHA_predam_summary, y =  Reversals) +
   scale_colour_discrete(na.translate = F)
  

```

Line graphs of all IHA variables, where x= year and y = each variable in succession.
Coloured by time period (pre/post dam).

```{r}

dontgraphme <- c("Year", "Time_Period", "Zero flow days", '1 Day Min', '3 Day Min', '7 Day Min', '30 Day Min', '90 Day Min', '1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max','October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September')

list_plots <- vector('list', length(cols))

for (col in names(IHA_both)[!names(IHA_both) %in% dontgraphme]) {
  print(ggplot(data = IHA_both, aes(x = Year, y = IHA_both[,col], group = 1 )) +
          geom_line(aes(colour = Time_Period) ) +
      geom_point(aes(colour = Time_Period)) +
  ggtitle(col)  +
    ylab(col) +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.position=c(0.87, 0.8), legend.title=element_text(size=8), legend.text=element_text(size=7))+
       scale_colour_discrete(na.translate = F)) 
   print(list_plots[col])

}


```
Boxplots of IHA variables: 

Y = flow and x = the IHA variables passed in to the function.

Coloured by time period (pre/post dam).

Minimum flow
```{r}

boxplot_min <- function(df) {

p <-  IHA_both %>% 
  na.omit() %>%
  select('Year', '1 Day Min', '3 Day Min', '7 Day Min', '30 Day Min', '90 Day Min', 'Time_Period') %>%
  pivot_longer(., cols = c('1 Day Min', '3 Day Min', '7 Day Min','30 Day Min' , '90 Day Min'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
geom_boxplot() +
  geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
       theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
  scale_fill_manual(values = c("orange","light blue"), 
                    labels = c("Pre Dam", "Post Dam")) +
  xlab("") +
  ylab("Flow") 

return(p)

}

```

Boxplots of IHA variables: 

Maximum Flow
```{r}


boxplot_max <- function(df) {


p <-  IHA_both %>% 
  na.omit() %>%
  select('Year', '1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max','Time_Period') %>%
  pivot_longer(., cols = c('1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max', '90 Day Max'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
geom_boxplot() +
  geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
  scale_fill_manual(values = c("orange","light blue"), 
                    labels = c("Pre Dam", "Post Dam")) +
  xlab("") +
  ylab("Flow") 

return(p)

}


```

Boxplots of IHA variables: 

Flow by month
```{r}

boxplot_month <- function(df) {
  
p <-  IHA_both %>% 
  na.omit() %>%
  select('Year', 'October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September', 'Time_Period') %>%
  pivot_longer(., cols = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
geom_boxplot() +
  #geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
  scale_fill_manual(values = c("orange","light blue"), 
                    labels = c("Pre Dam", "Post Dam")) +
 aes(x = fct_inorder(Var)) +
  xlab("Month") +
  ylab("Flow") 

return(p)

}

```

```{r}
IHA_output_predam %>% 
  na.omit() %>%
  select('Year', 'October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September', 'Time_Period') %>%
  pivot_longer(., cols = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September'), names_to = "Var", values_to = "Val") %>%
ggplot(aes(x = Var, 
           y = Val, 
           group = Var,
           color = Var)) +
  stat_summary(fun.y = "mean", geom = "point") # +
 # stat_summary(fun.y = "mean", geom = "line") +
   # stat_summary(fun.data = "mean_se", geom = "errorbar") # adding error bars (standard error)
```


```{r}
#Graphing IHA DATA IN LAYERS WITH TWO DIFFERENT DF##

ggplot(data = IHA_output_predam, aes(x=Year, y = Reversals, group = 1)) +
  geom_line(colour = 'blue') +
  geom_point(colour = 'blue') +
  geom_line(data = IHA_output_postdam, aes(x=Year, y = Reversals, group = 1), colour = 'orange') +
  geom_point(data = IHA_output_postdam, aes(x=Year, y = Reversals, group = 1), colour = 'orange')  +
  ggtitle( "07NB001 \nReversals") +
  theme_bw()+  
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(plot.title =  element_text(size = 10, hjust = .5)) +
  theme(legend.position = "top right") 

```



