---
title: "Code to Save"
output: html_notebook
---

##Code snippets for use in the WBNP EF project##

#How to get a list of data frames of station metadata:
```{r}

#metadata subset - list of df
stations_subset_metadata <- list()
for (st in stations_subset) {data <- tidyhydat::hy_stn_data_coll(station_number = (st))
stations_subset_metadata[[st]] <- data.frame(data)
}
#this worked, can get a metadata list of lists.

```

#Same thing but to get daily flows in a list of df.
```{r}

dfList <- list()
for (st in stations_all) {data <- tidyhydat::hy_daily(station_number = (st))
dfList[[st]] <- data.frame(data)
}

```

##Take station metadata and make the station numbers into a list so they can be used int he tidyhat package for station number

```{r}

#Post-dam stations- get rid of blanks and make into a list

#select only the list of pre-dam station numbers
postdam_stations <- stations_to_analyze %>%
select(post.dam..1969.1989)

#delete rows that are na or blanks
postdam_stations <- postdam_stations[!(is.na(postdam_stations$post.dam..1969.1989) | postdam_stations$post.dam..1969.1989==""), ]

#make as a list of characters
as.list(as.character(postdam_stations))

#use this variable in the tidyhat package for station_number =
```

#compare two lists of station numbers to see if any are in common
This is useful for finding the same pre and post dam stations

```{r}

unique(postdam_stations[postdam_stations %in% predam_stations])

```

##Extract data and put into a list of df. This uses trycatch so it will continue to extract data even if a station number is null

```{r}

#Flow only with trycatch 
dfFlow_subset <- list()
for (st in stations_subset) {data <- tryCatch(tidyhydat::hy_daily_flows(station_number = (st)), error = function(e) NULL)
if (!is.null(data)){
dfFlow_subset[[st]] <- data.frame(data) 
}
}

```

## checking that the above worked ##

```{r}

for (df in dfFlow_subset) {
  print(unique(df$STATION_NUMBER, df$Parameter))
  print(unique(df$Parameter)) 
}

```

## Splitting the list of data frames into two separate lists: ones that have "B' or "not B", which has to do with ice coverage

```{r}

open_water = list()
year_round = list()

for (df in dfList) {
  if("B" %in% df$Symbol) {
    open_water = append(open_water, df)
  } else {
    year_round = append(year_round, df)
  }  }


```

## QA/QC checks on the above open water and ice data to make sure they are correctly separated
```{r}

for (df in open_water) {
  if ("B" %in% df$Symbol){
    print("Open water data are correct")
  } else {
    print("Open water data are incorrectly separated, there are stations with Ice")
  }
}
#add exception for no df in list
```

##Find how many consecutive Nas there are in a row and flag somehow.

Note: this code hasn't been working- so far I have been doing the IHA calculations and seeing which months are missing, then going back and deleting those years
```{r}


for (df in dfFlow_subset) {
  if (length(is.na(df$Value)) >= 14.0) {
    print("There are 14 or more NAs in a row for these dates: ")
} else {
    print("there are not 14 or more NAs in a row") #need to mutate a new column with "flagged"
  } }

#x = is.na(dfList_subset$'07DA045'$Value)
#rle(x)$lengths[rle(x)$values]
```

#Still trying...sad face...
```{r}
flow_subset_onelist[max(with(rle(is.na(x)), lengths[values])), by = Symbol]

```


```{r}

is.na.rle <- rle(is.na(flow_subset_onelist[,1]))
is.na.rle$Values <- is.na.rle$Values & is.na.rle$lengths >= 14
flow_subset_onelist[!inverse.rle(is.na.rle),]



```

Tryingt to make two lists, one with Nas >14, one with <14

```{r}

find_consec_nas <- function(df) {
  nas <- is.na(df$Value)
  rle_nas = fle(nas)
  
  df$flag_col <-ifelse(rle_nas$Value[rle_nas$lengths >= 14],
                       "more than 14", "less than 14")
                       return(df)
}

dfLevel_subset <- lapply(dfLevel_subset, find_consec_nas)


```



