---
title: "E_Flows_Station_DATA"
output: html_notebook
---


Note: This file is somewhat depreciated and we are no longer using this code and method to find pre and post dam stations or any stations for any year, but instead using the tile plots (see station summaries file) to find which stations within certain years. 

See "PART 2- SINGLE STATION WORKFLOW" ~line 230 for the workflow to tidy, calculate IHA variables, and graph one station, or alternatively the finished product in the "FT_Smith_Stns" file


1) Install packages 

```{r}
install.packages("tidyhydat")
install.packages("IHA")
install.packages("Rtools")
install.packages("IHA", repos="http://R-Forge.R-project.org")
install.packages("plyr")
install.packages("caTools")
install.packages("tidyverse")
install.packages("zoo")
install.packages("lubridate")
install.packages("png")
install.packages("reticulate")
install.packages("miniconda")

```

Import libraries
```{r}
library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
```

2) Download the HYDAT Database and save it in the same place as this file

```{r}
download_hydat()
```

3) Read in the stations for the Slave River Watershed and get a list of the station numbers

```{r}
Slave_River_watershed <-  read_csv( "https://raw.githubusercontent.com/wendyamonk/WBNP_analyses/main/HYDAT_Slave_River_watershed.csv?token=GHSAT0AAAAAAB4LKYXMDYYNQDQGBQ6ITTFQY42MPCA" )
SLave_River_watershed_Stations <- Slave_River_watershed %>%
  select(STATION_NU) %>%
summarise(STATION_NU = paste(STATION_NU,collapse = " , "))
message(SLave_River_watershed_Stations)

#select only the station numbers
predam_stations <- Slave_River_watershed_Stations %>%
select(station_number)

```

Copy the above output and paste it into "stations_all" 
```{r}
stations_all <- c("07AA001", "07AA002", "07AA003", "07AA004", "07AA007", "07AA008", "07AA009", "07AA010", "07AB002", "07AC001", "07AC002", "07AC003", "07AC004", "07AC005", "07AC006", "07AC007", "07AC008", "07AD001", "07AD002", "07AD003", "07AD004", "07AD005", "07AD006", "07AD007", "07AD008", "07AD009", "07AD010", "07AE001", "07AE002", "07AE003", "07AF001", "07AF002", "07AF003", "07AF004", "07AF005", "07AF008", "07AF009", "07AF010", "07AF011", "07AF012", "07AF013", "07AF014", "07AF015", "07AF016", "07AF906", "07AF907", 
              "07AF909", "07AF910", "07AG001", "07AG002", "07AG003", "07AG004", "07AG005", "07AG006", "07AG007", "07AG008", "07AH001", "07AH002", "07AH003", "07BA001", "07BA002", "07BA003", "07BB001", "07BB002", "07BB003", "07BB004", "07BB005", "07BB006", "07BB007", "07BB008", "07BB009", "07BB010", "07BB011", "07BB012", "07BB013", "07BB014", "07BB903", "07BB914", "07BC001", "07BC002", "07BC003", "07BC004", "07BC005", "07BC006", "07BC007", "07BE001", "07BE002", "07BE003", "07BE004", "07BF001", "07BF002", "07BF004", 
              "07BF006", "07BF007", "07BF008", "07BF009", "07BF010", "07BF905", "07BG004", "07BH001", "07BH003", "07BJ001", "07BJ002", "07BJ003", "07BJ004", "07BJ005", "07BJ006", "07BK001", "07BK002", "07BK003", "07BK004", "07BK005", "07BK006", "07BK007", "07BK008", "07BK009", "07BK010", "07BK012", "07CA001", "07CA002", "07CA003", "07CA004", "07CA005", "07CA006", "07CA008", "07CA010", "07CA011", "07CA012", "07CA013", "07CA901", "07CB001", "07CB002", "07CC001", "07CC002", "07CD001", "07CD002", 
              "07CD003", "07CD004", "07CD005", "07CD006", "07CD007", "07CD008", "07CD009", "07CE001", "07CE002", "07CE003", "07CE004", "07CE005", "07CE006", "07CE007", "07CE008", "07CE010", "07CE013", "07CE906", "07DA001", "07DA002", "07DA003", "07DA004", "07DA005", "07DA006", "07DA007", "07DA008", "07DA009", "07DA010", "07DA011", "07DA012", "07DA013", "07DA014", "07DA015", "07DA016", "07DA017", "07DA018", "07DA019", "07DA020", "07DA021", "07DA022", "07DA023", "07DA024", "07DA025", "07DA026", "07DA027", "07DA028", 
              "07DA029", "07DA030", "07DA032", "07DA033", "07DA034", "07DA035", "07DA037", "07DA038", "07DA039", "07DA040", "07DA041", "07DA042", "07DA044", "07DA045", "07DB001", "07DB002", "07DB003", "07DB004", "07DB005", "07DB006", "07DC001", "07DC002", "07DC003", "07DC004", "07DD001", "07DD002", "07DD003", "07DD004", "07DD005", "07DD006", "07DD007", "07DD008", "07DD009", "07DD010", "07DD011", "07EA001", "07EA002", "07EA004", "07EA005", "07EA006", "07EA007", "07EB001", "07EB002", "07EC001", "07EC002", "07EC003", 
              "07EC004", "07ED001", "07ED002", "07ED003", "07EE002", "07EE003", "07EE004", "07EE005", "07EE007", "07EE008", "07EE009", "07EE010", "07EE011", "07EF001", "07EF002", "07EF003", "07EF004", "07FA001", "07FA003", "07FA004", "07FA005", "07FA006", "07FA007", "07FB001", "07FB002", "07FB003", "07FB004", "07FB005", "07FB006", "07FB007", "07FB008", "07FB009", "07FB010", "07FB011", "07FC001", "07FC002", "07FC003", "07FC004", "07FC005", "07FC006", "07FC007", "07FD001", "07FD002", "07FD003", "07FD004", "07FD005", 
              "07FD006", "07FD007", "07FD008", "07FD009", "07FD010", "07FD011", "07FD012", "07FD013", "07FD014", "07FD015", "07FD016", "07FD017", "07FD019", "07FD020", "07FD901", "07FD908", "07FD910", "07FD912", "07FD913", "07FD921", "07FD934", "07GA001", "07GA002", "07GB001", "07GB002", "07GB003", "07GC001", "07GC002", "07GD001", "07GD002", "07GD003", "07GD004", "07GE001", "07GE002", "07GE003", "07GE004", "07GE005", "07GE006", "07GE007", "07GF001", "07GF002", "07GF003", "07GF004", "07GF005", "07GF006", "07GF007", 
              "07GF008", "07GG001", "07GG002", "07GG003", "07GH001", "07GH002", "07GH003", "07GH004", "07GH005", "07GH906", "07GJ001", "07GJ004", "07GJ005", "07HA001", "07HA002", "07HA003", "07HA005", "07HA902", "07HA914", "07HB001", "07HB002", "07HC001", "07HC002", "07HC907", "07HD001", "07HF001", "07HF002", "07JA001", "07JA002", "07JA003", "07JB001", "07JB002", "07JC001", "07JC002", "07JC003", "07JD001", "07JD002", "07JD003", "07JD004", "07JF002", "07JF003", "07JF004", "07JF005", "07KA002", "07KC001", "07KC003", 
              "07KC004", "07KC005", "07KE001", "07KF001", "07KF002", "07KF003", "07KF004", "07KF005", "07KF006", "07KF007", "07KF008", "07KF010", "07KF013", "07KF014", "07KF015", "07LB001", "07LB002", "07LC002", "07LC003", "07LD001", "07LD002", "07LD003", "07LE001", "07LE002", "07LE003", "07LE004", "07MA003", "07MA004", "07MA005", "07MB001", "07MC001", "07MC002", "07MC003", "07MD001", "07MD002", "07NA001", "07NA002", "07NA003", "07NA004", "07NA005", "07NA007", "07NA008", "07NB001", "07NB002", "07NB003", "07NB004", 
              "07NB005", "07NB006", "07NB007", "07NB008", "07NC001", "07NC002", "07NC003", "07NC004", "07NC005", "07NC006", "07NC007", "07NC008", "07NC009", "07PB002", "07QC008")

```

Df of metadata for all stations
```{r}
Stations_metadata_onelist <- hy_stn_data_coll(station_number = c("07AA001", "07AA002", "07AA003", "07AA004", "07AA007", "07AA008", "07AA009", "07AA010", "07AB002", "07AC001", "07AC002", "07AC003", "07AC004", "07AC005", "07AC006", "07AC007", "07AC008", "07AD001", "07AD002", "07AD003", "07AD004", "07AD005", "07AD006", "07AD007", "07AD008", "07AD009", "07AD010", "07AE001", "07AE002", "07AE003", "07AF001", "07AF002", "07AF003", "07AF004", "07AF005", "07AF008", "07AF009", "07AF010", "07AF011", "07AF012", "07AF013", "07AF014", "07AF015", "07AF016", "07AF906", "07AF907", 
              "07AF909", "07AF910", "07AG001", "07AG002", "07AG003", "07AG004", "07AG005", "07AG006", "07AG007", "07AG008", "07AH001", "07AH002", "07AH003", "07BA001", "07BA002", "07BA003", "07BB001", "07BB002", "07BB003", "07BB004", "07BB005", "07BB006", "07BB007", "07BB008", "07BB009", "07BB010", "07BB011", "07BB012", "07BB013", "07BB014", "07BB903", "07BB914", "07BC001", "07BC002", "07BC003", "07BC004", "07BC005", "07BC006", "07BC007", "07BE001", "07BE002", "07BE003", "07BE004", "07BF001", "07BF002", "07BF004", 
              "07BF006", "07BF007", "07BF008", "07BF009", "07BF010", "07BF905", "07BG004", "07BH001", "07BH003", "07BJ001", "07BJ002", "07BJ003", "07BJ004", "07BJ005", "07BJ006", "07BK001", "07BK002", "07BK003", "07BK004", "07BK005", "07BK006", "07BK007", "07BK008", "07BK009", "07BK010", "07BK012", "07CA001", "07CA002", "07CA003", "07CA004", "07CA005", "07CA006", "07CA008", "07CA010", "07CA011", "07CA012", "07CA013", "07CA901", "07CB001", "07CB002", "07CC001", "07CC002", "07CD001", "07CD002", 
              "07CD003", "07CD004", "07CD005", "07CD006", "07CD007", "07CD008", "07CD009", "07CE001", "07CE002", "07CE003", "07CE004", "07CE005", "07CE006", "07CE007", "07CE008", "07CE010", "07CE013", "07CE906", "07DA001", "07DA002", "07DA003", "07DA004", "07DA005", "07DA006", "07DA007", "07DA008", "07DA009", "07DA010", "07DA011", "07DA012", "07DA013", "07DA014", "07DA015", "07DA016", "07DA017", "07DA018", "07DA019", "07DA020", "07DA021", "07DA022", "07DA023", "07DA024", "07DA025", "07DA026", "07DA027", "07DA028", 
              "07DA029", "07DA030", "07DA032", "07DA033", "07DA034", "07DA035", "07DA037", "07DA038", "07DA039", "07DA040", "07DA041", "07DA042", "07DA044", "07DA045", "07DB001", "07DB002", "07DB003", "07DB004", "07DB005", "07DB006", "07DC001", "07DC002", "07DC003", "07DC004", "07DD001", "07DD002", "07DD003", "07DD004", "07DD005", "07DD006", "07DD007", "07DD008", "07DD009", "07DD010", "07DD011", "07EA001", "07EA002", "07EA004", "07EA005", "07EA006", "07EA007", "07EB001", "07EB002", "07EC001", "07EC002", "07EC003", 
              "07EC004", "07ED001", "07ED002", "07ED003", "07EE002", "07EE003", "07EE004", "07EE005", "07EE007", "07EE008", "07EE009", "07EE010", "07EE011", "07EF001", "07EF002", "07EF003", "07EF004", "07FA001", "07FA003", "07FA004", "07FA005", "07FA006", "07FA007", "07FB001", "07FB002", "07FB003", "07FB004", "07FB005", "07FB006", "07FB007", "07FB008", "07FB009", "07FB010", "07FB011", "07FC001", "07FC002", "07FC003", "07FC004", "07FC005", "07FC006", "07FC007", "07FD001", "07FD002", "07FD003", "07FD004", "07FD005", 
              "07FD006", "07FD007", "07FD008", "07FD009", "07FD010", "07FD011", "07FD012", "07FD013", "07FD014", "07FD015", "07FD016", "07FD017", "07FD019", "07FD020", "07FD901", "07FD908", "07FD910", "07FD912", "07FD913", "07FD921", "07FD934", "07GA001", "07GA002", "07GB001", "07GB002", "07GB003", "07GC001", "07GC002", "07GD001", "07GD002", "07GD003", "07GD004", "07GE001", "07GE002", "07GE003", "07GE004", "07GE005", "07GE006", "07GE007", "07GF001", "07GF002", "07GF003", "07GF004", "07GF005", "07GF006", "07GF007", 
              "07GF008", "07GG001", "07GG002", "07GG003", "07GH001", "07GH002", "07GH003", "07GH004", "07GH005", "07GH906", "07GJ001", "07GJ004", "07GJ005", "07HA001", "07HA002", "07HA003", "07HA005", "07HA902", "07HA914", "07HB001", "07HB002", "07HC001", "07HC002", "07HC907", "07HD001", "07HF001", "07HF002", "07JA001", "07JA002", "07JA003", "07JB001", "07JB002", "07JC001", "07JC002", "07JC003", "07JD001", "07JD002", "07JD003", "07JD004", "07JF002", "07JF003", "07JF004", "07JF005", "07KA002", "07KC001", "07KC003", 
              "07KC004", "07KC005", "07KE001", "07KF001", "07KF002", "07KF003", "07KF004", "07KF005", "07KF006", "07KF007", "07KF008", "07KF010", "07KF013", "07KF014", "07KF015", "07LB001", "07LB002", "07LC002", "07LC003", "07LD001", "07LD002", "07LD003", "07LE001", "07LE002", "07LE003", "07LE004", "07MA003", "07MA004", "07MA005", "07MB001", "07MC001", "07MC002", "07MC003", "07MD001", "07MD002", "07NA001", "07NA002", "07NA003", "07NA004", "07NA005", "07NA007", "07NA008", "07NB001", "07NB002", "07NB003", "07NB004", 
              "07NB005", "07NB006", "07NB007", "07NB008", "07NC001", "07NC002", "07NC003", "07NC004", "07NC005", "07NC006", "07NC007", "07NC008", "07NC009", "07PB002", "07QC008"))

```
## FLOW DATA TRIAL AND ERROR -- to get stns to analyze ##

Pre-dam stations- get rid of blanks and make into a list

```{r} 

#select only the list of pre-dam station numbers
predam_stations <- stations_to_analyze %>%
select(pre.dam..1947.1967)

#delete rows that are na or blanks
predam_stations <- predam_stations[!(is.na(predam_stations$pre.dam..1947.1967) | predam_stations$pre.dam..1947.1967==""), ]

#make as a list of characters
as.list(as.character(predam_stations))

#use this variable in the tidyhat package for station_number = 

```


#this is now a function : "Stations_pre_Bennett" and can run this to find a list of stations for predam years

```{r}

#GETTING PREDAM STATIONS- 20 YRS AND PUT INTO A LIST 

Stations_flow_1947_67 <- Stations_metadata_onelist %>%
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT, OPERATION ) %>%
  #only flow data
filter(DATA_TYPE == 'Flow') %>%
  #only list years that start before 1967
filter(Year_from < 1967) %>%
  #only list years that end after 1940
filter(Year_to >= 1940)  %>%
group_by(STATION_NUMBER) %>%
  #add max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #only choose min dates that start before 1950
filter(min_date <= 1950) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, min_date, max_date)  %>%
  #get rid of max date and min date duplicates
  distinct()

View(Stations_flow_1947_67)

predam_flow_stn <- Stations_flow_1947_67 %>%
select(STATION_NUMBER)

View(predam_flow_stn)

#make as a list of characters
as.list(as.character(predam_flow_stn))


```

##NOW WE HAVE A LIST OF PREDAM FLOW STNS AND CAN START TO TIDY AND DO ANALYSES ON THEM ##

##NEXT LOOK AT POST DAM AND SEE IF ANY ARE IN COMMON -- doesn't look like it though.

##THEN FIND ONES IN FORT SMITH AREA

##NEXT WRITE CODE TO FIND POST-DAM STATIONS FROM THE FULL LIST #this is a function : "Stations_post_Bennett"

```{r}

Stations_flow_1969_89 <- Stations_metadata_onelist %>%
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT, OPERATION ) %>%
  #only flow data
filter(DATA_TYPE == 'Flow') %>%
  #only list years that start before 1990
filter(Year_from < 1990) %>%
  #only list years that end after 1969
filter(Year_to >= 1969)  %>%
group_by(STATION_NUMBER) %>%
  #add max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #only choose min dates that start before 1950
filter(min_date <= 1972) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, min_date, max_date)  %>%
  #get rid of max date and min date duplicates
  distinct()

#gives 135 stations

View(Stations_flow_1969_89)


postdam_flow_stn <- Stations_flow_1969_89 %>%
select(STATION_NUMBER)

View(postdam_flow_stn)

#make as a list of characters
as.list(as.character(postdam_flow_stn))



```

#put non csv stations from above code into a list and compare with the csv list 

```{r}

Stations_flow_1969_89_noncsv <- Stations_flow_1969_89  %>% 
  select(STATION_NUMBER)

Stations_flow_1969_89_noncsv <- Stations_flow_1969_89_noncsv[!(is.na(Stations_flow_1969_89_noncsv$STATION_NUMBER) | Stations_flow_1969_89_noncsv$STATION_NUMBER==""), ]

as.list(as.character(Stations_flow_1969_89_noncsv))

```

#compare two lists of station numbers to see if any are in common
This is useful for finding the same pre and post dam stations

```{r}

unique(postdam_stations[postdam_stations %in% predam_stations])

```

#PROCEED WITH COMPARING PRE AND POST DAM STATIONS USING TIDYING METHOD
```{r}

#predam_flow_stn vs postdam_flow_stn

unique(postdam_flow_stn[postdam_flow_stn %in% predam_flow_stn])

#this didn't work but stations in common are: 07AA007, 07BE001, 07ED001, 07EF001, 07FA001, 07FD001, 07FD002

#one near ft smith: 07NB001

```


##PART 2- SINGLE STATION WORKFLOW ##
#NOW TRY TO TIDY DATA AND GET A FORT SMITH DATA SET

```{r} 
Fort_Smith_Flows <- tidyhydat::hy_daily_flows(station_number = "07NB001")

```

#Get a df of predam data 

```{r} 
flow_data_FS <- Fort_Smith_Flows %>%
  select(Date, Value) %>%
  na.omit()

flow_data_predam_FS <- flow_data_FS %>%
  #filter(Date >= 1947-01-01)
  filter(Date < "1967-01-01" & Date > "1947-01-01")

#Make the dates into dates in R
#as.Date(flow_data$Date)

```

#Get a df of postdam data 

```{r} 

#post dam data first 
flow_data_FS <- Fort_Smith_Flows %>%
  select(Date, Value) %>%
  na.omit()

flow_data_postdam_FS <- flow_data_FS %>%
  filter(Date > "1969-01-01" & Date < "1990-01-01")

View(flow_data_postdam_FS) 

```

##Can see if there are missing years by finding the unique years

```{r} 

years <- format(flow_data_predam_FS$Date, "%Y")

unique_years <- unique(years)

all_years <- seq(min(unique_years), max(unique_years), by = 1)

missing_years <- setdiff(all_years, unique_years)

View(missing_years) #should be zero

any(missing_years) #says if there is anything in the list
```

#Make the dates into a day-month-year form
```{r}

##This worked
#This works but just gives date column
#flow_data_date$Date <- format(flow_data$Date, "%d-%m-%Y")

#Use mutate to keep both the date and value column 
flow_data_date_FS <- flow_data_FS %>%
mutate(Date = format(Date,"%d-%m-%Y"))


```


```{r}
## Convert data to zoo object
flow_data <- zoo(flow_data_date$Value, order.by = as.Date(as.character(flow_data_date$Date), format = "%d-%m-%Y"))

## Run IHA analyses
group1_output <- group1(flow_data, year = "water", FUN = median)
group2_output <- group2(flow_data, year = "water", mimic.tnc = TRUE)
group3_output <- group3(flow_data, year = "water", mimic.tnc = FALSE)
group4_output <- group4(flow_data, year = "water")
group5_output <- group5(flow_data, year = "water")

## Convert outputs
group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)


## Create output file
IHA_output <- bind_cols(list(group1_output, group2_output, group3_output,
                             group4_output, group5_output))

## Write output file - remember to change output file name
write.csv(IHA_output, "YourNewShinyOutput.csv")

```


```{r} 

#Graphing IHA DATA IN LAYERS WITH TWO DIFFERENT DF##

ggplot(data = IHA_output_predam, aes(x=Year, y = Reversals, group = 1)) +
  geom_line(colour = 'blue') +
  geom_point(colour = 'blue') +
  geom_line(data = IHA_output_postdam, aes(x=Year, y = Reversals, group = 1), colour = 'orange') +
  geom_point(data = IHA_output_postdam, aes(x=Year, y = Reversals, group = 1), colour = 'orange')  +
  ggtitle( "07NB001 \nReversals") +
  theme_bw()+  
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(plot.title =  element_text(size = 10, hjust = .5)) +
  theme(legend.position = "top right") 

```


#Getting station data for Wendy's post doc student
```{r}

PAD_hist <- read_csv("PAD_area_historic_gauge_data.csv")

Pad_stations <- PAD_hist %>%
select("Gauge_ID") %>%
summarize(Gauge_ID = paste(Gauge_ID, sep = " , "))
message(Pad_stations)

```


```{r}
PAD_metadata <- tidyhydat::hy_stn_data_coll(station_number = c("07DD001", "07DD011", "07KF015", "07KF003", "07KF002", "07MD001", "07NA001", "07NA007", "07NA008", "07KC005", "07KC001", "07DA042", "07DA040", "07DC004", "07DA041", "07DA023", "07DA024", "07DA029", "07DA034", "07DA028", "07DA033", "07DA037", "07DA025", "07DA035", "07DA030", "07DA027", "07DA026", "07DA045", "07DC003"))

PAD_metadata <- PAD_metadata %>%
  filter(DATA_TYPE == "Flow" | DATA_TYPE == "Water Level")

PAD_Flows <- tidyhydat::hy_daily_flows(station_number = c("07DD001", "07DD011", "07KF015", "07KF003", "07KF002", "07MD001", "07NA001", "07NA007", "07NA008", "07KC005", "07KC001", "07DA042", "07DA040", "07DC004", "07DA041", "07DA023", "07DA024", "07DA029", "07DA034", "07DA028", "07DA033", "07DA037", "07DA025", "07DA035", "07DA030", "07DA027", "07DA026", "07DA045", "07DC003"))

PAD_Water_Levels <- tidyhydat::hy_daily_levels(station_number = c("07DD001", "07DD011", "07KF015", "07KF003", "07KF002", "07MD001", "07NA001", "07NA007", "07NA008", "07KC005", "07KC001", "07DA042", "07DA040", "07DC004", "07DA041", "07DA023", "07DA024", "07DA029", "07DA034", "07DA028", "07DA033", "07DA037", "07DA025", "07DA035", "07DA030", "07DA027", "07DA026", "07DA045", "07DC003"))


write.csv(PAD_Flows, "PAD_historic_flows.csv")

write.csv(PAD_Water_Levels, "PAD_historic_water_levels.csv")


```
