
## FUNCTIONS USED TO TIDY STATION META DATA FOR FLOW ##


1) Returns a DF of all pre-Bennett dam stations from 1947-1967. Input: a df of the station metadata from hy_stn_data_coll in hydat.

```{r}

#need to edit so user inputs the df and the year range and it returns a DF, rather than pre and post function

Stations_pre_Bennett <- function(df, STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT) {
  
library(dplyr)
df <- df %>%
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT ) %>%
  #only flow data
filter(DATA_TYPE == 'Flow') %>%
  #only list years that start before 1967
filter(Year_from < 1967) %>%
  #only list years that end after 1940
filter(Year_to >= 1940)  %>%
group_by(STATION_NUMBER) %>%
  #add max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #only choose min dates that start before 1950
filter(min_date <= 1950) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, min_date, max_date)  %>%
  #get rid of max date and min date duplicates
  distinct()
  return(df)
}

```

2) Returns a DF of post-Bennett dam stations from 1969-1989. Input: a df of the station metadata from hy_stn_data_coll in hydat.

```{r}
#need to edit so user inputs the df and the year range and it returns a DF, rather than pre and post function

Stations_post_Bennett <- function(df,STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT ) {
  library(dplyr)
  df <- df %>%
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT, OPERATION ) %>%
  #only flow data
filter(DATA_TYPE == 'Flow') %>%
  #only list years that start before 1990
filter(Year_from < 1990) %>%
  #only list years that end after 1969
filter(Year_to >= 1969)  %>%
group_by(STATION_NUMBER) %>%
  #add max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #only choose min dates that start before 1950
filter(min_date <= 1972) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, min_date, max_date)  %>%
  #get rid of max date and min date duplicates
  distinct()
  return(df)
}


```

## FUNCTIONS USED TO TIDY FLOW DATA ##
## Prep for IHA calculations ## 

3) Returns a DF of pre-Bennett years (1947-67) from a df of one or more stations, that can then be put into IHA calculation sheet. Input: a single df of the flow data for one or more stations from hy_daily_flows in hydat.
```{r}

#need to edit so user inputs the df and the year range and it returns a DF, rather than having a pre and post function

flow_data_predam <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit()%>%
  filter(Date < "1966-10-01" & Date > "1947-10-01") %>%
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

```

4) Returns a DF of post-Bennett years (1969-89) from a df of one or more stations that can then be put into IHA calcs. Input: a single df of the flow data for one or more stations from hy_daily_flows in hydat.
 
```{r}

#need to edit so user inputs the df and the year range and it returns a DF, rather than having a pre and post function
flow_data_postdam <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit() %>%
  filter(Date >= "1968-10-01" & Date < "1991-10-01") %>%
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

```

5) Find out if there are missing years in the data set. Will return "false" if there are no missing years, or a df of missing years. Needs to be in the tidyhydat format of year-month-day, so use before run in function 3, 4 (Stations_post_Bennett etc).

```{r} 
yrs_missing <- function(df, Date) {
years <- format(df$Date, "%Y")
unique_years <- unique(years)
all_years <- seq(min(unique_years), max(unique_years), by = 1)
missing_years <- setdiff(all_years, unique_years)

View(missing_years) #should be zero
any(missing_years) #says if there is anything in the list
print(missing_years)
}

```

## IHA calculations ##

```{r}

calc_IHA <- function(data){
  
flow_data <-  {{data}} #%>%
#  select(Date, Value)
  

flow_data <- zoo(flow_data$Value, order.by = as.Date(as.character(flow_data$Date), format = "%d-%m-%Y"))
## Run IHA analyses
group1_output <- group1(flow_data, year = "water", FUN = median)
group2_output <- group2(flow_data, year = "water", mimic.tnc = TRUE)
group3_output <- group3(flow_data, year = "water", mimic.tnc = FALSE)
#group4_output <- group4(flow_data, year = "water")
group5_output <- group5(flow_data, year = "water")

## Convert outputs
group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
#group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)

## Create output dataframe 
#Need to figure out how to join all 4 dfs even if they are diff sizes - use reduce? 
#group4 data left out for now (low/high pulse numbers)

IHA_output <- bind_cols(list(group1_output, group2_output, group3_output, group5_output)) #group4_output

#make the years the column names instead of the rownames
IHA_output <- tibble::rownames_to_column(IHA_output, "Year") 

}

```


```{r}

```



```{r}
```


