
## FUNCTIONS USED TO TIDY STATION META DATA FOR FLOW ##


1) Returns a DF of all pre-Bennett dam stations from 1947-1967. Input: a df of the station metadata from hy_stn_data_coll in hydat.

```{r}
Stations_pre_Bennett <- function(df, STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT) {
  
library(dplyr)
df <- df %>%
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT ) %>%
  #only flow data
filter(DATA_TYPE == 'Flow') %>%
  #only list years that start before 1967
filter(Year_from < 1967) %>%
  #only list years that end after 1940
filter(Year_to >= 1940)  %>%
group_by(STATION_NUMBER) %>%
  #add max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #only choose min dates that start before 1950
filter(min_date <= 1950) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, min_date, max_date)  %>%
  #get rid of max date and min date duplicates
  distinct()
  return(df)


}

```

2) Returns a DF of post-Bennett dam stations from 1969-1989. Input: a df of the station metadata from hy_stn_data_coll in hydat.

```{r}
Stations_post_Bennett <- function(df,STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT ) {
  library(dplyr)
  df <- df %>%
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to, MEASUREMENT, OPERATION ) %>%
  #only flow data
filter(DATA_TYPE == 'Flow') %>%
  #only list years that start before 1990
filter(Year_from < 1990) %>%
  #only list years that end after 1969
filter(Year_to >= 1969)  %>%
group_by(STATION_NUMBER) %>%
  #add max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #only choose min dates that start before 1950
filter(min_date <= 1972) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, min_date, max_date)  %>%
  #get rid of max date and min date duplicates
  distinct()
  return(df)
}


```

## FUNCTIONS USED TO TIDY FLOW DATA ##
## Prep for IHA calculations ## 

3) Returns a DF of pre-Bennett years (1947-67) from a df of one or more stations, that can then be put into IHA calculation sheet. Input: a single df of the flow data for one or more stations from hy_daily_flows in hydat.
```{r}

flow_data_predam <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit()%>%
  filter(Date < "1966-10-01" & Date > "1947-10-01") %>%
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

```

4) Returns a DF of post-Bennett years (1969-89) from a df of one or more stations that can then be put into IHA calcs. Input: a single df of the flow data for one or more stations from hy_daily_flows in hydat.
 
```{r}

flow_data_postdam <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit() %>%
  filter(Date >= "1968-10-01" & Date < "1991-10-01") %>%
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

```

5) Find out if there are missing years in the data set. Will return "false" if there are no missing years, or a df of missing years. Needs to be in the tidyhydat format of year-month-day, so use before run in function 3, 4 (Stations_post_Bennett etc).

```{r} 
yrs_missing <- function(df, Date) {
years <- format(df$Date, "%Y")
unique_years <- unique(years)
all_years <- seq(min(unique_years), max(unique_years), by = 1)
missing_years <- setdiff(all_years, unique_years)

View(missing_years) #should be zero
any(missing_years) #says if there is anything in the list
print(missing_years)
}

```

```{r}

```


```{r}



```


```{r}



```

