## FUNCTIONS USED TO TIDY STATION META DATA FOR FLOW ##


Line graph of one IHA output variable at a time. Coloured by time period (pre/post dam).

```{r}

ggplot(data = IHA_both, aes(x =Year, y = Reversals, group = 1) ) +
          geom_line(aes(colour = Time_Period) ) +
      geom_point(aes(colour = Time_Period)) +
  ggtitle( "07NB001 \n Reversals") +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.position=c(0.87, 0.8), legend.title=element_text(size=8), legend.text=element_text(size=7)) +
  geom_line(data = IHA_predam_summary, y =  Reversals) +
   scale_colour_discrete(na.translate = F)
  

```

Line graphs of all IHA variables, where x= year and y = each variable in succession.
Coloured by time period (pre/post dam).

```{r}

dontgraphme <- c("Year", "Time_Period", "Zero flow days", '1 Day Min', '3 Day Min', '7 Day Min', '30 Day Min', '90 Day Min', '1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max','October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September')

list_plots <- vector('list', length(cols))

for (col in names(IHA_both)[!names(IHA_both) %in% dontgraphme]) {
  print(ggplot(data = IHA_both, aes(x = Year, y = IHA_both[,col], group = 1 )) +
          geom_line(aes(colour = Time_Period) ) +
      geom_point(aes(colour = Time_Period)) +
  ggtitle(col)  +
    ylab(col) +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.position=c(0.87, 0.8), legend.title=element_text(size=8), legend.text=element_text(size=7))+
       scale_colour_discrete(na.translate = F)) 
   print(list_plots[col])

}


```
Boxplots of IHA variables: 

Y = flow and x = the IHA variables passed in to the function.

Coloured by time period (pre/post dam).

Minimum flow
```{r}

boxplot_min <- function(df) {

p <-  IHA_both %>% 
  na.omit() %>%
  select('Year', '1 Day Min', '3 Day Min', '7 Day Min', '30 Day Min', '90 Day Min', 'Time_Period') %>%
  pivot_longer(., cols = c('1 Day Min', '3 Day Min', '7 Day Min','30 Day Min' , '90 Day Min'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
geom_boxplot() +
  geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
       theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
  scale_fill_manual(values = c("orange","light blue"), 
                    labels = c("Pre Dam", "Post Dam")) +
  xlab("") +
  ylab("Flow") 

return(p)

}

```

Boxplots of IHA variables: 

Maximum Flow
```{r}


boxplot_max <- function(df) {


p <-  IHA_both %>% 
  na.omit() %>%
  select('Year', '1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max','Time_Period') %>%
  pivot_longer(., cols = c('1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max', '90 Day Max'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
geom_boxplot() +
  geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
  scale_fill_manual(values = c("orange","light blue"), 
                    labels = c("Pre Dam", "Post Dam")) +
  xlab("") +
  ylab("Flow") 

return(p)

}


```

Boxplots of IHA variables: 

Flow by month
```{r}

boxplot_month <- function(df) {
  
p <-  IHA_both %>% 
  na.omit() %>%
  select('Year', 'October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September', 'Time_Period') %>%
  pivot_longer(., cols = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
geom_boxplot() +
  #geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
  scale_fill_manual(values = c("orange","light blue"), 
                    labels = c("Pre Dam", "Post Dam")) +
 aes(x = fct_inorder(Var)) +
  xlab("Month") +
  ylab("Flow") 

return(p)

}

```

## IGNORE THIS### 

Trying ribbons from here down. 

```{r}


 
```

Trying so hard to get a ribbon behind the boxplots....

```{r}
#DF of Month, 25%, 50%, 75%
df <- IHA_output_predam %>%
  select('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September') 
 IQ <- sapply(df, quantile, probs=c(0.25, 0.5, 0.75), na.rm = TRUE)
IHA_quartiles_predam <- as.data.frame(IQ)
df1 <- t(IHA_quartiles_predam)
df2 <- as.data.frame(df1)
df3 <- rownames_to_column(df2, "Month") 

write.csv(df3, "IQ-predam.csv")
```

Tidy in excel, make a new df and load back in

```{r}

IQ_pre <- read.csv("IQ-predam-edit.csv")

 IQ_pre %>% 
  select(Month, Variable, Value) %>%
 # filter(Variable == "IQ25") %>%
ggplot(aes(x=Month, y = Value)) +
  geom_line(aes(group= Variable, colour = Variable)) 

 ggplot(aes(x=Month, y = Value)) +
  geom_line(aes(x=Month, y = Value, group= Variable, colour = Variable)) 

 

```

make max/min columns to be able to use with geomribbon

```{r}

maxIQ <-IQ_pre %>%
  select(Month, Variable, Value)%>%
  filter(Variable == "IQ75") %>%
  mutate("max" = Value) %>%
  select(Month, max)
  
minIQ <- IQ_pre %>%
  select(Month, Variable, Value)%>%
filter(Variable == "IQ25")%>%
  mutate("min" = Value) %>%
  select(Month, min)

maxmin <- inner_join(maxIQ, minIQ, by= "Month") %>%
  mutate(newcol = c(1,2,3,4,5,6,7,8,9,10,11,12)) %>%
  select(newcol, max, min, Month)
  

```


```{r}
p +
  geom_area(maxIQ, mapping = aes(x= Month, y=Value, fill=Value))

```


```{r}


ggplot() +
 # geom_area(maxmin, mapping = aes(x= newcol, y=max)) +
 # geom_area(maxmin, mapping = aes(x=newcol, y=min))
  geom_ribbon(data=maxmin, aes(x= newcol, ymin= min, ymax= max, group=1, alpha = .2, fill = "light pink"))

```


```{r}

ggplot() +
    geom_line(data = IQ_pre, aes(x=Month, y = Value, group= Variable, colour = Variable)) 


```


```{r}
IHA_output_predam %>% 
  na.omit() %>%
  select('Year', 'October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September', 'Time_Period') %>%
  pivot_longer(., cols = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September'), names_to = "Var", values_to = "Val") %>%
ggplot(aes(x = Var, 
           y = Val, 
           group = Var,
           color = Var)) +
  stat_summary(fun.y = "mean", geom = "point") # +
 # stat_summary(fun.y = "mean", geom = "line") +
   # stat_summary(fun.data = "mean_se", geom = "errorbar") # adding error bars (standard error)
```


```{r}

ggplot(IHA_output_predam, aes(Time, Value)) + geom_smooth(stat = 'summary', fun.data = function(y) data.frame(ymin = quantile(y, .1), y = mean(y), ymax = quantile(y, .9)))


```


```{r}

```






