---
title: "R Notebook"
output: html_notebook
---

Import libraries
```{r}
library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
```

Tidy data for water level: Keep only the stations with 20 years of data, regardless of if they are c/s/m 
note: taking a different approach and subtracting the years of operation and then adding for each station, see #5
-note: need to add this to a separate water level r code file, keep these separate for now.


```{r}

Stations_metadata_wl_20yrsdata <- Stations_metadata_onelist %>%
  #getting rid of operation type (we don't care if its s/c/m for now)
select(STATION_NUMBER, DATA_TYPE, Year_from, Year_to ) %>%
  #only water level data
filter(DATA_TYPE == 'Water Level') %>%
  #keep station numbers together
group_by(STATION_NUMBER) %>%
  #make a new column with the max and min dates
mutate(min_date = min(Year_from),
            max_date = max(Year_to)) %>%
  #select only the max and min dates rather than the years to get rid of duplicate max and min dates
select(STATION_NUMBER, DATA_TYPE, max_date, min_date) %>%
  #get rid of max date and min date duplicates
  distinct() %>%
  #take the diff between max and min date
mutate(diff = max_date - min_date) %>%
  #keep only those with greater than or equal to 20 years of data
filter(diff >= 20)  
  
View(Stations_metadata_wl_20yrsdata)

Make a list of stations with 20 years or more of water level data to be able to use in the tidyhydat function hy_daily_levels and hy_stn_data_coll
-note: need to add this to a separate water level r code file, keep these separate for now.

```{r}
#select only the station numbers and make them characters
WL_ST_20yrs = Stations_metadata_wl_20yrsdata %>% 
select(STATION_NUMBER)
as.list(as.character(WL_ST_20yrs))

#make a concatenated list 
st_nu_WL <- c()
#loop through the df of only station numbers and append to the concatenated list
for (st in WL_ST_20yrs) {
 new_var = WL_ST_20yrs$STATION_NUMBER + ","
  #append(st_nu_WL, new_var)
  st_nu_WL = append(st_nu_WL, new_var)    # Appending new value to vector
}

#can use st_nu_WL as station number for hy_daily_levels to gather wl data for those stations

```

Test with just WL to see if can separate by c/s/m
This isn't working, ignore for now
-note: need to add this to a separate water level r code file, keep these separate for now.

```{r}
Stations_20yrs_metadata <- hy_stn_data_coll(station_number = st_nu_WL)

Stations_20yrs_wl_metadata <- Stations_wl_20yrs_metadata %>%
  #only water level data
filter(DATA_TYPE == 'Water Level') %>%
  summarize(STATION_NUMBER, OPERATION) %>%
  distinct()
#casewhen isn't working- it wont group see station numbers together
mutate(type = case_when(OPERATION == "Seasonal" ~ "Seasonal Only",
                          OPERATION == "Miscellaneous" ~ "Miscellaneous Only",
                          OPERATION == "Continuous" ~ "Continuous Only",
                          (OPERATION == "Seasonal" & OPERATION == "Continuous") ~ "Seasonal and Continuous ",
                          (OPERATION == "Seasonal" & OPERATION == "Miscellaneous") ~ "Seasonal and Miscellaneous",
                          (OPERATION == "Continuous" & OPERATION == "Miscellaneous") ~ "Continuous and Miscellaneous")
                          )

View(Stations_20yrs_wl_metadata)

#50 wl stations with varying amounts of C,S,M
```


