
## Data Visualization for stations on the PAD using Percent Change


```{r}

library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)

```

Station to practice on:
```{r}

pr_stn <- tidyhydat::hy_daily_flows(station_number = "07HA001")

```

Functions for predam and postdam flows
```{r}

#edit so user inputs dates they want, rather than having a pre/post dam flow function - do in the functions section

predam_flow <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit()%>%
  filter( Date >= "1915-10-01" & Date < "1931-10-01") %>% 
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

postdam_flow <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit() %>%
  filter(Date >= "1972-10-01" & Date < "1992-10-01") %>%
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

```

Getting flow for those years
```{r}

pr_stn_pre <- predam_flow(pr_stn, Date, Value)
pr_stn_post <- postdam_flow(pr_stn, Date, Value)

```

Predam IHA Variables 
```{r}

library(tibble)

IHA_output <- calc_IHA(pr_stn_pre)

IHA_output_pre <- IHA_output %>%
 add_column(Time_Period = "Pre Dam" )  %>%
  add_column('post_years'= c(1973, 1974, 1975, 1976, 1977, 1978, 1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1989)) %>%
mutate(post_years = as.character(post_years))

```

Postdam IHA Variables 

```{r}


IHA_output <- calc_IHA(pr_stn_post)

IHA_output_post <- IHA_output %>%
 # add_row(Year = c("1968", "1967") )%>%
 add_column(Time_Period = "Post Dam" )

```

Line graphs -  Max flow predam vs post dam
```{r}

ggplot() +
  geom_line(data = IHA_output_post, aes(x =Year, y = Max, group = 1), colour = "red", size = 0.75)  + 
      geom_point() + 
  annotate(geom="text", x=16, y=250, label="Maximum flow, post-dam, 1973-1992",
              color="red") +
  
  #geom_line(data = IHA_output_post, aes(x =Year, y = Min, group = 1), colour = "blue" ) + 
   #   geom_point() + 
  
  geom_line(data = IHA_output_pre, aes(x = post_years, y = Max, group = 1), colour = "black", linewidth = 1)  + 
      geom_point() +
  annotate(geom="text", x=14, y=140, label="Maximum flow, pre-dam: 1916-1931",
              color="black") +

  ggtitle( "Maximum River Flow Pre and Post Bennet Dam for a Station on the Peace River") +
  ylab("Flow")+
  xlab("Post-dam years") +
  theme_bw() +
   theme(plot.title = element_text(size = 10, hjust = .5, face = "bold")) # +
    #theme(axis.text.x = element_blank()) #+

  
    # theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  #  theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
#  theme(legend.position=c(0.87, 0.8), legend.title=element_text(size=8), legend.text=element_text(size=7)) +
#   scale_colour_discrete(na.translate = F)
```

Calculate percentiles, to get a range of variability within 0.33, 0.64 percentiles: 
```{r}
#function to calculate both quartiles, 0.33, 0.63, in one df
quant_calc <- function(x) {

quantiles <- quantile(x, c(0.33, 0.64))
names(quantiles) <- paste0(names(x), "_quantile")
return(quantiles)
}
#not as good for data tidying, otherwise need to manually rename columns
```

Calculate quartiles, rename columns, and then join dfs together:
```{r}

#functions to calculate each quartiles, 0.33, 0.63, in separate dfs

quant_calc_33 <- function(x){
  quant <- quantile(x, c(0.33))
  return(quant)
}

quant_calc_64 <- function(x){
  quant <- quantile(x, c(0.64))
  return(quant)
}

#function to rename column, use after calc quartiles 
rename_cols <- function(df, suffix) {
  names(df) <- paste0(names(df), suffix)
  return(df)
}
```

Rename columns and join 0.33 and 0.64 dfs together 
```{r}

name <- names(pre_33)

  pre_33 <- IHA_output_pre %>%
    select(-c("Low pulse length")) %>%
    summarise(across(October:Reversals, quant_calc_33)) %>%
    rename_cols("_q33") %>%
    add_row(.before = 2, .rows = 18) %>%
    fill(name) %>% 
  add_column('post_years'= c(1973, 1974, 1975, 1976, 1977, 1978, 1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1989, 1990, 1991, 1992))

name <- names(pre_64)
pre_64 <- IHA_output_pre %>%
  select(-c("Low pulse length")) %>%
  summarise(across(October:Reversals, quant_calc_64)) %>%
  rename_cols("_q64") %>%
  add_row(.before = 2, .rows = 18) %>% 
  fill(name) %>%
add_column('post_years'= c(1973, 1974, 1975, 1976, 1977, 1978, 1979, 1980, 1981, 1982, 1983, 1984, 1985, 1986, 1987, 1989, 1990, 1991, 1992))
  
#join together 
df_quartiles <- merge(pre_33, pre_64, by = "post_years" ) %>%
mutate(post_years = as.character(post_years))

```

Finally, graph an IHA variable for the post-dam period with the 0.33-0.64 percentiles from the pre-dam period 
```{r}
ggplot() +
  geom_line(data = IHA_output_post, aes(x =Year, y = Max, group = 1), colour = "red", size = 0.75)  + 
      geom_point() + 
   annotate(geom="text", x=16, y=250, label="Maximum flow, post-dam, 1973-1992",
              color="red") +
 #q33 line
  geom_line(data = df_quartiles, aes(x = post_years, y = Max_q33, group = 1), linetype = 'dotted', colour = "black", size = 0.5)  + 
  #q64 line
  geom_line(data = df_quartiles, aes(x = post_years, y = Max_q64, group = 1), linetype = 'dotted', colour = "black", size = 0.5)  

#maybe also add a .50 median line?
```
Graphs of mean percent change - old way that just looks at total percent change (ie one value of mean for post, pre, and then subtracting those). But keep this bargraph code to use later on 
(IGNORE BUT KEEP)
```{r, include = FALSE}

#reshape percent change df 
test <- data.frame(t(m)) %>%
  rename("Value" =  "t.m." ) %>%
  rownames_to_column("Variable")

#use this to round in case want to label bars
test$Value <- round (test$Value, digits = 0)

#barplot of mins and maxes percent change
ggplot(test, aes( x = Variable, y = Value) ) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ylab("Percent change") +
  xlab("IHA variable") +
  scale_x_discrete(limits = c('1 Day Min', '3 Day Min', '7 Day Min','30 Day Min', '90 Day Min', '1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max')) #+


```

```{r}

#barplot of percent change by month

ggplot(test, aes( x = Variable, y = Value) ) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ylab("Percent change") +
  xlab("IHA variable") +
scale_x_discrete(limits = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September'))

```

##Percent change: 

Calculate percent change, using all the annual values: (Post-dam annual values - pre-dam mean values) / pre-dam mean values
```{r}

#First get the mean output for the pre-dam IHA outputs
IHA_output_pre_mean <- IHA_output_pre %>%
  summarise(across(October:Reversals, mean)) 

#then make the IHA pre-dam mean dataframe the same length as the IHA post-dam dataframe
IHA_output_pre_mean_all <- IHA_output_pre_mean[rep(1, nrow(IHA_output_post)),]

#get rid of the time period and year
IHA_output_post_calc <- IHA_output_post %>%
  select(-c(Year, Time_Period))

#make a subset of the "Year" column for later
year_post <- IHA_output_post["Year"]

#do the calculation: IHA_output_postdam_annualvalues - IHA_output_pre_mean/IHA_output_pre_mean
percent_change_output <- ((IHA_output_post_calc - IHA_output_pre_mean_all)/IHA_output_pre_mean_all) * 100

#add the year df and percent change output df together
percent_change_years <- merge(year_post, percent_change_output, by.x = 0, by.y = 0)

```


Now, put all this into a function: 
```{r}

group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)

## Create output dataframe 
IHA_output <- bind_cols(list(group1_output, group2_output, group3_output, group4_output, group5_output))


#takes inputs for the IHA output dataframe for pre and post
percent_change <- function(datapre, datapost){
  
  IHA_output_pre_mean <- {{datapre}} %>%
  summarise(across(October:Reversals, mean)) 
  IHA_output_pre_mean_all <- IHA_output_pre_mean[rep(1, nrow({{datapost}})),]
  
  #good for error checking here, if datapost has year, timeperiod, then don't run this section of code
  IHA_output_post_calc <- {{datapost}} %>%
  select(-c(Year, Time_Period))
  
  year_post <- {{datapost}}["Year"]
  percent_change_output <- ((IHA_output_post_calc - IHA_output_pre_mean_all)/IHA_output_pre_mean_all) * 100
percent_change_years <- merge(year_post, percent_change_output, by.x = 0, by.y = 0)

}


test <- compute_percent_change(IHA_output_pre, IHA_output_post)
#First get the mean output for the pre-dam IHA outputs

#then make the IHA pre-dam mean dataframe the same length as the IHA post-dam dataframe

#get rid of the time period and year


#make a subset of the "Year" column for later

#do the calculation: IHA_output_postdam_annualvalues - IHA_output_pre_mean/IHA_output_pre_mean

#add the year df and percent change output df together

```

Graphs of the percent change output:

```{r}

ggplot() +
  geom_line(data = percent_change_years, aes(x = Year, y = Max), group = 1, colour = "red", size = 0.75)  + 
      geom_point() +
  annotate(geom="text", x=16, y=55, label="Maximum flow", color="red") +
  
   geom_line(data = percent_change_years, aes(x = Year, y = Min), group = 1, colour = "blue", size = 0.75)  + 
      geom_point() +
    annotate(geom="text", x=15, y=250, label="Minimum flow", color="blue") +

  
  ylab("Percent Change")

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

