
## Fort Smith Indicators of Hydrologic Alteration

This is an example of the data collection, manipulation, analysis and graphing for IHA variables for a station on the Slave River.

Indicators of Hydrologic Alteration (IHA) variables are a series of ecologically relevant calculations that can be used to understand changes in river hydrology over time.

This data is from the Water Survey of Canada database, and has historical and current hydrometric data from hundreds of stations across Canada. 

Here we have tidied, cleaned, calculated IHA variables, and graphed a few of them to show hydrometric variables before and after the installation of the Bennett Dam. 




```{r message=FALSE, warning=FALSE, echo=FALSE}
#install.packages("tidyhydat")
#install.packages("IHA")
#install.packages("Rtools")
#install.packages("IHA", repos="http://R-Forge.R-project.org")
#install.packages("plyr")
#install.packages("caTools")
#install.packages("tidyverse")
#install.packages("zoo")
#install.packages("lubridate")
#install.packages("png")
#install.packages("reticulate")
#install.packages("miniconda")
#install.packages("gridExtra")
```



```{r message=FALSE, warning=FALSE, echo=FALSE}
library(tidyhydat)
library(IHA)
library(png)
library(dplyr)
library(tidyverse)
library(reticulate)
library(zoo)
library(lubridate)
library(ggplot2)
library(sf)
library(mapview)
library(gridExtra)
```



```{r message=FALSE, warning=FALSE, echo=FALSE}
#run this once
#download_hydat()
```

```{r, echo=FALSE}
flow_data_predam_FS <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit()%>%
  filter(Date < "1966-10-01" & Date >= "1959-10-01") %>% 
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}

flow_data_postdam <- function(df,Date, Value) {
library(dplyr)
df <- df %>%
  select(Date, Value) %>%
  na.omit() %>%
  filter(Date >= "1968-10-01" & Date < "1991-10-01") %>%
  mutate(Date = format(Date,"%d-%m-%Y"))
  return(df)
}
```


```{r, echo=FALSE}
Fort_Smith_Flows <- tidyhydat::hy_daily_flows(station_number = "07NB001")
```


```{r, echo=FALSE}
latlong_07NB001 <- hy_stations(station_number= "07NB001" ) %>% select(LATITUDE, LONGITUDE)
```


#### Map of Station 07NB001 

We chose this station because it's near Fort Smith along the Slave River.
```{r, echo=FALSE}
sdf <- st_as_sf(latlong_07NB001, coords = c("LONGITUDE", "LATITUDE"),  crs = 4326)
mapview(sdf, map.types = "CartoDB.Positron")
```


```{r, echo=FALSE}
FL_FS_predam <- flow_data_predam_FS(Fort_Smith_Flows) 

FL_FS_postdam <- flow_data_postdam(Fort_Smith_Flows)
```


```{r, echo=FALSE}

## Convert data to zoo object to prep for IHA analyses
flow_data <- zoo(FL_FS_predam$Value, order.by = as.Date(as.character(FL_FS_predam$Date), format = "%d-%m-%Y"))

## Run IHA analyses
group1_output <- group1(flow_data, year = "water", FUN = median)
group2_output <- group2(flow_data, year = "water", mimic.tnc = TRUE)
group3_output <- group3(flow_data, year = "water", mimic.tnc = FALSE)
group4_output <- group4(flow_data, year = "water")
group5_output <- group5(flow_data, year = "water")

## Convert outputs
group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)

#deleting last years because it has an extra year not in the data
group1_output <- group1_output[-15,]
group2_output <- group2_output[-15,]
group3_output <- group3_output[-15,]
group4_output <- group4_output[-15,]
group5_output <- group5_output[-15,]

## Create output dataframe 
IHA_output <- bind_cols(list(group1_output, group2_output, group3_output, group4_output, group5_output))

library(tibble)
IHA_output_predam <- tibble::rownames_to_column(IHA_output, "Year") %>%
  add_row(Year = c("1968", "1967") )%>%
  add_column(Time_Period = "Pre Dam" )

```


```{r echo=FALSE}
## Convert data to zoo object to prep for IHA analyses
flow_data <- zoo(FL_FS_postdam$Value, order.by = as.Date(as.character(FL_FS_postdam$Date), format = "%d-%m-%Y"))

## Run IHA analyses
group1_output <- group1(flow_data, year = "water", FUN = median)
group2_output <- group2(flow_data, year = "water", mimic.tnc = TRUE)
group3_output <- group3(flow_data, year = "water", mimic.tnc = FALSE)
group4_output <- group4(flow_data, year = "water")
group5_output <- group5(flow_data, year = "water")

## Convert outputs
group1_output <- as.data.frame(group1_output)
group2_output <- group2_output[,-1]
group3_output <- as.data.frame(group3_output)
group4_output <- as.data.frame(group4_output)
group5_output <- as.data.frame(group5_output)

## Create output dataframe 
IHA_output <- bind_cols(list(group1_output, group2_output, group3_output, group4_output, group5_output))

library(tibble)
IHA_output_postdam <- tibble::rownames_to_column(IHA_output, "Year") %>%
  add_column(Time_Period = "Post Dam" )
```

#### IHA ANALYSES 
To calculate the IHA variables, we used the IHA package from the Nature Conservancy.

This is what some of the output variables look like: 

```{r, echo=FALSE}
IHA_both <- rbind(IHA_output_predam, IHA_output_postdam) %>%
  add_row(Year = c("1968", "1967") ) 
head(IHA_both, 2)

```



#### Flow Reversals, pre and post dam

All other IHA variables can be graphed this way.

```{r, warning = FALSE, echo=FALSE}

ggplot(data = IHA_both, aes(x =Year, y = Reversals, group = 1) ) +
          geom_line(aes(colour = Time_Period) ) +
      geom_point(aes(colour = Time_Period)) +
  #ggtitle( "07NB001 \nReversals") +
  ylab("Number of Reversals")+
  theme_bw() +
   theme(plot.title =element_text(size = 10, hjust = .5, face = "bold")) +
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.position=c(0.87, 0.8), legend.title=element_text(size=8), legend.text=element_text(size=7)) +
   scale_colour_discrete(na.translate = F)
  
```

#### Minimum flows  
```{r, echo=FALSE}

cols <- c("orange", "light blue")

IHA_both %>% 
  na.omit() %>%
  select('Year', '1 Day Min', '3 Day Min', '7 Day Min', '30 Day Min', '90 Day Min', 'Time_Period') %>%
  pivot_longer(., cols = c('1 Day Min', '3 Day Min', '7 Day Min','30 Day Min', '90 Day Min'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period), levels = c('1 Day Min', '3 Day Min', '7 Day Min','30 Day Min' , '90 Day Min')) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
  stat_boxplot(geom = "errorbar", width = 0.25) + 
  geom_boxplot() + 
scale_x_discrete(limits = c('1 Day Min', '3 Day Min', '7 Day Min','30 Day Min', '90 Day Min')) +
  scale_fill_manual(values = cols) +  # Fill colors
#  geom_point() +
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
       theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
 # scale_fill_manual(values = c("orange","light blue"), 
  #                 labels = c("Pre Dam", "Post Dam")) +
  xlab("") +
  ylab("Flow") 
```

#### Maximum flows  
```{r, echo=FALSE}

cols <- c("orange", "light blue")

IHA_both %>% 
  na.omit() %>%
  select('Year', '1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max','Time_Period') %>%
  pivot_longer(., cols = c('1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max', '90 Day Max'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
  stat_boxplot(geom = "errorbar", width = 0.25) + 
  geom_boxplot() + 
scale_x_discrete(limits = c('1 Day Max', '3 Day Max', '7 Day Max', '30 Day Max' , '90 Day Max')) +
  scale_fill_manual(values = cols) +  # Fill colors
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
 # scale_fill_manual(values = c("orange","light blue"), 
  #                  labels = c("Pre Dam", "Post Dam")) +
  xlab("") +
  ylab("Flow") 

```


#### Monthly flows
```{r, echo=FALSE}

col <- c("light blue", "orange")

IHA_both %>% 
  na.omit() %>%
  select('Year', 'October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September', 'Time_Period') %>%
  pivot_longer(., cols = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September'), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = Time_Period)) +
  facet_grid(~factor(Time_Period, levels=c('Pre Dam','Post Dam'))) +
  stat_boxplot(geom = "errorbar", width = 0.25) + 
  geom_boxplot() + 
scale_x_discrete(limits = c('October', 'November', 'December', 'January' , 'February', 'March', 'April', 'May', 'June','July', 'August', 'September')) +
  scale_fill_manual(values = cols) +  # Fill colors
  theme_bw() +
    theme(plot.title =  element_text(size = 10, hjust = .5, face = "bold")) +
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle( "Station 07NB001") +
    theme(legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  theme(legend.title=element_text(size=8), legend.text=element_text(size=7))+
    aes(x = fct_inorder(Var)) +
  xlab("Month") +
  ylab("Flow") 
 
```
