# Freeze and Thaw Timing for Sites on the Peace-Athabasca Delta (PAD) 

Data is from hydrometric stations managed by the Water Survey of Canada. Stations tidied from database using tidyhydat in R. 

Workflow for the calculations and graphing for the freeze-up dates, freshet onset, and continuous ice coverage for a couple of stations on the PAD. 

Functions for these calculations can be found in "Ice_Variable_Functions.R"

 
```{r, warning= FALSE, include= FALSE}
library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)
library(grid)
library(gridExtra)
library(viridis)
library(sqldf)
library(caTools)
library(dataRetrieval)
#install.packages("IHA", repos="http://R-Forge.R-project.org", dependencies = TRUE)

```

#1) get data 

```{r}

stn <- tidyhydat::hy_daily_flows(station_number = c('07BE001', '07HA001'))

#all years greater than 1971 (only seasonal data before 1971)
stn <- stn %>%
  mutate(Date = ymd(Date), Year = format_ISO8601(Date, precision = "y")) %>%
  filter(Year >= 1971)

```


```{r, echo = FALSE}
#Function to Calculate Day of the Year

#function sequences by number of days in each water year
  
day_of_year_calc <- function(data){
  
  df <- data.frame(waterYear = character(), sequence = character())
  
  for (i in unique({{data}}$waterYear)){
    df_subset <- {{data}}[{{data}}$waterYear == i,]
    days <- seq(1:nrow(df_subset))
    temp_df <- data.frame(waterYear = i, day_of_year = days)
    df <- rbind(df, temp_df)
  }
    df2 <- cbind(df, {{data}}) 
    df3 <- df2[, !duplicated(colnames(df2))]

  return(df3)
}

```

#2) Calculate water year and day of the water year 
```{r} 

#add water year
stn_wy <- addWaterYear(stn) %>%
    mutate(waterYear = as.character(waterYear)) %>%
  select(-Year)

#apply the doy calc by station number. If there is only one station, this function can be applied directly to the df.
stn_wy_split <- split(stn_wy, stn_wy$STATION_NUMBER )
stn_wy_split_doy <- lapply(stn_wy_split, day_of_year_calc)
stn_wy_doy <- bind_rows(stn_wy_split_doy, .id = "Station_Number") 

stn_wy_doy <- stn_wy_doy %>%
  select(-c(STATION_NUMBER))

rm(stn_wy_split, stn_wy_split_doy, stn_wy)

```

```{r, echo = FALSE, eval = FALSE} 
#test stations- just one year

stn_07BK007 <- tidyhydat::hy_daily_flows(station_number = '07BK007')

stn_07BK007_1972 <- stn_07BK007 %>%
  #get a year column
mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y")) %>%
  #just 1972 for now 
  filter(Year == "1972")

rm(stn_07BK007, stn_07BK007_1972)

stn_07BE001 <- stn_wy_doy %>% #"07BE001" "07HA001"
filter(Station_Number == "07BE001")

stn_07HA001 <- stn_wy_doy %>% #"07BE001" "07HA001"
filter(Station_Number == "07HA001")


```


```{r, echo = FALSE}
#Group 1 Function: Ice coverage  

Group_1_ice_cover <- function(data) {
#function returns a df for the length of ice coverage, per calendar year
#works with one station, multiple years
#for multiple stations, split-apply-combine station
  lst <- list()

    for (i in unique({{data}}$waterYear)) {
  
      length_B_date  <- max(rle({{data}}$Symbol[{{data}}$waterYear == i] == "B")[[1]]) 
  #append each value to a list
      length_B_date <- length_B_date - 1
      lst[[i]] <- length_B_date
    }
  
  Ice_coverage_wy <- data.frame(waterYear = names(lst), Ice_coverage_wy = unlist(lst))
  rownames(Ice_coverage_wy) <-NULL
  return(Ice_coverage_wy)
  
}

#test
#Ice_coverage_07BE001 <- Group_1_ice_cover(stn_07BE001)
#edit function so that it takes water year and calendar year arguments

```


```{r, echo= FALSE}
#Group 2 Function: Freeze and Thaw Dates

Group_2_freeze_thaw <- function(data) {
  #This function calculates the freeze and thaw dates and their flow values, using the     longest consecutive run of "B" Symbols in the df

  start_date_lst <- list()
  end_date_lst <- list()
  start_flow_lst <- list()
  end_flow_lst <- list()
  start_doy_lst <- list()
  end_doy_lst <- list()

  for (i in unique({{data}}$waterYear)){
 
    df_subset <- {{data}}[{{data}}$waterYear == i,]
  
    #calc rle for B symbol
    rle_m = rle(df_subset$Symbol == "B")
  
    #find index for max run of B symbols
    max_run_index <- which.max(rle_m$lengths)
  
    #find end start and index of max run of B symbols
    end <- cumsum(rle_m$lengths)[max_run_index]
    start <- end - rle_m$lengths[max_run_index] +1

    #find date at end, start index of the max run 
    date_end = df_subset$Date[end]
    date_start = df_subset$Date[start]
  
    #find flow at end, start index
    flow_end = df_subset$Value[end]
    flow_start = df_subset$Value[start]
    
    #find doy at end, start index
    doy_end = df_subset$day_of_year[end]
    doy_start = df_subset$day_of_year[start]
    
    #append dates to the list
    start_date_lst[[i]] <- date_start
    end_date_lst[[i]] <- date_end
  
    #append flows to the list
    start_flow_lst[[i]] <- flow_start
    end_flow_lst[[i]] <- flow_end
    
    #append doy to the list
    start_doy_lst[[i]] <- doy_start
    end_doy_lst[[i]] <- doy_end
    
  }

    Freeze_Date <- as.Date(unlist(start_date_lst))
    Thaw_Date <- as.Date(unlist(end_date_lst))
    Flow_Freeze <- unlist(start_flow_lst)
    Flow_Thaw <- unlist(end_flow_lst)
    Freeze_DOY <-unlist(start_doy_lst)
    Thaw_DOY <- unlist(end_doy_lst)

    df <- cbind.data.frame(Freeze_Date,Freeze_DOY,Flow_Freeze,Thaw_Date,Thaw_DOY, Flow_Thaw)

    Ice_coverage_dates_flow <- rownames_to_column(df, "waterYear")
    return(Ice_coverage_dates_flow)
}

#test
#Fr_thaw_stn_07BE001 <- Group_2_freeze_thaw(stn_07BE001)
#edit function so that it takes water year and calendar year arguments
#edit function for error checking

```


```{r, echo = FALSE}
#Group 3 Function: Onset of Freshet

Group_3_freshet <- function(data) {
  index <- 0
  f_index <- 16
  date_lst <- list()
  flow_lst <- list()
  stn_nu <- list()
  doy_lst <- list()

  for (i in unique({{data}}$waterYear)) { #first loop
    #subset data by year, resetting at each year
    index = 0  
    f_index = 16 
    df_subset <- {{data}}[{{data}}$waterYear == i,]
   df_subset <- df_subset %>%
     #data tidying:delete dates before Feb 12, so rolling mean calc starts on March 1
    mutate(Date = as.Date(Date)) %>%
    #filter(month(Date) %in% c(3,4,5,6))
    mutate(new_col = format(Date,"%m-%d")) %>%
    filter(month(Date) >= 2 & month(Date) < 7) %>% 
    filter(!(new_col %in% c("02-01", "02-02", "02-03", "02-04", "02-05", "02-06", "02-07", "02-08", "02-09", "02-10", "02-11")))
   
    #calc rolling 16 day mean
    rollmn <- rollmean(df_subset$Value, k = 16, width = 16)
    #rollmn <- as.data.frame(rollmn)
    
      for (j in rollmn) { #second loop
        #increment index
        index = index + 1
        f_index = f_index + 1
      
        #find rolling mean value at the index and multiply by 1.5
        rollmnvalue <- rollmn[index] #roll mean = 0.81
        rollmnvalue1.5 <- rollmnvalue*1.5 #1.215
        
        #get the flow value at that index
        flowvalue <- df_subset$Value[f_index] #flow = 0.654
        
          if (flowvalue > rollmnvalue1.5 & f_index < 123 ) { #third loop
          #append date, flow value to a list using the index numbers
          dt <- df_subset$Date[f_index]
          fl <- df_subset$Value[f_index]
          st <- df_subset$Station_Number[f_index]
          doy <- df_subset$day_of_year[f_index]
         # print(dt) 
        #  print(fl)
        #  print("end") 
          date_lst[[i]] <- dt 
          flow_lst[[i]] <- fl 
          stn_nu[[i]] <- st
          doy_lst[[i]] <- doy
          Freshet_Date <- as.Date(unlist(date_lst))
          Freshet_Flow <- unlist(flow_lst)
          Station_Number <- unlist(stn_nu)
          Day_of_year <- unlist(doy_lst)
          df <- cbind.data.frame(Station_Number, Freshet_Date, Day_of_year, Freshet_Flow)
          break
          }
        
      }
        
  }
        Freshet_dates_flow <- rownames_to_column(df, "waterYear")
        return(Freshet_dates_flow)
}

```

#3) Calculate and tidy Ice Variables( Group 1 and Group 2) for each station
```{r}

#make a list of dfs- split by station number
stn_wy_doy_split <- split(stn_wy_doy, stn_wy_doy$STATION_NUMBER )

#apply ice variable function to all dfs in the list
stn_wy_doy_g1 <- lapply(stn_wy_doy_split, Group_1_ice_cover)
stn_wy_doy_g2 <- lapply(stn_wy_doy_split, Group_2_freeze_thaw)
#stn_wy_doy_g3 <- lapply(stn_wy_doy_split, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_wy_doy_g1, .id = "Station_Number") 
output2 <- bind_rows(stn_wy_doy_g2, .id = "Station_Number")
#output3 <- bind_rows(stn_wy_doy_g3, .id = "Station_Number")


df_final <- merge(output1, output2)

#Get rid of first and last years, they are always wrong when calculated using water year
df_final<- df_final %>%
  filter(waterYear >= 1972 & waterYear < 2021)

head(df_final, 5)

```


```{r, echo = FALSE, eval = FALSE}

Ice_coverage <- Group_1_ice_cover(stn_wy_doy)
Fr_thaw <- Group_2_freeze_thaw(stn_wy_doy)

#combine results df and remove duplicate rows
Ice_calc_complete <- cbind(Fr_thaw, Ice_coverage)
Ice_calc_complete<- Ice_calc_complete[, !duplicated(colnames(Ice_calc_complete))]

#remove first row,it's always incomplete due to water year cycle
Ice_calc_complete <- Ice_calc_complete[-1,] 
#and remove last year too

#Get rid of first and last years, they are always wrong when calculated using water year
df_final <- df_final[-1,]
x <- as.numeric(nrow(df_final))
df_final <- df_final[-x,]


```

## Days of Total continuous ice coverage per water year

```{r, echo = FALSE}

pal <- c("#E69F00", "#56B4E9","#999999", "#009E73")

df_final %>%
ggplot(aes(y= Ice_coverage_wy, x = factor(waterYear), group = Station_Number, colour = Station_Number )) +
  geom_line(linewidth = .75, linetype = "F1") +
    annotate(geom = "text", x = 45, y = 195, label = "Athabasca River, \n Athabasca", colour = "#E69F00") + #07BE001
  annotate(geom = "text", x = 45, y = 62, label = "Peace River, \n AB", colour = "#56B4E9") +    #07HA001
 scale_x_discrete(breaks = seq(1965, 2021, 5)) +
 ylim(0, 200) +
   theme_classic() +
#ggtitle("Days of Total Continuous Ice Coverage per Water Year") +
  xlab(" ") +
  ylab("Days of Ice Coverage ") +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_manual(values = pal)  +
  theme(legend.position = "None")

```

## Timing of yearly Freeze-up dates by day of water year
("Higher" freeze date means the river freezes later in the season)

```{r, echo = FALSE}

pal <- c("#E69F00", "#56B4E9","#999999", "#009E73")

df_final %>%
ggplot(aes(y= Freeze_DOY, x = factor(waterYear), group = Station_Number, colour = Station_Number )) +
  geom_line(linewidth = .75, linetype = "F1") +
    annotate(geom = "text", x = 45, y = 40, label = "Athabasca River, \n Athabasca", colour = "#E69F00") + #07BE001
  annotate(geom = "text", x = 45, y = 210, label = "Peace River, \n AB", colour = "#56B4E9") +    #07HA001
 scale_x_discrete(breaks = seq(1965, 2021, 5)) +
 ylim(0, 300) +
   theme_classic() +
#ggtitle("Freeze Date: Day of the Water Year ") +
  xlab(" ") +
  ylab("Freeze Date (day of year) ") +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_manual(values = pal)  +
  theme(legend.position = "None")

```

## Timing of yearly Thaw dates by day of water year
("lower" date means the river thaws earlier in the season)

```{r, echo =  FALSE}

pal <- c("#E69F00", "#56B4E9","#999999", "#009E73")

df_final %>%
ggplot(aes(y= Thaw_DOY, x = factor(waterYear), group = Station_Number, colour = Station_Number )) +
  geom_line(linewidth = .75, linetype = "F1") +
    annotate(geom = "text", x = 45, y = 450, label = "Athabasca River, \n Athabasca", colour = "#E69F00") + #07BE001
  annotate(geom = "text", x = 45, y = 325, label = "Peace River, \n AB", colour = "#56B4E9") +    #07HA001
 scale_x_discrete(breaks = seq(1965, 2021, 5)) +
 ylim(300, 450) +
   theme_classic() +
#ggtitle("Thaw Date: Day of the Water Year ") +
  xlab(" ") +
  ylab("Thaw Date (day of year) ") +
  theme(plot.title = element_text(hjust = 0.5) ) +
scale_colour_manual(values = pal)  +
  theme(legend.position = "None")

```


```{r, echo=FALSE, eval=FALSE}

#Rolling 16 day mean calcs - just for one year 

rollmn <- rollmean(stn_07BK007_1972$Value, k = 16, width = 16)
rollmnvalue <- rollmn[1] #0.81
rollmnvalue1.5 <- rollmnvalue*1.5 #1.215
x <- stn_07BK007_1972$Value[16] #0.654

#when x > rollmnvalue1.5, freshet starts. #no freshet yet


```

```{r, echo= FALSE, eval=FALSE}
#for one year, from March 1 onward
#stn_07BK007_1972_summer <- stn_07BK007_1972 %>%
 #filter(Date >= "1972-03-01")

stn_07BK007_wy_doy_no1971 <- stn_07BK007_wy_doy %>%
  filter(waterYear > 1971 & waterYear < 2022)

#View(stn_07BK007_wy_doy_no1971)
```

#freshet calc no function
```{r, echo= FALSE, eval=FALSE}

#calculate rolling mean, not in a function

index <- 0
f_index <- 16
date_lst <- list()
flow_lst <- list()
stn_nu <- list()
doy_lst <- list()

for (i in unique(stn_07HA001$waterYear)) { #first loop
  #subset data by year, resetting at each year
  index = 0  
  f_index = 16 
  df_subset <- stn_07HA001[stn_07HA001$waterYear == i,]
  df_subset <- df_subset %>%
    mutate(Date = as.Date(Date)) %>%
    #filter(month(Date) %in% c(3,4,5,6))
    mutate(new_col = format(Date,"%m-%d")) %>%
    filter(month(Date) >= 2 & month(Date) < 7) %>% 
    filter(!(new_col %in% c("02-01", "02-02", "02-03", "02-04", "02-05", "02-06", "02-07", "02-08", "02-09", "02-10", "02-11")))
 #if (nrow(df_subset) < 123) {next}
  
  #calc rolling 16 day mean
  rollmn <- rollmean(df_subset$Value, k = 16, width = 16)

    for (j in rollmn) { #second loop
      #increment index
      index = index + 1
      f_index = f_index + 1
      
      #find rolling mean value at the index and multiply by 1.5
      rollmnvalue <- rollmn[index] #roll mean = 0.81
      rollmnvalue1.5 <- rollmnvalue*1.5 #1.215
      
      #get the flow value at that index
      flowvalue <- df_subset$Value[f_index] #flow = 0.654
      
        if (flowvalue > rollmnvalue1.5 & f_index < 123) { #third loop
        #append date, flow value to a list using the index numbers
        dt <- df_subset$Date[f_index]
        fl <- df_subset$Value[f_index]
        st <- df_subset$Station_Number[f_index]
        doy <- df_subset$day_of_year[f_index]
        print(dt) 
        print(fl)
        print("end") 
        date_lst[[i]] <- dt 
        flow_lst[[i]] <- fl 
        stn_nu[[i]] <- st
        doy_lst[[i]] <- doy
        Freshet_Date <- as.Date(unlist(date_lst))
        Freshet_Flow <- unlist(flow_lst)
        Station_Number <- unlist(stn_nu)
        Day_of_year <- unlist(doy_lst)
        df <- cbind.data.frame(Station_Number, Freshet_Date, Day_of_year, Freshet_Flow)
        break
        }
     # else() #append column with NA for freshet date, doy, and value?
      
    }
      Freshet_dates_flow1 <- rownames_to_column(df, "waterYear")

}

#debug:
#need to make sure loop 3 is skipped and "NA" if index is greater than 122, ie if freshet doesn't happen
#error checking within function
#water year vs calendar year
#how to handle missing years and not stop loop (ie use "next")
#make sure that if flow never goes above 1.5 times the rm the loop doesn't use the last value but writes "NA"
```


#freshet function - working
```{r}

#Group 3 Function: Onset of Freshet

Group_3_freshet <- function(data) {
  index <- 0
  f_index <- 16
  date_lst <- list()
  flow_lst <- list()
  stn_nu <- list()
  doy_lst <- list()

  for (i in unique({{data}}$waterYear)) { #first loop
    #subset data by year, resetting at each year
    index = 0  
    f_index = 16 
    df_subset <- {{data}}[{{data}}$waterYear == i,]
   df_subset <- df_subset %>%
     
    mutate(Date = as.Date(Date)) %>%
    #filter(month(Date) %in% c(3,4,5,6))
    mutate(new_col = format(Date,"%m-%d")) %>%
    filter(month(Date) >= 2 & month(Date) < 7) %>% 
    filter(!(new_col %in% c("02-01", "02-02", "02-03", "02-04", "02-05", "02-06", "02-07", "02-08", "02-09", "02-10", "02-11")))
   
    #calc rolling 16 day mean
    rollmn <- rollmean(df_subset$Value, k = 16, width = 16)
    #rollmn <- as.data.frame(rollmn)
    
      for (j in rollmn) { #second loop
        #increment index
        index = index + 1
        f_index = f_index + 1
      
        #find rolling mean value at the index and multiply by 1.5
        rollmnvalue <- rollmn[index] #roll mean = 0.81
        rollmnvalue1.5 <- rollmnvalue*1.5 #1.215
        
        #get the flow value at that index
        flowvalue <- df_subset$Value[f_index] #flow = 0.654
        
          if (flowvalue > rollmnvalue1.5 & f_index < 123 ) { #third loop
          #append date, flow value to a list using the index numbers
          dt <- df_subset$Date[f_index]
          fl <- df_subset$Value[f_index]
          st <- df_subset$Station_Number[f_index]
          doy <- df_subset$day_of_year[f_index]
         # print(dt) 
        #  print(fl)
        #  print("end") 
          date_lst[[i]] <- dt 
          flow_lst[[i]] <- fl 
          stn_nu[[i]] <- st
          doy_lst[[i]] <- doy
          Freshet_Date <- as.Date(unlist(date_lst))
          Freshet_Flow <- unlist(flow_lst)
          Station_Number <- unlist(stn_nu)
          Day_of_year <- unlist(doy_lst)
          df <- cbind.data.frame(Station_Number, Freshet_Date, Day_of_year, Freshet_Flow)
          break
          }
        
      }
        
  }
        Freshet_dates_flow <- rownames_to_column(df, "waterYear")
        return(Freshet_dates_flow)
}

#test
Freshet_stn_07BE001<- Group_3_freshet(stn_07BE001)
Freshet_stn_07HA001 <- Group_3_freshet(stn_07HA001)

```

#freshet calc multiple stations - working
```{r}

stn_wy_doy_no1971 <- stn_wy_doy %>%
  filter(waterYear > 1971 & waterYear < 2021)

stn_wy_doy_split <- split(stn_wy_doy, stn_wy_doy$Station_Number )

#apply ice variable function to all dfs in the list
stn_wy_doy_g1 <- lapply(stn_wy_doy_split, Group_1_ice_cover)
stn_wy_doy_g2 <- lapply(stn_wy_doy_split, Group_2_freeze_thaw)
stn_wy_doy_g3 <- lapply(stn_wy_doy_split, Group_3_freshet)

#Combine all outputs into one df
output1 <- bind_rows(stn_wy_doy_g1, .id = "Station_Number") 
output2 <- bind_rows(stn_wy_doy_g2, .id = "Station_Number")
output3 <- bind_rows(stn_wy_doy_g3, .id = "Station_Number")

df_final <- merge(output1, output2)

#Get rid of first and last years, they are always wrong when calculated using water year
df_final<- df_final %>%
  filter(waterYear >= 1972 & waterYear < 2021)

head(df_final, 5)

```



```{r}


```


```{r}


```


