
## Data Availability for Water Survey of Canada Stations

These graphs visualize the availability of water flow and level data for over 400 water survey stations within the Peace Athabasca Delta. This work was done to support the analysis of environmental flows within Wood Buffalo National Park in Canada. 

I graphed these stations to be able to more easily select water stations for analyses. Rivers experience natural variation over time, both seasonally and even from year to year. The timing, volume, length of time, and variability of river flow is called a river's hydrological regime. 

In order to understand how a river or a watershed's hydrological regime is changing over time, you often need multiple points of data along several watersheds, with daily flow measurements over many consecutive years. 

Visualizing which stations have overlapping data in different watersheds will make it much easier for me and others on my team to pick water stations to analyze. 

I graphed these stations by watershed and sorted from upstream to downstream. They are separated by data type: water level and flow.

Open-source hydrological data from the Water Survey of Canada: 
https://www.canada.ca/en/environment-climate-change/services/water-overview/quantity/monitoring/survey.html 

Open-sourced R packages used to tidy data and create tile plots: tidyverse, tidyhydat, zoo, lubridate, ggplot

Source code: https://github.com/Jacqui-123/EFlows-Project/blob/main/Station_summaries.Rmd 

```{r, echo = FALSE, include=FALSE}

library(png)
library(reticulate)
library(IHA)
library(tidyverse)
library(tidyhydat)
library(zoo)
library(lubridate)
library(ggplot2)

```


```{r, echo = FALSE, include=FALSE}
#1) Dataframe of all daily flow for all stations 

ALL_FLOWS <- tidyhydat::hy_daily_flows(station_number = c("07AA001", "07AA002", "07AA003", "07AA004", "07AA007", "07AA008", "07AA009", "07AA010", "07AB002", "07AC001", "07AC002", "07AC003", "07AC004", "07AC005", "07AC006", "07AC007", "07AC008", "07AD001", "07AD002", "07AD003", "07AD004", "07AD005", "07AD006", "07AD007", "07AD008", "07AD009", "07AD010", "07AE001", "07AE002", "07AE003", "07AF001", "07AF002", "07AF003", "07AF004", "07AF005", "07AF008", "07AF009", "07AF010", "07AF011", "07AF012", "07AF013", "07AF014", "07AF015", "07AF016", "07AF906", "07AF907", 
              "07AF909", "07AF910", "07AG001", "07AG002", "07AG003", "07AG004", "07AG005", "07AG006", "07AG007", "07AG008", "07AH001", "07AH002", "07AH003", "07BA001", "07BA002", "07BA003", "07BB001", "07BB002", "07BB003", "07BB004", "07BB005", "07BB006", "07BB007", "07BB008", "07BB009", "07BB010", "07BB011", "07BB012", "07BB013", "07BB014", "07BB903", "07BB914", "07BC001", "07BC002", "07BC003", "07BC004", "07BC005", "07BC006", "07BC007", "07BE001", "07BE002", "07BE003", "07BE004", "07BF001", "07BF002", "07BF004", 
              "07BF006", "07BF007", "07BF008", "07BF009", "07BF010", "07BF905", "07BG004", "07BH001", "07BH003", "07BJ001", "07BJ002", "07BJ003", "07BJ004", "07BJ005", "07BJ006", "07BK001", "07BK002", "07BK003", "07BK004", "07BK005", "07BK006", "07BK007", "07BK008", "07BK009", "07BK010", "07BK012", "07CA001", "07CA002", "07CA003", "07CA004", "07CA005", "07CA006", "07CA008", "07CA010", "07CA011", "07CA012", "07CA013", "07CA901", "07CB001", "07CB002", "07CC001", "07CC002", "07CD001", "07CD002", 
              "07CD003", "07CD004", "07CD005", "07CD006", "07CD007", "07CD008", "07CD009", "07CE001", "07CE002", "07CE003", "07CE004", "07CE005", "07CE006", "07CE007", "07CE008", "07CE010", "07CE013", "07CE906", "07DA001", "07DA002", "07DA003", "07DA004", "07DA005", "07DA006", "07DA007", "07DA008", "07DA009", "07DA010", "07DA011", "07DA012", "07DA013", "07DA014", "07DA015", "07DA016", "07DA017", "07DA018", "07DA019", "07DA020", "07DA021", "07DA022", "07DA023", "07DA024", "07DA025", "07DA026", "07DA027", "07DA028", 
              "07DA029", "07DA030", "07DA032", "07DA033", "07DA034", "07DA035", "07DA037", "07DA038", "07DA039", "07DA040", "07DA041", "07DA042", "07DA044", "07DA045", "07DB001", "07DB002", "07DB003", "07DB004", "07DB005", "07DB006", "07DC001", "07DC002", "07DC003", "07DC004", "07DD001", "07DD002", "07DD003", "07DD004", "07DD005", "07DD006", "07DD007", "07DD008", "07DD009", "07DD010", "07DD011", "07EA001", "07EA002", "07EA004", "07EA005", "07EA006", "07EA007", "07EB001", "07EB002", "07EC001", "07EC002", "07EC003", 
              "07EC004", "07ED001", "07ED002", "07ED003", "07EE002", "07EE003", "07EE004", "07EE005", "07EE007", "07EE008", "07EE009", "07EE010", "07EE011", "07EF001", "07EF002", "07EF003", "07EF004", "07FA001", "07FA003", "07FA004", "07FA005", "07FA006", "07FA007", "07FB001", "07FB002", "07FB003", "07FB004", "07FB005", "07FB006", "07FB007", "07FB008", "07FB009", "07FB010", "07FB011", "07FC001", "07FC002", "07FC003", "07FC004", "07FC005", "07FC006", "07FC007", "07FD001", "07FD002", "07FD003", "07FD004", "07FD005", 
              "07FD006", "07FD007", "07FD008", "07FD009", "07FD010", "07FD011", "07FD012", "07FD013", "07FD014", "07FD015", "07FD016", "07FD017", "07FD019", "07FD020", "07FD901", "07FD908", "07FD910", "07FD912", "07FD913", "07FD921", "07FD934", "07GA001", "07GA002", "07GB001", "07GB002", "07GB003", "07GC001", "07GC002", "07GD001", "07GD002", "07GD003", "07GD004", "07GE001", "07GE002", "07GE003", "07GE004", "07GE005", "07GE006", "07GE007", "07GF001", "07GF002", "07GF003", "07GF004", "07GF005", "07GF006", "07GF007", 
              "07GF008", "07GG001", "07GG002", "07GG003", "07GH001", "07GH002", "07GH003", "07GH004", "07GH005", "07GH906", "07GJ001", "07GJ004", "07GJ005", "07HA001", "07HA002", "07HA003", "07HA005", "07HA902", "07HA914", "07HB001", "07HB002", "07HC001", "07HC002", "07HC907", "07HD001", "07HF001", "07HF002", "07JA001", "07JA002", "07JA003", "07JB001", "07JB002", "07JC001", "07JC002", "07JC003", "07JD001", "07JD002", "07JD003", "07JD004", "07JF002", "07JF003", "07JF004", "07JF005", "07KA002", "07KC001", "07KC003", 
              "07KC004", "07KC005", "07KE001", "07KF001", "07KF002", "07KF003", "07KF004", "07KF005", "07KF006", "07KF007", "07KF008", "07KF010", "07KF013", "07KF014", "07KF015", "07LB001", "07LB002", "07LC002", "07LC003", "07LD001", "07LD002", "07LD003", "07LE001", "07LE002", "07LE003", "07LE004", "07MA003", "07MA004", "07MA005", "07MB001", "07MC001", "07MC002", "07MC003", "07MD001", "07MD002", "07NA001", "07NA002", "07NA003", "07NA004", "07NA005", "07NA007", "07NA008", "07NB001", "07NB002", "07NB003", "07NB004", 
              "07NB005", "07NB006", "07NB007", "07NB008", "07NC001", "07NC002", "07NC003", "07NC004", "07NC005", "07NC006", "07NC007", "07NC008", "07NC009", "07PB002", "07QC008") )

```


```{r, echo = FALSE, include= FALSE}
#Test cleaning code on a subset of stations - this works 

test_flows_daily <- tidyhydat::hy_daily_flows(station_number =  c("07AF909", "07AF910", "07AG001", "07AG002", "07AG003", "07AG004", "07AG005", "07AG006", "07AG007", "07AG008", "07AH001", "07AH002", "07AH003", "07BA001", "07BA002", "07BA003", "07BB001", "07BB002", "07BB003", "07BB004", "07BB005", "07BB006", "07BB007", "07BB008", "07BB009", "07BB010", "07BB011", "07BB012", "07BB013", "07BB014", "07BB903", "07BB914"))

test_flows_daily <- test_flows_daily %>%
group_by(STATION_NUMBER) %>%
  #make a complete set of days for each year 
  complete(Date = seq.Date(as.Date("1912/1/1"), as.Date("2020/12/31"), by="day")) %>%
  #make a separate column for year
  mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y"))
  

test_flows_daily_test <- test_flows_daily %>%
  group_by(STATION_NUMBER, Year) %>%
  count(is.na(Value)) %>%
  rename("Is_NA" = "is.na(Value)")  %>%
  pivot_wider(names_from = Is_NA, values_from = n) %>%
  rename('Number_values' = "FALSE",
         "Number_NAs" = "TRUE" )  %>%
  select(-c(Number_NAs)) %>%
  #na.omit(Number_values) %>%
 mutate("Watershed" = str_extract(STATION_NUMBER, ".{2}(.)") )%>%
mutate_if(is.numeric, ~replace(., is.na(.), 0))

```

```{r, echo = FALSE, include = FALSE}
#Graphing test for one watershed - this works and looks good, now make a function

test_flows_daily_test %>%
  filter(Watershed == "07A") %>%
 ggplot(aes(x = Year, y = STATION_NUMBER, fill = Number_values)) +
  geom_tile(colour = "white", lwd = 0.5, linetype = 1) +
  theme_bw() +
    scale_y_discrete(limits=rev) +
  theme(axis.text.y =  element_text( size = 6, face = 'bold')) +
    scale_x_discrete(breaks = seq(1912, 2023, 2)) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 7, face = 'bold')) +
  theme(axis.text.y =  element_text(size = 6)) +
  #labs(fill = "Number of Days \n in each year") +
    theme(legend.position = 'none') +
  ylab('') +
  scale_fill_gradientn(colours = hcl.colors(5, palette = "Blues", rev = TRUE) )
 
```


```{r, echo = FALSE, include = FALSE}

ALL_FLOWS_y <- ALL_FLOWS %>%
group_by(STATION_NUMBER) %>%
  #make a complete set of days for each year 
  complete(Date = seq.Date(as.Date("1912/1/1"), as.Date("2020/12/31"), by="day")) %>%
  #make a separate column for year
  mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y"))

```


```{r, echo = FALSE, include = FALSE}

ALL_FLOWS <- ALL_FLOWS_y %>%
  group_by(STATION_NUMBER, Year) %>%
  count(is.na(Value)) %>%
  rename("Is_NA" = "is.na(Value)")  %>%
  pivot_wider(names_from = Is_NA, values_from = n) %>%
  rename('Number_values' = "FALSE",
         "Number_NAs" = "TRUE" )  %>%
  select(-c(Number_NAs)) %>%
mutate("Watershed" = str_extract(STATION_NUMBER, ".{2}(.)") ) %>%
mutate_if(is.numeric, ~replace(., is.na(.), 0))

#length(unique(ALL_FLOWS$STATION_NUMBER))

```


```{r, echo = FALSE, include = FALSE}

#Function to create tile plots 

tileplot <- function(data, x, y, fill, z, m, n, o) {
  p <- data %>%
    filter(Watershed == z | Watershed == m | Watershed == n | Watershed == o) %>%
    ggplot(aes({{x}}, {{y}},fill = {{fill}} )) +
  geom_tile( colour = "white", lwd = 0.5, linetype = 1) +
  theme_bw() +
    scale_y_discrete(limits=rev)+
 #scale_x_discrete(guide = guide_axis(n.dodge = 2)) + #dodge every 2
     theme(axis.text.y =  element_text( size = 7, face = 'bold')) +
    scale_x_discrete(breaks = seq(1912, 2022, 2)) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 7, face = 'bold')) +
       # theme(legend.position = 'none') +
    theme(legend.text = element_text(size = 6),
          legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(.5, 'cm'), legend.title = element_text(size=7)) +
    ylab('') +
  scale_fill_gradientn(colours = hcl.colors(5, palette = "Blues3", rev = TRUE) ) +
    labs(fill = "Number of Days \n in each year") 
    #labs(title = {{title}}) +
    print(p)

}


```


## Plots of flow data availability for each watershed:

"07" refers to the main drainage basin (in this case, the Peace-Athabasca Delta), and then the letter after that represents a different watershed. There are 15 different watersheds here, one on each graph: "07A", "07B", "07C", "07D", "07E", "07F", "07G", "07H", "07J", "07K", "07L", "07M", "07N", "07P", "07Q"

```{r,echo = FALSE,  fig.height = 6.0, fig.width= 8}

tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07A', '07A', '07A', '07A' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07B', '07B' ,'07B' ,'07B'  )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07C', '07C', '07C', '07C' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07D', '07D', '07D','07D' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07E', '07E','07E','07E' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07F', '07F' ,'07F' ,'07F' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07G', '07G','07G','07G' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07H', '07J', '07H', '07J')
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07K', '07L', '07K', '07L' )
tileplot(ALL_FLOWS, Year, STATION_NUMBER, Number_values, '07M', '07N', '07P', '07Q' )

```



```{r, echo = FALSE, include = FALSE}

ALL_LEVELS <- tidyhydat::hy_daily_levels(station_number = c("07AA001", "07AA002", "07AA003", "07AA004", "07AA007", "07AA008", "07AA009", "07AA010", "07AB002", "07AC001", "07AC002", "07AC003", "07AC004", "07AC005", "07AC006", "07AC007", "07AC008", "07AD001", "07AD002", "07AD003", "07AD004", "07AD005", "07AD006", "07AD007", "07AD008", "07AD009", "07AD010", "07AE001", "07AE002", "07AE003", "07AF001", "07AF002", "07AF003", "07AF004", "07AF005", "07AF008", "07AF009", "07AF010", "07AF011", "07AF012", "07AF013", "07AF014", "07AF015", "07AF016", "07AF906", "07AF907", 
              "07AF909", "07AF910", "07AG001", "07AG002", "07AG003", "07AG004", "07AG005", "07AG006", "07AG007", "07AG008", "07AH001", "07AH002", "07AH003", "07BA001", "07BA002", "07BA003", "07BB001", "07BB002", "07BB003", "07BB004", "07BB005", "07BB006", "07BB007", "07BB008", "07BB009", "07BB010", "07BB011", "07BB012", "07BB013", "07BB014", "07BB903", "07BB914", "07BC001", "07BC002", "07BC003", "07BC004", "07BC005", "07BC006", "07BC007", "07BE001", "07BE002", "07BE003", "07BE004", "07BF001", "07BF002", "07BF004", 
              "07BF006", "07BF007", "07BF008", "07BF009", "07BF010", "07BF905", "07BG004", "07BH001", "07BH003", "07BJ001", "07BJ002", "07BJ003", "07BJ004", "07BJ005", "07BJ006", "07BK001", "07BK002", "07BK003", "07BK004", "07BK005", "07BK006", "07BK007", "07BK008", "07BK009", "07BK010", "07BK012", "07CA001", "07CA002", "07CA003", "07CA004", "07CA005", "07CA006", "07CA008", "07CA010", "07CA011", "07CA012", "07CA013", "07CA901", "07CB001", "07CB002", "07CC001", "07CC002", "07CD001", "07CD002", 
              "07CD003", "07CD004", "07CD005", "07CD006", "07CD007", "07CD008", "07CD009", "07CE001", "07CE002", "07CE003", "07CE004", "07CE005", "07CE006", "07CE007", "07CE008", "07CE010", "07CE013", "07CE906", "07DA001", "07DA002", "07DA003", "07DA004", "07DA005", "07DA006", "07DA007", "07DA008", "07DA009", "07DA010", "07DA011", "07DA012", "07DA013", "07DA014", "07DA015", "07DA016", "07DA017", "07DA018", "07DA019", "07DA020", "07DA021", "07DA022", "07DA023", "07DA024", "07DA025", "07DA026", "07DA027", "07DA028", 
              "07DA029", "07DA030", "07DA032", "07DA033", "07DA034", "07DA035", "07DA037", "07DA038", "07DA039", "07DA040", "07DA041", "07DA042", "07DA044", "07DA045", "07DB001", "07DB002", "07DB003", "07DB004", "07DB005", "07DB006", "07DC001", "07DC002", "07DC003", "07DC004", "07DD001", "07DD002", "07DD003", "07DD004", "07DD005", "07DD006", "07DD007", "07DD008", "07DD009", "07DD010", "07DD011", "07EA001", "07EA002", "07EA004", "07EA005", "07EA006", "07EA007", "07EB001", "07EB002", "07EC001", "07EC002", "07EC003", 
              "07EC004", "07ED001", "07ED002", "07ED003", "07EE002", "07EE003", "07EE004", "07EE005", "07EE007", "07EE008", "07EE009", "07EE010", "07EE011", "07EF001", "07EF002", "07EF003", "07EF004", "07FA001", "07FA003", "07FA004", "07FA005", "07FA006", "07FA007", "07FB001", "07FB002", "07FB003", "07FB004", "07FB005", "07FB006", "07FB007", "07FB008", "07FB009", "07FB010", "07FB011", "07FC001", "07FC002", "07FC003", "07FC004", "07FC005", "07FC006", "07FC007", "07FD001", "07FD002", "07FD003", "07FD004", "07FD005", 
              "07FD006", "07FD007", "07FD008", "07FD009", "07FD010", "07FD011", "07FD012", "07FD013", "07FD014", "07FD015", "07FD016", "07FD017", "07FD019", "07FD020", "07FD901", "07FD908", "07FD910", "07FD912", "07FD913", "07FD921", "07FD934", "07GA001", "07GA002", "07GB001", "07GB002", "07GB003", "07GC001", "07GC002", "07GD001", "07GD002", "07GD003", "07GD004", "07GE001", "07GE002", "07GE003", "07GE004", "07GE005", "07GE006", "07GE007", "07GF001", "07GF002", "07GF003", "07GF004", "07GF005", "07GF006", "07GF007", 
              "07GF008", "07GG001", "07GG002", "07GG003", "07GH001", "07GH002", "07GH003", "07GH004", "07GH005", "07GH906", "07GJ001", "07GJ004", "07GJ005", "07HA001", "07HA002", "07HA003", "07HA005", "07HA902", "07HA914", "07HB001", "07HB002", "07HC001", "07HC002", "07HC907", "07HD001", "07HF001", "07HF002", "07JA001", "07JA002", "07JA003", "07JB001", "07JB002", "07JC001", "07JC002", "07JC003", "07JD001", "07JD002", "07JD003", "07JD004", "07JF002", "07JF003", "07JF004", "07JF005", "07KA002", "07KC001", "07KC003", 
              "07KC004", "07KC005", "07KE001", "07KF001", "07KF002", "07KF003", "07KF004", "07KF005", "07KF006", "07KF007", "07KF008", "07KF010", "07KF013", "07KF014", "07KF015", "07LB001", "07LB002", "07LC002", "07LC003", "07LD001", "07LD002", "07LD003", "07LE001", "07LE002", "07LE003", "07LE004", "07MA003", "07MA004", "07MA005", "07MB001", "07MC001", "07MC002", "07MC003", "07MD001", "07MD002", "07NA001", "07NA002", "07NA003", "07NA004", "07NA005", "07NA007", "07NA008", "07NB001", "07NB002", "07NB003", "07NB004", 
              "07NB005", "07NB006", "07NB007", "07NB008", "07NC001", "07NC002", "07NC003", "07NC004", "07NC005", "07NC006", "07NC007", "07NC008", "07NC009", "07PB002", "07QC008") )


```


```{r, echo = FALSE, include = FALSE}

ALL_LEVELS <- ALL_LEVELS %>%
group_by(STATION_NUMBER) %>%
  #make a complete set of days for each year 
  complete(Date = seq.Date(as.Date("1912/1/1"), as.Date("2020/12/31"), by="day")) %>%
  #make a separate column for year
  mutate(Date = ymd(Date),  Year = format_ISO8601(Date, precision = "y"))


```


```{r, echo = FALSE, include = FALSE}

ALL_LEVELS <- ALL_LEVELS %>%
  group_by(STATION_NUMBER, Year) %>%
  count(is.na(Value)) %>%
  rename("Is_NA" = "is.na(Value)")  %>%
  pivot_wider(names_from = Is_NA, values_from = n) %>%
  rename('Number_values' = "FALSE",
         "Number_NAs" = "TRUE" )  %>%
  select(-c(Number_NAs)) %>%
 mutate("Watershed" = str_extract(STATION_NUMBER, ".{2}(.)") ) %>%
mutate_if(is.numeric, ~replace(., is.na(.), 0))

#length(unique(ALL_LEVELS$STATION_NUMBER))
```


```{r, echo = FALSE, include = FALSE}

tileplot <- function(data, x, y, fill, z, m, n, o) {
  p <- data %>%
    filter(Watershed == z | Watershed == m | Watershed == n | Watershed == o) %>%
    ggplot(aes({{x}}, {{y}},fill = {{fill}} )) +
  geom_tile( colour = "white", lwd = 0.5, linetype = 1) +
  theme_bw() +
    scale_y_discrete(limits=rev)+
 #scale_x_discrete(guide = guide_axis(n.dodge = 2)) + #dodge every 2
    theme(axis.text.y =  element_text( size = 7, face = 'bold')) +
    scale_x_discrete(breaks = seq(1912, 2022, 2)) +
    theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size= 7, face = 'bold')) +
       # theme(legend.position = 'none') +
    theme(legend.text = element_text(size = 6),
          legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(.5, 'cm'), legend.title = element_text(size=7)) +
    ylab('') +
  scale_fill_gradientn(colours = hcl.colors(5, palette = "Purples", rev = TRUE) ) +
    labs(fill = "Number of Days \n in each year") 
    #labs(title = {{title}}) +
    print(p)    }

```

## Plots of water level data availability for each watershed: 
There are 15 different watersheds here, one on each graph: "07A", "07B", "07C", "07D", "07E", "07F", "07G", "07H", "07J", "07K", "07L", "07M", "07N", "07P", "07Q"

```{r,echo = FALSE, fig.height = 5.0, fig.width= 8}
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07A', '07A', '07A', '07A' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07B' , '07B','07B' ,'07B' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07C', '07C', '07C', '07C' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07D', '07D', '07D','07D' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07E', '07E','07E','07E'  )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07F', '07F' ,'07F' ,'07F' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07G', '07G','07G','07G' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07H', '07J', '07H', '07J' )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07K', '07L', '07K', '07L'  )
tileplot(ALL_LEVELS, Year, STATION_NUMBER, Number_values, '07M', '07N', '07N', '07Q' )
```


